### 2017-03-19 latex/paper.tex
Initial commit

\documentclass[Afour,sageh,times]{sage/sagej}

\usepackage{moreverb,url}
\usepackage{csvsimple}
\usepackage[colorlinks,bookmarksopen,bookmarksnumbered,citecolor=red,urlcolor=red]{hyperref}

\bibliographystyle{sage/SageV}

\newcommand\BibTeX{{\rmfamily B\kern-.05em \textsc{i\kern-.025em b}\kern-.08em
T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

\def\volumeyear{2017}

\begin{document}
\runninghead{Boeing}
\title{A Multi-Scale Analysis of 27,000 Urban Street Networks}
\author{Geoff Boeing\affilnum{1}}
\affiliation{\affilnum{1}Department of City and Regional Planning, University of California, Berkeley}
\corrauth{Geoff Boeing,
Department of City and Regional Planning,
University of California, Berkeley,
United States}
\email{gboeing@berkeley.edu}

\begin{abstract}
This paper serves illustrative purposes as it presents wide-ranging preliminary empirical findings on U.S. urban form and street network characteristics, using OSMnx. It also demonstrates the scalability and flexibility of OSMnx as a new research tool. This paper addresses past limitations by conducting an analysis of street networks with large sample sizes, with clearly defined network definitions and extents for reproducibility, and using non-planar, directed graphs. In particular, it examines urban street networks focusing on structure, connectedness, centrality, and robustness. In total it cross-sectionally analyzes 497 urbanized areas' street networks, 19,655 cities' and towns' street networks, and 6,857 neighborhoods' street networks. These sample sizes are larger than those in any previous study and the preliminary empirical findings illustrate the use of OSMnx as a new research platform.
\end{abstract}

\keywords{network analysis, street network, urban design, urban form, urban planning}

\maketitle

\section{Introduction}
On May 20, 1862, Abraham Lincoln signed the Homestead Act into law, making land across the United States Midwest and Great Plains available for free to all applicants \cite{porterfield_homestead_2005}. Under its auspices over the next 70 years, the federal government distributed 270 million acres of public land (10\% of the entire U.S. landmass) to private owners in the form of 1.6 million homesteads \cite{lee_kansas_1979, sherraden_inclusion_2005}. Towns with gridiron street networks sprang up rapidly across the Great Plains and Midwest, due to both the prevailing urban design paradigm of the day and the standardized rectilinear town plats used repeatedly to lay out instant new cities (Southworth and Ben-Joseph 1997). Through path dependence, the spatial signatures of these land use laws, design paradigms, and planning instruments can still be seen today in these cities' urban forms and street networks.

A cross-sectional analysis of American urban form through its street networks at metropolitan, municipal, and neighborhood scales can reveal these artifacts and histories across the nation. Network analysis is a natural approach to the study of cities as growing, complex systems that self-organize to fill space with a capillary circulation network (Masucci et al. 2009). The empirical literature on street network analysis is growing ever richer, but suffers from some limitations. First, sample sizes tend to be fairly small due to data availability, gathering, and processing constraints. Second, reproducibility is difficult when the dozens of decisions that go into analysis – such as spatial extents, topological simplification and correction, definitions of nodes and edges, etc. – are ad hoc or only partly reported. Third, and related to the first two, studies frequently oversimplify to planar or undirected primal graphs for tractability, or use dual graphs despite the loss of geographic and metric information. This paper addresses these limitations by conducting an analysis of street networks at multiple scales, with large sample sizes, with clearly defined network definitions and extents for reproducibility, and using non-planar, directed graphs. In particular, it examines urban street networks – represented here as primal, non-planar, weighted multidigraphs with possible self-loops – focusing on structure, connectedness, centrality, and resilience.

Most studies in the street network literature that conduct topological and/or metric analysis tend to have sample sizes ranging around 5 to 50 networks. This paper instead conducts a large analysis of over 27,000 urban street networks at multiple overlapping scales across the United States. Namely, it examines the street networks of every U.S. incorporated city and town, urbanized area, and Zillow-defined neighborhood. In total, this study presented uses OSMnx to download, construct, and analyze 497 urbanized areas' street networks, 19,655 cities' and towns' street networks, and 6,857 neighborhoods' street networks. It uses these street networks to conduct four analyses: at the metropolitan scale, at the municipal scale, at the neighborhood scale, and with a case study looking deeper at the neighborhood-scale street networks in the city of San Francisco.

This paper is organized as follows. In the next section, it presents the data sources, tools, and methods used to collect, construct, and analyze these street networks. Then it presents findings of the analyses at three spatial scales: metropolitan, municipal, and neighborhood. Next it presents a case study of the neighborhood scale in San Francisco. Finally, it concludes with a discussion of these findings and their implications for street network analysis, urban form complexity, and city planning.

\section{Methodology}
This study uses OSMnx to download, construct, correct, analyze, and visualize street networks at metropolitan, municipal, and neighborhood scales. To define the study sites and their spatial boundaries, this study uses three sets of geometries. The first defines the metropolitan-scale study sites using the 2016 national TIGER/Line shapefile of U.S. census bureau urban areas, released 19 August 2016. Census-defined urban areas comprise a set of census tracts that meet a minimum density threshold (U.S. Census Bureau 2010). In particular, we retain only the urbanized areas subset in this data set (i.e., areas with greater than 50,000 population), discarding the smaller urban clusters subset. The second set of geometries defines the municipal-scale study sites using 51 TIGER/Line shapefiles (again, 2016) of U.S. census bureau places for all 50 states plus the District of Columbia. In particular, we discard the subset of census-designated places in this data set, instead retaining all the incorporated cities and towns in the United States.

The third and final set of geometries defines the neighborhood-scale study sites using 42 Zillow neighborhood shapefiles. These shapefiles contain the geometries of neighborhoods in 41 states plus the District of Columbia. This is a fairly new data set comprising nearly 7,000 neighborhoods in large American cities, but as Schernthanner et al. (2016) point out, Zillow does not publish the methodology it uses to construct these geometries. However, despite its newness it already has a track record in the academic literature. For instance, Besbris et al. (2015) used Zillow neighborhood boundaries to examine neighborhood stigma, and Albrecht and Abramowitz (2014) used these data to study neighborhood-level poverty in New York.

For all of these geometries, we use OSMnx to download the (drivable, public) street network contained within the geometry. First it buffers each geometry by 0.5 km, then downloads the street network within this buffered geometry. Next it constructs a street network from this data, corrects the topology, calculates street counts per node (this ensures that intersections are not considered cul-de-sacs simply because an incident edge connects to a node outside the desired polygon), then truncates the network to the original, desired polygon. OSMnx saves each of these networks to disk as GraphML and shapefiles. Finally, it automatically calculates metric and topological measures for each. However, in a planar graph, topological and metric properties are somewhat interrelated (Masucci et al. 2009).

Average street length, the mean edge length in the undirected representation of the graph, serves as a linear proxy for block size and indicates how fine-grained or coarse-grained the network is. Node density is the number of nodes divided by the area covered by the network. Intersection density is the node density of the set of nodes with more than one street emanating from them (thus excluding dead-ends). The edge density is the sum of all edge lengths divided by the area, and the physical street density is the sum of all edges in the undirected representation of the graph divided by the area. These density measures all provide further indication of how fine-grained the network is. Finally, the average circuity represents the average ratio between an edge length and the straight-line distance between the two nodes it links.

The average node degree, or mean number of edges incident to each node, quantifies how well the nodes are connected on average. Similarly, but more concretely, the average streets per node measures the mean number of physical streets (i.e., edges in the undirected representation of the graph) that emanate from each intersection and dead-end. This adapts the average node degree for physical form rather than directed circulation. The statistical and spatial distributions of number of streets per intersection characterize the type, prevalence, and dispersion of intersection connectedness in the network. Connectivity measures the minimum number of nodes or edges that must be removed from a connected graph to disconnect it. The average node connectivity of a network – the mean number of internally node-disjoint paths between each pair of nodes in the graph – represents the expected number of nodes that must be removed to disconnect a randomly selected pair of non-adjacent nodes (Beineke et al. 2002).

The clustering coefficient of a node is the ratio of the number of edges between its neighbors to the maximum possible number of edges that could exist between these neighbors. The weighted clustering coefficient weights this ratio by edge length and the average clustering coefficient is the mean of the clustering coefficients of all the nodes in the network. Centrality indicates the importance of nodes in a network. Betweenness centrality evaluates the number of shortest paths that pass through each node (Barthélemy 2004). The maximum betweenness centrality in a network specifies the proportion of shortest paths that pass through the most important node. This is an indicator of resilience: networks with a high maximum betweenness centrality are more prone to failure or inefficiency should this single choke point fail. Closeness centrality ranks nodes as more central if they are on average closer to all other nodes. Finally, PageRank ranks nodes based on the structure of incoming links and the rank of the source node.

In total, this study cross-sectionally analyzes 27,009 networks: 497 urbanized areas' street networks, 19,655 cities' and towns' street networks, and 6,857 neighborhoods' street networks. These sample sizes are larger than those in any previous study. In the following sections, we present the findings of these analyses at the metropolitan scale, the municipal scale, the neighborhood scale, and through a case study looking deeper at the neighborhood-scale street networks in the city of San Francisco.

\section{Analysis of Metropolitan-Scale Street Networks}

\begin{table*}
\caption{Measures of central tendency and dispersion for selected measures of the 497 urbanized area street networks.}
\label{table01}
\csvautotabular{media/table01.csv}
\end{table*}

Table \ref{table01} shows the considerable variation in street network characteristics across the entire data set of 497 urbanized areas. This is unsurprising: urbanized areas span a wide spectrum of sizes, from the Delano, CA Urbanized Area's 26 km2 to the New York--Newark, NY--NJ--CT Urbanized Area's 8,937 km2 – thus, density and count-based measures demonstrate substantial variance. Further, these urbanized areas span a wide spectrum of terrains, development eras and paradigms, and cultures.

\begin{figure}
\includegraphics[width=0.45\textwidth]{media/fig01.png}
\caption{Intersection density per urbanized area (top) and average streets per node per urbanized area (bottom) from lowest (pale yellow) to highest (dark red), in the contiguous U.S.}
\label{fig01}
\end{figure}

Nevertheless, looking across the data set provides a sense of the breadth of American metropolitan street networks. New York's urbanized area – America's largest – has 373,309 intersections and 79 million meters of linear street (or 417,570 and 83.4 million if you include service roads). Delano, CA's urbanized area – America's smallest – has 874 intersections and 222,328 meters of linear street (or 964 and 231,000 meters if you include service roads). The typical (Table \ref{table01}, \emph{median}) American urbanized area is approximately 185 km2 in land area, has 5,830 intersections, and 1.3 million linear meters of street. Its street network is about 7.4\% more circuitous than straight-line, as-the-crow-flies edges between nodes would be. The most circuitous network is 14\% more circuitous than straight-line would be, and least is only 2\%. Looking at network complexity in terms of density and connectivity, in the typical urbanized area, the average street segment length (a proxy for block size) is 160 meters. The longest average street segment is the 226-meter average of urbanized Danbury, CT. Puerto Rican cities hold the top four positions for shortest average street segment length, but among the 50 states plus DC, the shortest average street segment is the 125.3-meter average of urbanized Tracy, CA, indicating a much finer street network. The urbanized area of Portland, Oregon, with its famously compact walkable blocks, ranks second at 125.5 meters on average.

\begin{figure}
\includegraphics[width=0.45\textwidth]{media/fig02.png}
\caption{Distribution of node types in 9 urbanized areas, with number of streets emanating from the node on the \emph{x}-axis and proportion of nodes of this type on the \emph{y}-axis }
\label{fig02}
\end{figure}

The typical urbanized area has 26 intersections per km2. Both the densest and the sparsest are in the deep south: the sparsest is 12.5 (Gainesville, GA urbanized area) and the densest is 49.4 (New Orleans urbanized area). However, New Orleans is an anomaly in the deep south. Figure \ref{fig01} depicts the intersection density in each American urbanized area, from lowest density in dark red to highest density in light yellow. The map makes clear how the highest intersection densities are concentrated to the west of the Mississippi River. The lowest intersection densities are concentrated in a belt running from Louisiana, up through the Carolinas and Appalachians, and into New England. In general, only the largest cities on the east coast (e.g., Boston, New York, Philadelphia, Washington) and Florida escape this trend.

The distribution of node types (i.e., intersections and dead ends) provides a clear indicator of network connectedness. The typical urbanized area has 2.8 streets per intersection on average: lots of 3-way intersections, fewer cul-de-sacs, and even fewer 4-way intersections. The grid-like San Angelo, TX urbanized area has the most streets per node (3.2) on average, and (outside of Puerto Rico, which contains the seven lowest urbanized areas) the sprawling, disconnected Lexington Park, MD urbanized area has the fewest (2.2). These two urban areas fit the trend seen in the spatial distribution across the U.S. in Figure \ref{fig01}: urbanized areas in the great plains and Midwest have particularly high numbers of streets per node on average, indicating more grid-like, connected networks. Cities in the southern and western U.S. tend to have fewer streets per node, reflecting more dead-ends and a disconnected network. This finding is discussed in more detail in the upcoming section.

In the typical urbanized area, 18\% of nodes are 4-way intersections, 59\% are 3-way intersections, and 21\% are dead-ends. However, this distribution varies somewhat between urbanized areas. Examining a small sample of 9 urbanized areas, chosen to maximize variance, reveals this in clearer detail. In Figure \ref{fig02}, urban Atlanta and Chattanooga have very high proportions of dead-ends, each over 30\% of all nodes, and very few 4-way intersections, indicating a disconnected street pattern. The urban areas of Phoenix, Boston, Detroit, and Chattanooga have particularly high proportions of 3-way intersections, each over 60\%, indicating a prevalence of T-intersections. Conversely, Chicago, New Orleans, Duluth, and Lubbock have high proportions of 4-way intersections, indicating more grid-like connected networks. But what is perhaps most notable about Figure \ref{fig02} is that these nine urbanized areas, despite being chosen to maximize variance, are overwhelmingly similar to each other. At the metropolitan scale, every large American urban agglomeration is characterized by a preponderance of 3-way intersections.

The relationship between fine-grained networks and connectedness/grid-ness is, however, not clear-cut. Intersection density has only a weak, positive linear relationship with the proportion of 4-way intersections in the urbanized area (r2=0.17). But the relationship between network circuity and grid-ness is somewhat clearer. Average circuity has a negative linear relationship with the proportion of 4-way intersections in the urbanized area (r2=0.43).

\begin{table*}
\caption{Selected street network stats for the 30 largest urbanized areas (by land area).}
\label{table02}
\csvautotabular{media/table02.csv}
\end{table*}

Due to the substantial variation in urbanized area size, from 25 to 9,000 km2, the preceding analysis covers a wide swath of metropolitan place types. To better compare apples-to-apples, Table \ref{table02} focuses on the 30 largest urban areas cross-sectionally to examine how their metric and topological measures compare. This provides more consistent spatial scales and extents, while offering a window into the similarities and differences in the built forms of America's largest urban agglomerations.

Among these urban areas, Milwaukee has the least circuitous network (6\% more circuitous than straight-line edges would be), and Orlando has the most (12\%). San Juan and Atlanta have the fewest streets per node on average (2.36 and 2.45, respectively), while Milwaukee has the most (3.03). Cincinnati has both the lowest intersection density (18/km2) and street density (6.1 km/km2) while Denver has the highest intersection density (40.6/km2) and Miami and Los Angeles have the highest street density (10.6 km/km2, apiece). In other words, Cincinnati has a particularly coarse-grained network with few connections and paths. This can also be seen in the average street segment length, a proxy for block size: Cincinnati has the second highest (186 m), bested only by Cleveland (198 m). In contrast, the two lowest are Denver's 138-meter average and San Juan's 131-meter average.

These metropolitan-scale analyses consider trends in the built form at the scale of broad self-organized human systems and urbanized regions. However, they aggregate multiple heterogenous neighborhoods and municipalities – the scales of human life, urban design projects, and planning jurisdiction – into single units of analysis. To disaggregate and analyze finer characteristics, the following sections examine municipal- and neighborhood-scale street networks across the United States.

\section{Analysis of Municipal-Scale Street Networks}

\begin{figure}[h]
\includegraphics[width=0.45\textwidth]{media/fig03.png}
\caption{The linear relationship between total street length and number of nodes in the street networks of 6,857 US neighborhoods (top) and 19,655 US cities and towns (bottom).}
\label{fig03}
\end{figure}

\begin{figure}[h]
\includegraphics[width=0.45\textwidth]{media/fig04.png}
\caption{Top: contiguous US states by median average street segment length in city and town street networks, from longest (pale yellow) to shortest (dark red). Bottom: contiguous US states by median average number of streets per node in municipal street networks, from lowest (pale yellow) to highest (dark red).}
\label{fig04}
\end{figure}

Table \ref{table03} shows similarly great variation in street network characteristics across the entire data set of 19,655 cities and towns. Again, this data set comprises the street networks of every incorporated city and town in the United States. Following the recent work by Barthélemy and Flammini (2008) and Strano et al. (2013), we examine the relationship between the total street length L and the number of nodes n. Barthélemy and Flammini proposed a model of cities in which L and n scale as n1/2, and Strano et al. claimed to confirm this finding empirically with a small sample of ten European cities' street networks. However, their small sample size limits may the generalizability and interpretability of the finding. To investigate this empirically, we examine the relationship between the total street length L and the number of nodes n for 19,655 U.S. cities and towns and find a strong linear relationship (r2=0.98), contradicting Strano et al., as depicted in Figure \ref{fig03}. A nearly identical linear relationship is found at the metropolitan and neighborhood scales.

\begin{table*}
\caption{Selected summary stats for every incorporated city and town in the US.}
\label{table03}
\csvautotabular{media/table03.csv}
\end{table*}

Moreover, previous findings (e.g., Masucci et al. 2009) suggest the distribution of street segment lengths in an urban street network follows a power-law distribution. However, we find that these networks by and large follow a lognormal distribution instead. This finding makes sense: most street networks are not scale-free. While there are very few very long street segments (say, 1 km), more medium-length segments (say, 250 m), and many short segments (say, 80 m), there are very few very short segments (say, 10 m). This theoretical illustration suggests the lognormal distribution this analysis typically finds across municipal U.S. street networks. One exception, of course, lies in consistently sized, orthogonal grids filling a city's incorporated spatial extents. Such distributions are extremely peaked around a single value: the linear length of a grid block.

This analysis finds that such cities are not uncommon, particularly between the Mississippi River and the Rocky Mountains. These Great Plains states are characterized by a unique street network form that is both orthogonal and reasonably dense. The former is partly the result of topography (flat, plains terrain that allows idealized grids) and design history (platting and development during the late nineteenth century) that favor orthogonal grids, as discussed earlier. The latter seems to result from the fact that most towns across the Great Plains exhibit minimal suburban sprawl. Thus, the municipal boundaries snugly embrace the grid-like street network, without extending to accommodate the vast peripheral belt of twentieth century sprawl, circuity, and loops and lollipops that characterizes cities in California that were settled in the same era but later subjected to substantial urban sprawl.

\begin{table*}
\caption{Median values, aggregated by state plus DC, of selected measures of the municipal-scale street networks for every city and town in the US.}
\label{table04}
\csvautotabular{media/table04.csv}
\end{table*}

For example, if we measure connectedness in terms of average number of streets per node at the city-scale and then aggregate these cities by state (Table \ref{table04}), we find Nebraska, Kansas, South Dakota, Montana, North Dakota, Oklahoma, and Iowa have, in order, the highest medians (Figure \ref{fig04}). This indicates the most grid-like networks. If we measure density and connectedness in terms of intersection density – previously identified as an emergent property of complex network organization – at the city-scale and then aggregate these cities by state, we find Rhode Island, Nebraska, New Jersey, Kansas, and Montana have, in order, the highest medians. We again see three Great Plains states near the top, alongside densely populated East Coast states. Nebraska also has the smallest block sizes (measured via the proxy of average street segment length) while the largest are concentrated in the deep South, upper New England, and Utah (Figure \ref{fig04}).

Municipal boundaries vary greatly in their extents around the built-up area. For example, while Rhode Island averages 56 intersections/km2 in its cities and towns, Alaska averages only 1.3. This is an artifact of Alaska's municipal boundaries often extending thousands of square kilometers beyond the actual built-up area. In fact, Alaska has four cities (Anchorage, Juneau, Sitka, and Wrangell) with such large municipal extents that their land areas exceed that of the state of Rhode Island. These state-level aggregations of municipal-scale street network characteristics show clear variation across the country that reflect topography, economies, culture, planning paradigms, and settlement eras. But they also aggregate and thus obfuscate the variation within each state and within each city. To explore these smaller-scale differences, the following section examines street networks at the neighborhood scale.

\section{Analysis of Neighborhood-Scale Street Networks}

Thus far, we have examined every urban street network in the United States at the metropolitan and municipal scales. This analysis has focused on the complexity on the network in terms of density, resilience, and connectedness. While the metropolitan scale captures the emergent character of the wider region's complex system, and the municipal scale captures planning decisions made by a single city government, the neighborhood scale best represents the scale of individual urban design interventions into the urban form. Further, this scale more commonly reflects individual designs, eras, and paradigms in street network development than the ``many hands, many eras" evolution of form at larger scales.

\begin{table*}
\caption{Selected summary stats for all the neighborhood-scale street networks.}
\label{table05}
\csvautotabular{media/table05.csv}
\end{table*}

Table \ref{table05} presents summary statistics for this data set. Compared to the summary statistics presented at the metropolitan scale (Table \ref{table01}) and the municipal scale (Table \ref{table03}), here we see much greater variance. This is expected, given the smaller network sizes at the neighborhood scale. A few neighborhoods have no intersections within their Zillow-defined boundaries, resulting in a minimum intersection density of 0 across the data set. Meanwhile, the small neighborhood of Cottages North in Davis, California has the highest intersection density in the country, 444/km2, largely as an artifact of its small area as the denominator.

Nationwide, the typical neighborhood averages 2.9 streets per intersection, reflecting the prevalence of 3-way intersections in the U.S., discussed earlier. The median proportions of each node type are 14.5\% for cul-de-sacs, 57.4\% for 3-way intersections, and 23.4\% for 4-way intersections. The typical neighborhood averages 135-meter street segment lengths and 46.4 intersections per km2. At the neighborhood scale (sample size of 6,857 in this analysis) we again find the same strong linear relationship between total street length and the number of nodes in a network (r2=0.98), as seen in Figure \ref{fig03}, contradicting the smaller-sample findings of Strano et al. (2013).

Due to the extreme values seen – resulting from the large variance in neighborhood size – we can filter the data set to examine only large neighborhoods (i.e., neighborhoods with area greater than the median value across the data set). In this filtered set, the five neighborhoods with the highest intersection density are all in central Philadelphia. Central neighborhoods are common at the top of this list, including Point Breeze, Philadelphia; Central Boston; Central City, New Orleans; Downtown Tampa; and Downtown Portland (Figure \ref{fig05}). The three neighborhoods with the lowest intersection density are on the outskirts of Anchorage, Alaska. In this filtered set, the neighborhoods with the greatest average number of streets per node tend to be older neighborhoods with orthogonal grids, such as Virginia Park, Tampa; Outer Sunset, San Francisco; and New Orleans' French Quarter. The neighborhoods with the lowest tend to be sprawling and often hilly suburbs far from the urban core, such as Scholl Canyon in Glendale, California or Sonoma Ranch in San Antonio, Texas.

\begin{figure}
\includegraphics[width=0.45\textwidth]{media/fig05.png}
\caption{Square-mile comparisons of central cities and their suburbs. Left: top, downtown Philadelphia; bottom, suburban King of Prussia. Middle: top, downtown Portland; bottom, suburban Beaverton. Right: top, downtown San Francisco; bottom, suburban Concord.}
\label{fig05}
\end{figure}

For comparison, Figure \ref{fig05} compares one square mile of the centers of Philadelphia, Portland, and San Francisco with one square mile of each of their suburbs. The connectedness and density of the central cities is clear, as is the disconnectivity of their suburbs. In fact, the suburbs have more in common with one another – despite being hundreds or thousands of miles apart – than they do with their central city neighbors, suggesting that land use and an era's prevailing design paradigm is paramount to geographical localism and regional context. The top row of Figure \ref{fig05} represents an era of urban planning and development that preceded the automobile, while the bottom row reflects the exclusionary zoning and mid-late twentieth century era of automobility in residential suburb design – namely the ``loops and lollipops" and the ``lollipops on a stick" design patterns identified by Southworth and Ben-Joseph (1997).

\section{Discussion}

A complex spatial network is a network, embedded in space, that has a nontrivial topology. In other words, its structure and organization is neither fully regular nor fully random. Of the network measures, this empirical study has presented preliminary findings that emphasize street network complexity in terms of density, resilience, and connectedness. Older, denser, and more self-organized networks, such as those at the heart of Boston or lower Manhattan are complex in terms of category I complexity. Sprawling, disconnected suburban neighborhoods rank low on all measures of complexity, with the exception that their high circuity can lend itself to disorder (i.e., category I). The orthogonal grid we see in the downtowns of Portland and San Francisco have high density (i.e., intersection and street densities), connectedness (i.e., average number of streets per node), and order (based on circuity and statistical dispersion of node types), but low resilience in the presence of one-way streets, measured by maximum betweenness centralities and average node connectivity increases when switching from one-way to bi-directional edges. These latter neighborhoods are complex in terms of category II and category III complexity. Their connectedness, density, and orderliness balances at the midpoint embraced by category II. However, their extreme regularity of block sizes, streets per node, and orthogonality best represents category III.

Another critical takeaway of this analysis is that scale matters. The median average circuity is lower across the neighborhoods data set than across the municipal set, which in turn is lower than across the urbanized areas set. Conversely, the median average number of streets per node is higher across the neighborhoods data set than across the municipal set, which in turn is higher than across the urbanized areas set. The median intersection density per km2 is about 83\% higher in the neighborhoods data set than in the municipal or urbanized areas sets. These findings make sense: the Zillow neighborhood boundaries focus on large, core cities with older and denser street networks. The municipal boundaries only include incorporated cities and towns – discarding small census-designated places and unincorporated communities. The urbanized area boundaries include far-flung sprawling suburbs.

The characteristics of an urban network for a city fundamentally depend on what city means: municipal boundaries, urbanized areas, or just certain neighborhoods? The first is a merely legal definition, but also captures the scope of city planning authority and decision-making for top-down interventions into a street network. The second captures a wider self-organized human system and its emergent built form, but tends to aggregate multiple non-homogenous built forms together into a single unit of analysis. The third captures the nature of the local built form and lived experience, but at the expense of a broad view of the urban system and metropolitan-scale trip-taking. In short, multiple scales in concert provide planners and scholars a clearer view of the urban form and the topological and metric complexity of the street network than any single scale can.

We find a strong linear relationship, invariant across scales, between total street length, L, and the number of nodes, n, in a network that contradicts some previous findings in the literature that relied on small sample sizes. We also find that most networks empirically demonstrate a lognormal distribution of street segment lengths, contradicting some previous findings in the literature, as discussed earlier. However, we believe our empirical finding makes more sense theoretically and is supported by the large-sample data at multiple scales. An obvious exception to lognormal distribution lies in those networks that exhibit substantial uniformity across the entire network. At the neighborhood scale, examples include downtown neighborhoods with similarly consistent orthogonal grids, such as that of Portland, Oregon. At the municipal scale, examples include towns in the Great Plains that have orthogonal grids with consistent block sizes, platted at one time, and never subjected to sprawl.

These findings are interesting for the practice and history of planning. The spatial signatures of the Homestead Act, successive land use regulations, urban design paradigms, and planning instruments remain clearly visible today in these cities' urban forms and street networks due to path dependence. When comparing the median municipal street networks of each state, Nebraska has the lowest circuity, the highest average number of streets per node, the second shortest average street segment length, and the second highest intersection density for similar reasons. These preliminary findings point to how street networks across the Great Plains developed all at once, but grew very little afterwards – unlike, for instance, most cities in California that were settled in the same era but later subjected to sprawl.

This finding suggests future research should incorporate a temporal analysis that this present study does not do with its cross-sectional data. This empirical analysis emphasized network structure. Expanding the study of complex urban form by examining the other types of complexity – and further linking structural complexity to the temporal complexity of dynamics and processes – lies ahead as critical future work.

In total, this paper analyzed 497 urbanized areas' street networks, 19,655 cities' and towns' street networks, and 6,857 neighborhoods' street networks. These sample sizes were larger than those in any previous study. It looked at both metric and topological measures of the structure and complexity of these networks – particularly focusing on density, connectedness, and resilience. These preliminary empirical findings demonstrate the use of OSMnx as a new piece of research infrastructure. They suggest to urban planners new methods for acquiring and analyzing street network data, including new methods for evaluating network resilience and resilience gains with betweenness centralities and average node connectivities. Finally, this study has made all of these network datasets – for 497 urbanized areas, 19,655 cities and towns, and 6,857 neighborhoods – along with all of their attribute data and complexity measures available in an online public repository for other researchers to study and repurpose.

\begin{acks}
Redacted
\end{acks}

\begin{thebibliography}{}

\bibliography{references}

\end{thebibliography}

\end{document}

### 2017-03-19 latex/sage/Sage_LaTeX_Guidelines.tex
Initial commit

% sage_latex_guidelines.tex V1.10, 24 June 2016

\documentclass[Afour,sageh,times]{sagej}

\usepackage{moreverb,url}

\usepackage[colorlinks,bookmarksopen,bookmarksnumbered,citecolor=red,urlcolor=red]{hyperref}

\newcommand\BibTeX{{\rmfamily B\kern-.05em \textsc{i\kern-.025em b}\kern-.08em
T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

\def\volumeyear{2016}

\begin{document}

\runninghead{Smith and Wittkopf}

\title{A demonstration of the \LaTeXe\ class file for
\itshape{SAGE Publications}}

\author{Alistair Smith\affilnum{1} and Hendrik Wittkopf\affilnum{2}}

\affiliation{\affilnum{1}Sunrise Setting Ltd, UK\\
\affilnum{2}SAGE Publications Ltd, UK}

\corrauth{Alistair Smith, Sunrise Setting Ltd
Brixham Laboratory,
Freshwater Quarry,
Brixham, Devon,
TQ5~8BA, UK.}

\email{alistair.smith@sunrise-setting.co.uk}

\begin{abstract}
This paper describes the use of the \LaTeXe\
\textsf{\journalclass} class file for setting papers to be
submitted to a \textit{SAGE Publications} journal.
The template can be downloaded \href{http://www.uk.sagepub.com/repository/binaries/SAGE LaTeX template.zip}{here}.
\end{abstract}

\keywords{Class file, \LaTeXe, \textit{SAGE Publications}}

\maketitle

\section{Introduction}
Many authors submitting to research journals use \LaTeXe\ to
prepare their papers. This paper describes the
\textsf{\journalclass} class file which can be used to convert
articles produced with other \LaTeXe\ class files into the correct
form for submission to \textit{SAGE Publications}.

The \textsf{\journalclass} class file preserves much of the
standard \LaTeXe\ interface so that any document which was
produced using the standard \LaTeXe\ \textsf{article} style can
easily be converted to work with the \textsf{\journalclassshort}
style. However, the width of text and typesize will vary from that
of \textsf{article.cls}; therefore, \textit{line breaks will change}
and it is likely that displayed mathematics and tabular material
will need re-setting.

In the following sections we describe how to lay out your code to
use \textsf{\journalclass} to reproduce much of the typographical look of
the \textit{SAGE} journal that you wish to submit to. However, this paper is not a guide to
using \LaTeXe\ and we would refer you to any of the many books
available (see, for example, \cite{R1}, \cite{R2} and \cite{R3}).

\section{The three golden rules}
Before we proceed, we would like to stress \textit{three golden
rules} that need to be followed to enable the most efficient use
of your code at the typesetting stage:
\begin{enumerate}
\item[(i)] keep your own macros to an absolute minimum;

\item[(ii)] as \TeX\ is designed to make sensible spacing
decisions by itself, do \textit{not} use explicit horizontal or
vertical spacing commands, except in a few accepted (mostly
mathematical) situations, such as \verb"\," before a
differential~d, or \verb"\quad" to separate an equation from its
qualifier;

\item[(iii)] follow the journal reference style.
\end{enumerate}

\section{Getting started} The \textsf{\journalclassshort} class file should run
on any standard \LaTeXe\ installation. If any of the fonts, style
files or packages it requires are missing from your installation,
they can be found on the \textit{\TeX\ Collection} DVDs or downloaded from
CTAN.

\begin{figure*}
\setlength{\fboxsep}{0pt}%
\setlength{\fboxrule}{0pt}%
\begin{center}
\begin{boxedverbatim}
\documentclass[<options>]{sagej}

\begin{document}

\runninghead{<Author surnames>}

\title{<Initial capital only>}

\author{<An Author\affilnum{1},
Someone Else\affilnum{2} and
Perhaps Another\affilnum{1}>}

\affiliation{<\affilnum{1}First and third authors' affiliation\\
\affilnum{2}Second author affiliation>}

\corrauth{<Corresponding author's name and full postal address>}

\email{<Corresponding author's email address>}

\begin{abstract}
<Text>
\end{abstract}

\keywords{<List keywords>}

\maketitle

\section{Introduction}
.
.
.
\end{boxedverbatim}
\end{center}
\caption{Example header text.\label{F1}}
\end{figure*}

\section{The article header information}
The heading for any file using \textsf{\journalclass} is shown in
Figure~\ref{F1}. You must select options for the trim/text area and
the reference style of the journal you are submitting to.
The choice of \verb+options+ are listed in Table~\ref{T1}.

\begin{table}[h]
\small\sf\centering
\caption{The choice of options.\label{T1}}
\begin{tabular}{lll}
\toprule
Option&Trim and font size&Columns\\
\midrule
\texttt{shortAfour}& 210 $\times$ 280 mm, 10pt& Double column\\
\texttt{Afour} &210 $\times$ 297 mm, 10pt& Double column\\
\texttt{MCfour} &189 $\times$ 246 mm, 10pt& Double column\\
\texttt{PCfour} &170 $\times$ 242 mm, 10pt& Double column\\
\texttt{Royal} &156 $\times$ 234 mm, 10pt& Single column\\
\texttt{Crown} &7.25 $\times$ 9.5 in, 10pt&Single column\\
\texttt{Review} & 156 $\times$ 234 mm, 12pt & Single column\\
\bottomrule
\end{tabular}\\[10pt]
\begin{tabular}{ll}
\toprule
Option&Reference style\\
\midrule
\texttt{sageh}&SAGE Harvard style (author-year)\\
\texttt{sagev}&SAGE Vancouver style (superscript numbers)\\
\texttt{sageapa}&APA style (author-year)\\
\bottomrule
\end{tabular}
\end{table}

For example, if your journal is short A4 sized, uses Times fonts and has Harvard style references then you would need\\
{\small\verb+\documentclass[ShortAfour,times,sageh]{sagej}+}

Most \textit{SAGE} journals are published using Times fonts but if for any reason you have a problem using Times you can
easily resort to Computer Modern fonts by removing the
\verb"times" option.

\subsection{`Review' option}
Some journals (for example, \emph{Journal of the Society for Clinical Trials}) require that 
papers are set single column and with a larger font size to help with the review process. 
If this is a requirement for the journal that you are submitting to, just add the \verb+Review+ option to the \verb+\documenclass[]{sagej}+ line.

\subsection{Remarks}
\begin{enumerate}
\item[(i)] In \verb"\runninghead" use `\textit{et~al.}' if there
are three or more authors.

\item[(ii)] For multiple author papers please note the use of \verb"\affilnum" to
link names and affiliations. The corresponding author details need to be included using the
\verb+\corrauth+ and \verb+\email+ commands.

\item[(iii)] For submitting a double-spaced manuscript, add
\verb"doublespace" as an option to the documentclass line.

\item[(iv)] The abstract should be capable of standing by itself,
in the absence of the body of the article and of the bibliography.
Therefore, it must not contain any reference citations.

\item[(v)] Keywords are separated by commas.

\item[(vi)] If you are submitting to a \textit{SAGE} journal that requires numbered sections (for example, IJRR), please add the command
  \verb+\setcounter{secnumdepth}{3}+ just above the \verb+\begin{document}+ line.

\end{enumerate}

\section{The body of the article}

\subsection{Mathematics} \textsf{\journalclass} makes the full
functionality of \AmS\/\TeX\ available. We encourage the use of
the \verb"align", \verb"gather" and \verb"multline" environments
for displayed mathematics. \textsf{amsthm} is used for setting
theorem-like and proof environments. The usual \verb"\newtheorem"
command needs to be used to set up the environments for your
particular document.

\subsection{Figures and tables} \textsf{\journalclass} includes the
\textsf{graphicx} package for handling figures.

Figures are called in as follows:
\begin{verbatim}
\begin{figure}
\centering
\includegraphics{<figure name>}
\caption{<Figure caption>}
\end{figure}
\end{verbatim}

For further details on how to size figures, etc., with the
\textsf{graphicx} package see, for example, \cite{R1}
or \cite{R3}.

The standard coding for a table is shown in Figure~\ref{F2}.

\begin{figure}
\setlength{\fboxsep}{0pt}%
\setlength{\fboxrule}{0pt}%
\begin{center}
\begin{boxedverbatim}
\begin{table}
\small\sf\centering
\caption{<Table caption.>}
\begin{tabular}{<table alignment>}
\toprule
<column headings>\\
\midrule
<table entries
(separated by & as usual)>\\
<table entries>\\
.
.
.\\
\bottomrule
\end{tabular}
\end{table}
\end{boxedverbatim}
\end{center}
\caption{Example table layout.\label{F2}}
\end{figure}

\subsection{Cross-referencing}
The use of the \LaTeX\ cross-reference system
for figures, tables, equations, etc., is encouraged
(using \verb"\ref{<name>}" and \verb"\label{<name>}").

\subsection{End of paper special sections}
Depending on the requirements of the journal that you are submitting to,
there are macros defined to typeset various special sections.

The commands available are:
\begin{verbatim}
\begin{acks}
To typeset an
  "Acknowledgements" section.
\end{acks}
\end{verbatim}

\begin{verbatim}
\begin{biog}
To typeset an
  "Author biography" section.
\end{biog}
\end{verbatim}

\begin{verbatim}
\begin{biogs}
To typeset an
  "Author Biographies" section.
\end{biogs}
\end{verbatim}

%\newpage

\begin{verbatim}
\begin{dci}
To typeset a "Declaration of
  conflicting interests" section.
\end{dci}
\end{verbatim}

\begin{verbatim}
\begin{funding}
To typeset a "Funding" section.
\end{funding}
\end{verbatim}

\begin{verbatim}
\begin{sm}
To typeset a
  "Supplemental material" section.
\end{sm}
\end{verbatim}

\subsection{Endnotes}
Most \textit{SAGE} journals use endnotes rather than footnotes, so any notes should be coded as \verb+\endnote{<Text>}+.
Place the command \verb+\theendnotes+ just above the Reference section to typeset the endnotes.

To avoid any confusion for papers that use Vancouver style references,  footnotes/endnotes should be edited into the text.

\subsection{References}
Please note that the files \textsf{SageH.bst} and \textsf{SageV.bst} are included with the class file
for those authors using \BibTeX. 
The files work in a completely standard way, and you just need to uncomment one of the lines in the below example depending on what style you require:
\begin{verbatim}
%%Harvard (name/date)
%\bibliographystyle{SageH}
%%Vancouver (numbered)
%\bibliographystyle{SageV} 
\bibliography{<YourBibfile.bib>} 
\end{verbatim}

%\section{Support for \textsf{\journalclass}}
%We offer on-line support to participating authors. Please contact
%us via e-mail at \dots
%
%We would welcome any feedback, positive or otherwise, on your
%experiences of using \textsf{\journalclass}.

\section{Copyright statement}
Please  be  aware that the use of  this \LaTeXe\ class file is
governed by the following conditions.

\subsection{Copyright}
Copyright \copyright\ \volumeyear\ SAGE Publications Ltd,
1 Oliver's Yard, 55 City Road, London, EC1Y~1SP, UK. All
rights reserved.

\subsection{Rules of use}
This class file is made available for use by authors who wish to
prepare an article for publication in a \textit{SAGE Publications} journal.
The user may not exploit any
part of the class file commercially.

This class file is provided on an \textit{as is}  basis, without
warranties of any kind, either express or implied, including but
not limited to warranties of title, or implied  warranties of
merchantablility or fitness for a particular purpose. There will
be no duty on the author[s] of the software or SAGE Publications Ltd
to correct any errors or defects in the software. Any
statutory  rights you may have remain unaffected by your
acceptance of these rules of use.

\begin{acks}
This class file was developed by Sunrise Setting Ltd,
Brixham, Devon, UK.\\
Website: \url{http://www.sunrise-setting.co.uk}
\end{acks}

\begin{thebibliography}{99}
\bibitem[Kopka and Daly(2003)]{R1}
Kopka~H and Daly~PW (2003) \textit{A Guide to \LaTeX}, 4th~edn.
Addison-Wesley.

\bibitem[Lamport(1994)]{R2}
Lamport~L (1994) \textit{\LaTeX: a Document Preparation System},
2nd~edn. Addison-Wesley.

\bibitem[Mittelbach and Goossens(2004)]{R3}
Mittelbach~F and Goossens~M (2004) \textit{The \LaTeX\ Companion},
2nd~edn. Addison-Wesley.

\end{thebibliography}

\end{document}

### 2017-03-19 latex/paper.tex
clean up references and figs

@@
 
-\bibliographystyle{sage/SageV}
+\bibliographystyle{plainnat}
+\setcitestyle{numbers}
 
@@
 \def\volumeyear{2017}
-
 \begin{document}
@@
 \author{Geoff Boeing\affilnum{1}}
-\affiliation{\affilnum{1}Department of City and Regional Planning, University of California, Berkeley}
+\affiliation{\affilnum{1}PhD Candidate, Department of City and Regional Planning, University of California, Berkeley}
 \corrauth{Geoff Boeing,
@@
 \begin{abstract}
-This paper serves illustrative purposes as it presents wide-ranging preliminary empirical findings on U.S.
-urban form and street network characteristics, using OSMnx.
+This paper serves illustrative purposes as it presents wide-ranging preliminary empirical findings on US urban form and street network characteristics, using OSMnx.
 It also demonstrates the scalability and flexibility of OSMnx as a new research tool.
@@
 On May 20, 1862, Abraham Lincoln signed the Homestead Act into law, making land across the United States Midwest and Great Plains available for free to all applicants \cite{porterfield_homestead_2005}.
-Under its auspices over the next 70 years, the federal government distributed 270 million acres of public land (10\% of the entire U.S.
-landmass) to private owners in the form of 1.6 million homesteads \cite{lee_kansas_1979, sherraden_inclusion_2005}.
-Towns with gridiron street networks sprang up rapidly across the Great Plains and Midwest, due to both the prevailing urban design paradigm of the day and the standardized rectilinear town plats used repeatedly to lay out instant new cities (Southworth and Ben-Joseph 1997).
+Under its auspices over the next 70 years, the federal government distributed 270 million acres of public land (10\% of the entire US landmass) to private owners in the form of 1.6 million homesteads \cite{lee_kansas_1979, sherraden_inclusion_2005}.
+Towns with gridiron street networks sprang up rapidly across the Great Plains and Midwest, due to both the prevailing urban design paradigm of the day and the standardized rectilinear town plats used repeatedly to lay out instant new cities \cite{southworth_streets_1997}.
 Through path dependence, the spatial signatures of these land use laws, design paradigms, and planning instruments can still be seen today in these cities' urban forms and street networks.
@@
 A cross-sectional analysis of American urban form through its street networks at metropolitan, municipal, and neighborhood scales can reveal these artifacts and histories across the nation.
-Network analysis is a natural approach to the study of cities as growing, complex systems that self-organize to fill space with a capillary circulation network (Masucci et al.
-2009).
-The empirical literature on street network analysis is growing ever richer, but suffers from some limitations.
+Network analysis is a natural approach to the study of cities as growing, complex systems that self-organize to fill space with a capillary circulation network \cite{masucci_random_2009}.
+The empirical literature on street network analysis is growing ever richer, but suffers from some limitations---discussed in \cite{boeing_osmnx:_2017} and summarized here.
 First, sample sizes tend to be fairly small due to data availability, gathering, and processing constraints.
-Second, reproducibility is difficult when the dozens of decisions that go into analysis – such as spatial extents, topological simplification and correction, definitions of nodes and edges, etc.
-– are ad hoc or only partly reported.
+Second, reproducibility is difficult when the dozens of decisions that go into analysis---such as spatial extents, topological simplification and correction, definitions of nodes and edges, etc.---are ad hoc or only partly reported.
 Third, and related to the first two, studies frequently oversimplify to planar or undirected primal graphs for tractability, or use dual graphs despite the loss of geographic and metric information.
+
 This paper addresses these limitations by conducting an analysis of street networks at multiple scales, with large sample sizes, with clearly defined network definitions and extents for reproducibility, and using non-planar, directed graphs.
-In particular, it examines urban street networks – represented here as primal, non-planar, weighted multidigraphs with possible self-loops – focusing on structure, connectedness, centrality, and resilience.
-
+In particular, it examines urban street networks---represented here as primal, non-planar, weighted multidigraphs with possible self-loops---focusing on structure, connectedness, centrality, and resilience.
 Most studies in the street network literature that conduct topological and/or metric analysis tend to have sample sizes ranging around 5 to 50 networks.
 This paper instead conducts a large analysis of over 27,000 urban street networks at multiple overlapping scales across the United States.
-Namely, it examines the street networks of every U.S.
-incorporated city and town, urbanized area, and Zillow-defined neighborhood.
-In total, this study presented uses OSMnx to download, construct, and analyze 497 urbanized areas' street networks, 19,655 cities' and towns' street networks, and 6,857 neighborhoods' street networks.
+Namely, it examines the street networks of every US incorporated city and town, urbanized area, and Zillow-defined neighborhood.
+In total, this study presented uses OSMnx \cite{boeing_osmnx:_2017} to download, construct, and analyze 497 urbanized areas' street networks, 19,655 cities' and towns' street networks, and 6,857 neighborhoods' street networks.
 It uses these street networks to conduct four analyses: at the metropolitan scale, at the municipal scale, at the neighborhood scale, and with a case study looking deeper at the neighborhood-scale street networks in the city of San Francisco.
@@
 Then it presents findings of the analyses at three spatial scales: metropolitan, municipal, and neighborhood.
-Next it presents a case study of the neighborhood scale in San Francisco.
 Finally, it concludes with a discussion of these findings and their implications for street network analysis, urban form complexity, and city planning.
@@
 \section{Methodology}
-This study uses OSMnx to download, construct, correct, analyze, and visualize street networks at metropolitan, municipal, and neighborhood scales.
+This study uses OSMnx to download, construct, correct, analyze, and visualize street networks at metropolitan, municipal, and neighborhood scales \cite{boeing_osmnx:_2017}.
 To define the study sites and their spatial boundaries, this study uses three sets of geometries.
-The first defines the metropolitan-scale study sites using the 2016 national TIGER/Line shapefile of U.S.
-census bureau urban areas, released 19 August 2016.
-Census-defined urban areas comprise a set of census tracts that meet a minimum density threshold (U.S.
-Census Bureau 2010).
+The first defines the metropolitan-scale study sites using the 2016 national TIGER/Line shapefile of US census bureau urban areas, released 19 August 2016.
+Census-defined urban areas comprise a set of census tracts that meet a minimum density threshold \cite{u.s._census_bureau_2010_2010}.
 In particular, we retain only the urbanized areas subset in this data set (i.e., areas with greater than 50,000 population), discarding the smaller urban clusters subset.
-The second set of geometries defines the municipal-scale study sites using 51 TIGER/Line shapefiles (again, 2016) of U.S.
-census bureau places for all 50 states plus the District of Columbia.
+The second set of geometries defines the municipal-scale study sites using 51 TIGER/Line shapefiles (again, 2016) of US census bureau places for all 50 states plus the District of Columbia.
 In particular, we discard the subset of census-designated places in this data set, instead retaining all the incorporated cities and towns in the United States.
@@
 These shapefiles contain the geometries of neighborhoods in 41 states plus the District of Columbia.
-This is a fairly new data set comprising nearly 7,000 neighborhoods in large American cities, but as Schernthanner et al.
-(2016) point out, Zillow does not publish the methodology it uses to construct these geometries.
+This is a fairly new data set comprising nearly 7,000 neighborhoods in large American cities, but as \citet{schernthanner_spatial_2016} point out, Zillow does not publish the methodology it uses to construct these geometries.
 However, despite its newness it already has a track record in the academic literature.
-For instance, Besbris et al.
-(2015) used Zillow neighborhood boundaries to examine neighborhood stigma, and Albrecht and Abramowitz (2014) used these data to study neighborhood-level poverty in New York.
+For instance, \citet{besbris_effect_2015} used Zillow boundaries to examine neighborhood stigma, and \citet{albrecht_indicator_2014} used these data to study neighborhood-level poverty in New York.
 
@@
 Finally, it automatically calculates metric and topological measures for each.
-However, in a planar graph, topological and metric properties are somewhat interrelated (Masucci et al.
-2009).
+However, in a planar graph, topological and metric properties are somewhat interrelated \cite{masucci_random_2009}.
 
@@
 Connectivity measures the minimum number of nodes or edges that must be removed from a connected graph to disconnect it.
-The average node connectivity of a network – the mean number of internally node-disjoint paths between each pair of nodes in the graph – represents the expected number of nodes that must be removed to disconnect a randomly selected pair of non-adjacent nodes (Beineke et al.
-2002).
+The average node connectivity of a network---the mean number of internally node-disjoint paths between each pair of nodes in the graph---represents the expected number of nodes that must be removed to disconnect a randomly selected pair of non-adjacent nodes \cite{beineke_average_2002}.
 
@@
 Centrality indicates the importance of nodes in a network.
-Betweenness centrality evaluates the number of shortest paths that pass through each node (Barthélemy 2004).
+Betweenness centrality evaluates the number of shortest paths that pass through each node \cite{barthelemy_betweenness_2004}.
 The maximum betweenness centrality in a network specifies the proportion of shortest paths that pass through the most important node.
@@
 
-\begin{table*}
-\caption{Measures of central tendency and dispersion for selected measures of the 497 urbanized area street networks.}
+\begin{table*}[h]
+\caption{Central tendency and statistical dispersion for selected measures of all US urbanized areas' street networks.}
 \label{table01}
@@
 Table \ref{table01} shows the considerable variation in street network characteristics across the entire data set of 497 urbanized areas.
-This is unsurprising: urbanized areas span a wide spectrum of sizes, from the Delano, CA Urbanized Area's 26 km2 to the New York--Newark, NY--NJ--CT Urbanized Area's 8,937 km2 – thus, density and count-based measures demonstrate substantial variance.
+This is unsurprising: urbanized areas span a wide spectrum of sizes, from the Delano, CA Urbanized Area's 26 km\textsuperscript{2} to the New York--Newark, NY--NJ--CT Urbanized Area's 8,937 km\textsuperscript{2}---thus, density and count-based measures demonstrate substantial variance.
 Further, these urbanized areas span a wide spectrum of terrains, development eras and paradigms, and cultures.
 
-\begin{figure}
-\includegraphics[width=0.45\textwidth]{media/fig01.png}
-\caption{Intersection density per urbanized area (top) and average streets per node per urbanized area (bottom) from lowest (pale yellow) to highest (dark red), in the contiguous U.S.}
+\begin{figure*}[h]
+\includegraphics[width=1\textwidth]{media/fig01.png}
+\caption{Intersection density per urbanized area (left) and average streets per node per urbanized area (right) from lowest (pale yellow) to highest (dark red), in the contiguous US}
 \label{fig01}
-\end{figure}
+\end{figure*}
 
 Nevertheless, looking across the data set provides a sense of the breadth of American metropolitan street networks.
-New York's urbanized area – America's largest – has 373,309 intersections and 79 million meters of linear street (or 417,570 and 83.4 million if you include service roads).
-Delano, CA's urbanized area – America's smallest – has 874 intersections and 222,328 meters of linear street (or 964 and 231,000 meters if you include service roads).
-The typical (Table \ref{table01}, \emph{median}) American urbanized area is approximately 185 km2 in land area, has 5,830 intersections, and 1.3 million linear meters of street.
+New York's urbanized area---America's largest-–-has 373,309 intersections and 79 million meters of linear street (or 417,570 and 83.4 million if including service roads).
+Delano, CA's urbanized area---America's smallest---has 874 intersections and 222,328 meters of linear street (or 964 and 231,000 meters if including service roads).
+The typical (Table \ref{table01}, \emph{median}) American urbanized area is approximately 185 km\textsuperscript{2} in land area, has 5,830 intersections, and 1.3 million linear meters of street.
 Its street network is about 7.4\% more circuitous than straight-line, as-the-crow-flies edges between nodes would be.
@@
 
-\begin{figure}
+\begin{figure}[h]
 \includegraphics[width=0.45\textwidth]{media/fig02.png}
@@
 
-The typical urbanized area has 26 intersections per km2.
+The typical urbanized area has 26 intersections per km\textsuperscript{2}.
 Both the densest and the sparsest are in the deep south: the sparsest is 12.5 (Gainesville, GA urbanized area) and the densest is 49.4 (New Orleans urbanized area).
@@
 The grid-like San Angelo, TX urbanized area has the most streets per node (3.2) on average, and (outside of Puerto Rico, which contains the seven lowest urbanized areas) the sprawling, disconnected Lexington Park, MD urbanized area has the fewest (2.2).
-These two urban areas fit the trend seen in the spatial distribution across the U.S.
-in Figure \ref{fig01}: urbanized areas in the great plains and Midwest have particularly high numbers of streets per node on average, indicating more grid-like, connected networks.
-Cities in the southern and western U.S.
-tend to have fewer streets per node, reflecting more dead-ends and a disconnected network.
+These two urban areas fit the trend seen in the spatial distribution across the US in Figure \ref{fig01}: urbanized areas in the great plains and Midwest have particularly high numbers of streets per node on average, indicating more grid-like, connected networks.
+Cities in the southern and western US tend to have fewer streets per node, reflecting more dead-ends and a disconnected network.
 This finding is discussed in more detail in the upcoming section.
@@
 The relationship between fine-grained networks and connectedness/grid-ness is, however, not clear-cut.
-Intersection density has only a weak, positive linear relationship with the proportion of 4-way intersections in the urbanized area (r2=0.17).
+Intersection density has only a weak, positive linear relationship with the proportion of 4-way intersections in the urbanized area ($r^{2}=0.17$).
 But the relationship between network circuity and grid-ness is somewhat clearer.
-Average circuity has a negative linear relationship with the proportion of 4-way intersections in the urbanized area (r2=0.43).
+Average circuity has a negative linear relationship with the proportion of 4-way intersections in the urbanized area ($r^{2}=0.43$).
 
-\begin{table*}
-\caption{Selected street network stats for the 30 largest urbanized areas (by land area).}
+\begin{table*}[h]
+\caption{Selected measures of the 30 largest (by land area) urbanized areas' street networks.}
 \label{table02}
@@
 
-Due to the substantial variation in urbanized area size, from 25 to 9,000 km2, the preceding analysis covers a wide swath of metropolitan place types.
+Due to the substantial variation in urbanized area size, from 25 to 9,000 km\textsuperscript{2}, the preceding analysis covers a wide swath of metropolitan place types.
 To better compare apples-to-apples, Table \ref{table02} focuses on the 30 largest urban areas cross-sectionally to examine how their metric and topological measures compare.
@@
 San Juan and Atlanta have the fewest streets per node on average (2.36 and 2.45, respectively), while Milwaukee has the most (3.03).
-Cincinnati has both the lowest intersection density (18/km2) and street density (6.1 km/km2) while Denver has the highest intersection density (40.6/km2) and Miami and Los Angeles have the highest street density (10.6 km/km2, apiece).
+Cincinnati has both the lowest intersection density (18/km\textsuperscript{2}) and street density (6.1 km/km\textsuperscript{2}) while Denver has the highest intersection density (40.6/km\textsuperscript{2}) and Miami and Los Angeles have the highest street density (10.6 km/km\textsuperscript{2}, apiece).
 In other words, Cincinnati has a particularly coarse-grained network with few connections and paths.
@@
 These metropolitan-scale analyses consider trends in the built form at the scale of broad self-organized human systems and urbanized regions.
-However, they aggregate multiple heterogenous neighborhoods and municipalities – the scales of human life, urban design projects, and planning jurisdiction – into single units of analysis.
+However, they aggregate multiple heterogenous neighborhoods and municipalities---the scales of human life, urban design projects, and planning jurisdiction---into single units of analysis.
 To disaggregate and analyze finer characteristics, the following sections examine municipal- and neighborhood-scale street networks across the United States.
@@
 
-\begin{figure}[h]
-\includegraphics[width=0.45\textwidth]{media/fig03.png}
-\caption{The linear relationship between total street length and number of nodes in the street networks of 6,857 US neighborhoods (top) and 19,655 US cities and towns (bottom).}
-\label{fig03}
-\end{figure}
-
-\begin{figure}[h]
-\includegraphics[width=0.45\textwidth]{media/fig04.png}
-\caption{Top: contiguous US states by median average street segment length in city and town street networks, from longest (pale yellow) to shortest (dark red).
-Bottom: contiguous US states by median average number of streets per node in municipal street networks, from lowest (pale yellow) to highest (dark red).}
-\label{fig04}
-\end{figure}
-
 Table \ref{table03} shows similarly great variation in street network characteristics across the entire data set of 19,655 cities and towns.
 Again, this data set comprises the street networks of every incorporated city and town in the United States.
-Following the recent work by Barthélemy and Flammini (2008) and Strano et al.
-(2013), we examine the relationship between the total street length L and the number of nodes n.
-Barthélemy and Flammini proposed a model of cities in which L and n scale as n1/2, and Strano et al.
-claimed to confirm this finding empirically with a small sample of ten European cities' street networks.
-However, their small sample size limits may the generalizability and interpretability of the finding.
-To investigate this empirically, we examine the relationship between the total street length L and the number of nodes n for 19,655 U.S.
-cities and towns and find a strong linear relationship (r2=0.98), contradicting Strano et al., as depicted in Figure \ref{fig03}.
-A nearly identical linear relationship is found at the metropolitan and neighborhood scales.
+Following recent work by \citet{barthelemy_modeling_2008} and \citet{strano_urban_2013}, we examine the relationship between the total street length $L$ and the number of nodes $n$.
+The former proposed a model of cities in which $L$ and $n$ scale nonlinearly as $n^{1/2}$, and the latter recently confirmed this empirically with a small sample of ten European cities' street networks.
+However, their small sample size may limit the generalizability and interpretability of their finding.
+To investigate this empirically, we examine the relationship between the total street length $L$ and the number of nodes $n$ for 19,655 US cities and towns and find a strong \emph{linear} relationship ($r^{2}=0.98$), as depicted in Figure \ref{fig03}, contradicting Strano et al.
+We also find a similar linear relationship at the metropolitan and neighborhood scales.
 
-\begin{table*}
+\begin{table*}[h]
 \caption{Selected summary stats for every incorporated city and town in the US.}
@@
 
-Moreover, previous findings (e.g., Masucci et al.
-2009) suggest the distribution of street segment lengths in an urban street network follows a power-law distribution.
-However, we find that these networks by and large follow a lognormal distribution instead.
-This finding makes sense: most street networks are not scale-free.
-While there are very few very long street segments (say, 1 km), more medium-length segments (say, 250 m), and many short segments (say, 80 m), there are very few very short segments (say, 10 m).
-This theoretical illustration suggests the lognormal distribution this analysis typically finds across municipal U.S.
-street networks.
-One exception, of course, lies in consistently sized, orthogonal grids filling a city's incorporated spatial extents.
+\begin{figure}[h]
+	\includegraphics[width=0.45\textwidth]{media/fig03-v.png}
+	\caption{The linear relationship between total street length and number of nodes in the street networks of 6,857 US neighborhoods (top) and 19,655 US cities and towns (bottom).}
+	\label{fig03}
+\end{figure}
+
+\begin{figure*}[h]
+	\includegraphics[width=1\textwidth]{media/fig04.png}
+	\caption{Left: contiguous US states by median of mean streets per node in municipal street networks, from lowest (pale yellow) to highest (dark red).
+Right: contiguous US states by median of mean street segment length in municipal street networks, from longest (pale yellow) to shortest (dark red).}
+	\label{fig04}
+\end{figure*}
+
+Moreover, previous findings \citep[e.g.,][]{masucci_random_2009} suggest the distribution of street segment lengths in an urban street network follows a power-law distribution.
+However, we find that these networks instead generally follow a lognormal distribution, which makes theoretical sense as most street networks are not truly scale-free.
+For example, a typical street network might comprise very few very long street segments (e.g., 1 km), more medium-length segments (e.g., 250 m), and many short segments (e.g., 80 m), but very few very short segments (e.g., 10 m).
+This theoretical illustration suggests the lognormal distribution that this study generally finds across municipal US street networks.
+An exception, of course, lies in consistently-sized orthogonal grids filling a city's incorporated spatial extents.
 Such distributions are extremely peaked around a single value: the linear length of a grid block.
 
-This analysis finds that such cities are not uncommon, particularly between the Mississippi River and the Rocky Mountains.
-These Great Plains states are characterized by a unique street network form that is both orthogonal and reasonably dense.
+This study finds that such cities are not uncommon, particularly between the Mississippi River and the Rocky Mountains: the Great Plains states are characterized by a unique street network form that is both orthogonal and reasonably dense.
 The former is partly the result of topography (flat, plains terrain that allows idealized grids) and design history (platting and development during the late nineteenth century) that favor orthogonal grids, as discussed earlier.
-The latter seems to result from the fact that most towns across the Great Plains exhibit minimal suburban sprawl.
-Thus, the municipal boundaries snugly embrace the grid-like street network, without extending to accommodate the vast peripheral belt of twentieth century sprawl, circuity, and loops and lollipops that characterizes cities in California that were settled in the same era but later subjected to substantial urban sprawl.
+The latter results from the fact that most towns across the Great Plains exhibit minimal suburban sprawl.
+Thus, the municipal boundaries snugly embrace the grid-like street network, without extending to accommodate the vast peripheral belt of twentieth century sprawl, circuity, and ``loops and lollipops" \cite{southworth_streets_1997} that characterizes cities in California that were settled in the same era but later subjected to substantial suburbanization.
 
-\begin{table*}
+\begin{table*}[h]
 \caption{Median values, aggregated by state plus DC, of selected measures of the municipal-scale street networks for every city and town in the US.}
@@
 This indicates the most grid-like networks.
-If we measure density and connectedness in terms of intersection density – previously identified as an emergent property of complex network organization – at the city-scale and then aggregate these cities by state, we find Rhode Island, Nebraska, New Jersey, Kansas, and Montana have, in order, the highest medians.
+If we measure density and connectedness in terms of intersection density---previously identified as an emergent property of complex network organization---at the city-scale and then aggregate these cities by state, we find Rhode Island, Nebraska, New Jersey, Kansas, and Montana have, in order, the highest medians.
 We again see three Great Plains states near the top, alongside densely populated East Coast states.
@@
 Municipal boundaries vary greatly in their extents around the built-up area.
-For example, while Rhode Island averages 56 intersections/km2 in its cities and towns, Alaska averages only 1.3.
+For example, while Rhode Island averages 56 intersections/km\textsuperscript{2} in its cities and towns, Alaska averages only 1.3.
 This is an artifact of Alaska's municipal boundaries often extending thousands of square kilometers beyond the actual built-up area.
@@
 
-\begin{table*}
+\begin{table*}[h]
 \caption{Selected summary stats for all the neighborhood-scale street networks.}
@@
 A few neighborhoods have no intersections within their Zillow-defined boundaries, resulting in a minimum intersection density of 0 across the data set.
-Meanwhile, the small neighborhood of Cottages North in Davis, California has the highest intersection density in the country, 444/km2, largely as an artifact of its small area as the denominator.
+Meanwhile, the small neighborhood of Cottages North in Davis, California has the highest intersection density in the country, 444/km\textsuperscript{2}, largely as an artifact of its small area as the denominator.
 
-Nationwide, the typical neighborhood averages 2.9 streets per intersection, reflecting the prevalence of 3-way intersections in the U.S., discussed earlier.
+Nationwide, the typical neighborhood averages 2.9 streets per intersection, reflecting the prevalence of 3-way intersections in the US, discussed earlier.
 The median proportions of each node type are 14.5\% for cul-de-sacs, 57.4\% for 3-way intersections, and 23.4\% for 4-way intersections.
-The typical neighborhood averages 135-meter street segment lengths and 46.4 intersections per km2.
-At the neighborhood scale (sample size of 6,857 in this analysis) we again find the same strong linear relationship between total street length and the number of nodes in a network (r2=0.98), as seen in Figure \ref{fig03}, contradicting the smaller-sample findings of Strano et al.
-(2013).
+The typical neighborhood averages 135-meter street segment lengths and 46.4 intersections per km\textsuperscript{2}.
+At the neighborhood scale (sample size of 6,857 in this analysis) we again find the same strong \emph{linear} relationship between total street length and the number of nodes in a network ($r^{2}=0.98$), as seen in Figure \ref{fig03}, contradicting the smaller-sample findings of \citet{strano_urban_2013}.
 
-Due to the extreme values seen – resulting from the large variance in neighborhood size – we can filter the data set to examine only large neighborhoods (i.e., neighborhoods with area greater than the median value across the data set).
+Due to the extreme values seen---resulting from the large variance in neighborhood size---we can filter the data set to examine only large neighborhoods (i.e., with area greater than the median value across the data set).
 In this filtered set, the five neighborhoods with the highest intersection density are all in central Philadelphia.
@@
 
-\begin{figure}
-\includegraphics[width=0.45\textwidth]{media/fig05.png}
+\begin{figure*}[h]
+	\includegraphics[width=1\textwidth]{media/fig05.png}
 	\caption{Square-mile comparisons of central cities and their suburbs.
-Left: top, downtown Philadelphia; bottom, suburban King of Prussia.
-Middle: top, downtown Portland; bottom, suburban Beaverton.
-Right: top, downtown San Francisco; bottom, suburban Concord.}
+Left: top, downtown Philadelphia; bottom, its suburb, King of Prussia.
+Middle: top, downtown Portland; bottom, its suburb, Beaverton.
+Right: top, downtown San Francisco; bottom, its suburb Concord.}
 	\label{fig05}
-\end{figure}
+\end{figure*}
 
@@
 The connectedness and density of the central cities is clear, as is the disconnectivity of their suburbs.
-In fact, the suburbs have more in common with one another – despite being hundreds or thousands of miles apart – than they do with their central city neighbors, suggesting that land use and an era's prevailing design paradigm is paramount to geographical localism and regional context.
-The top row of Figure \ref{fig05} represents an era of urban planning and development that preceded the automobile, while the bottom row reflects the exclusionary zoning and mid-late twentieth century era of automobility in residential suburb design – namely the ``loops and lollipops" and the ``lollipops on a stick" design patterns identified by Southworth and Ben-Joseph (1997).
+In fact, the suburbs have more in common with one another---despite being hundreds or thousands of miles apart---than they do with their central city neighbors, suggesting that land use and an era's prevailing design paradigm is paramount to geographical localism and regional context.
+The top row of Figure \ref{fig05} represents an era of urban planning and development that preceded the automobile, while the bottom row reflects the exclusionary zoning and mid-late twentieth century era of automobility in residential suburb design---namely the ``loops and lollipops" and the ``lollipops on a stick" design patterns identified by \citet{southworth_streets_1997}.
 
@@
 Conversely, the median average number of streets per node is higher across the neighborhoods data set than across the municipal set, which in turn is higher than across the urbanized areas set.
-The median intersection density per km2 is about 83\% higher in the neighborhoods data set than in the municipal or urbanized areas sets.
+The median intersection density per km\textsuperscript{2} is about 83\% higher in the neighborhoods data set than in the municipal or urbanized areas sets.
 These findings make sense: the Zillow neighborhood boundaries focus on large, core cities with older and denser street networks.
-The municipal boundaries only include incorporated cities and towns – discarding small census-designated places and unincorporated communities.
+The municipal boundaries only include incorporated cities and towns---discarding small census-designated places and unincorporated communities.
 The urbanized area boundaries include far-flung sprawling suburbs.
@@
 When comparing the median municipal street networks of each state, Nebraska has the lowest circuity, the highest average number of streets per node, the second shortest average street segment length, and the second highest intersection density for similar reasons.
-These preliminary findings point to how street networks across the Great Plains developed all at once, but grew very little afterwards – unlike, for instance, most cities in California that were settled in the same era but later subjected to sprawl.
+These preliminary findings point to how street networks across the Great Plains developed all at once, but grew very little afterwards---unlike, for instance, most cities in California that were settled in the same era but later subjected to sprawl.
 
@@
 This empirical analysis emphasized network structure.
-Expanding the study of complex urban form by examining the other types of complexity – and further linking structural complexity to the temporal complexity of dynamics and processes – lies ahead as critical future work.
+Expanding the study of complex urban form by examining the other types of complexity---and further linking structural complexity to the temporal complexity of dynamics and processes---lies ahead as critical future work.
 
@@
 These sample sizes were larger than those in any previous study.
-It looked at both metric and topological measures of the structure and complexity of these networks – particularly focusing on density, connectedness, and resilience.
+It looked at both metric and topological measures of the structure and complexity of these networks---particularly focusing on density, connectedness, and resilience.
 These preliminary empirical findings demonstrate the use of OSMnx as a new piece of research infrastructure.
 They suggest to urban planners new methods for acquiring and analyzing street network data, including new methods for evaluating network resilience and resilience gains with betweenness centralities and average node connectivities.
-Finally, this study has made all of these network datasets – for 497 urbanized areas, 19,655 cities and towns, and 6,857 neighborhoods – along with all of their attribute data and complexity measures available in an online public repository for other researchers to study and repurpose.
+Finally, this study has made all of these network datasets---for 497 urbanized areas, 19,655 cities and towns, and 6,857 neighborhoods---along with all of their attribute data and complexity measures available in an online public repository for other researchers to study and repurpose.
 
@@
 
-\begin{thebibliography}{}
-
 \bibliography{references}
 
-\end{thebibliography}
-
 \end{document}

### 2017-03-21 latex/paper.tex
update cite style

@@
-\documentclass[Afour,sageh,times]{sage/sagej}
+\documentclass{article}
 
-\usepackage{moreverb,url}
+\usepackage{graphicx}
 \usepackage{csvsimple}
-\usepackage[colorlinks,bookmarksopen,bookmarksnumbered,citecolor=red,urlcolor=red]{hyperref}
+\usepackage[numbers]{natbib}
+\usepackage[affil-it]{authblk}
 
-\bibliographystyle{plainnat}
-\setcitestyle{numbers}
+\bibliographystyle{apalike}
 
-\newcommand\BibTeX{{\rmfamily B\kern-.05em \textsc{i\kern-.025em b}\kern-.08em
-T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
-
-\def\volumeyear{2017}
 \begin{document}
-\runninghead{Boeing}
-\title{A Multi-Scale Analysis of 27,000 Urban Street Networks}
-\author{Geoff Boeing\affilnum{1}}
-\affiliation{\affilnum{1}PhD Candidate, Department of City and Regional Planning, University of California, Berkeley}
-\corrauth{Geoff Boeing,
-Department of City and Regional Planning,
-University of California, Berkeley,
-United States}
-\email{gboeing@berkeley.edu}
+\title{A Multi-Scale Analysis of 27,000 Urban Street Networks\footnote{This is a journal article manuscript currently under peer review.}}
+\author{Geoff Boeing\footnote{Email address: gboeing@berkeley.edu; corresponding author.}}
+\affil{Department of City and Regional Planning\\University of California, Berkeley}
+\date{\today}
+\maketitle
 
 \begin{abstract}
-This paper serves illustrative purposes as it presents wide-ranging preliminary empirical findings on US urban form and street network characteristics, using OSMnx.
-It also demonstrates the scalability and flexibility of OSMnx as a new research tool.
-This paper addresses past limitations by conducting an analysis of street networks with large sample sizes, with clearly defined network definitions and extents for reproducibility, and using non-planar, directed graphs.
-In particular, it examines urban street networks focusing on structure, connectedness, centrality, and robustness.
-In total it cross-sectionally analyzes 497 urbanized areas' street networks, 19,655 cities' and towns' street networks, and 6,857 neighborhoods' street networks.
-These sample sizes are larger than those in any previous study and the preliminary empirical findings illustrate the use of OSMnx as a new research platform.
+OpenStreetMap presents an emerging source of worldwide geospatial urban data useful to researchers.
+This paper uses the OSMnx software to automatically download and analyze 27,000 street networks from OpenStreetMap at metropolitan, municipal, and neighborhood scales---that is, every urbanized area, incorporated place, and Zillow neighborhood in the US.
+In turn it presents wide-ranging empirical findings on US urban form and street network characteristics.
+This paper addresses past limitations by conducting an analysis of street networks with an extremely large sample sizes, with clearly defined network definitions and extents for reproducibility, and using non-planar, directed graphs.
+Its analysis emphasizes urban street networks measures relevant to urban design and morphology, including structure, connectedness, centrality, and robustness.
+In total we cross-sectionally analyze the street networks of 497 urbanized areas, 19,655 cities and towns, and 6,857 neighborhoods.
+These sample sizes are larger than those in any previous study and illustrate the use of OSMnx and OpenStreetMap as a new urban form research platform.
 \end{abstract}
-
-\keywords{network analysis, street network, urban design, urban form, urban planning}
-
-\maketitle
 
@@
 
-\begin{table*}[h]
+\begin{table*}[hp]
 \caption{Central tendency and statistical dispersion for selected measures of all US urbanized areas' street networks.}
@@
 
-\begin{figure}[h]
-\includegraphics[width=0.45\textwidth]{media/fig02.png}
-\caption{Distribution of node types in 9 urbanized areas, with number of streets emanating from the node on the \emph{x}-axis and proportion of nodes of this type on the \emph{y}-axis }
+\begin{figure*}[hp]
+\includegraphics[width=1\textwidth]{media/fig02.png}
+\caption{Distribution of node types in 9 urbanized areas, with number of streets emanating from the node on the $x$-axis and proportion of nodes of this type on the $y$-axis }
 \label{fig02}
-\end{figure}
+\end{figure*}
 
@@
 
-\begin{table*}[h]
+\begin{table*}[hp]
 \caption{Selected measures of the 30 largest (by land area) urbanized areas' street networks.}
@@
 
-\begin{table*}[h]
+\begin{table*}[hp]
 \caption{Selected summary stats for every incorporated city and town in the US.}
@@
 
-\begin{figure}[h]
-	\includegraphics[width=0.45\textwidth]{media/fig03-v.png}
+\begin{figure*}[h]
+	\includegraphics[width=1\textwidth]{media/fig03.png}
 	\caption{The linear relationship between total street length and number of nodes in the street networks of 6,857 US neighborhoods (top) and 19,655 US cities and towns (bottom).}
 	\label{fig03}
-\end{figure}
+\end{figure*}
 
@@
 
-\begin{table*}[h]
+\begin{table*}[hp]
 \caption{Median values, aggregated by state plus DC, of selected measures of the municipal-scale street networks for every city and town in the US.}
@@
 
-\begin{table*}[h]
+\begin{table*}[hp]
 \caption{Selected summary stats for all the neighborhood-scale street networks.}
@@
 Middle: top, downtown Portland; bottom, its suburb, Beaverton.
-Right: top, downtown San Francisco; bottom, its suburb Concord.}
+Right: top, downtown San Francisco; bottom, its suburb, Concord.}
 	\label{fig05}
@@
 
-\begin{acks}
-Redacted
-\end{acks}
-
 \bibliography{references}

### 2017-03-26 latex/paper.tex
clean up tables

@@
 \usepackage{graphicx}
-\usepackage{csvsimple}
+\usepackage{geometry}
+\usepackage{booktabs}
+\usepackage{pdflscape}
 \usepackage[numbers]{natbib}
@@
 
-\begin{table*}[hp]
-\caption{Central tendency and statistical dispersion for selected measures of all US urbanized areas' street networks.}
-\label{table01}
-\csvautotabular{media/table01.csv}
-\end{table*}
-
 Table \ref{table01} shows the considerable variation in street network characteristics across the entire data set of 497 urbanized areas.
@@
 
-\begin{figure*}[h]
+\begin{figure}
 \includegraphics[width=1\textwidth]{media/fig01.png}
@@
 \label{fig01}
-\end{figure*}
+\end{figure}
 
@@
 
-\begin{figure*}[hp]
-\includegraphics[width=1\textwidth]{media/fig02.png}
+\begin{table}
+\centering
+\caption{Central tendency and statistical dispersion for selected measures of all US urbanized areas' street networks.}
+\label{table01}
+\begin{tabular}{llllll}
+	\toprule
+	measure                                     & mean           & std dev        & min            & median         & max            \\
+	\midrule
+	Area (km2)                                  & 460.657        & 858.125        & 25.685         & 184.898        & 8937.429       \\
+	Avg of the avg neighborhood degree          & 2.886          & 0.109          & 2.626          & 2.875          & 3.228          \\
+	Avg of the avg weighted neighborhood degree & 0.032          & 0.018          & 0.021          & 0.03           & 0.321          \\
+	Avg circuity                                & 1.076          & 0.019          & 1.023          & 1.074          & 1.14           \\
+	Avg clustering coefficient                  & 0.042          & 0.009          & 0.015          & 0.042          & 0.071          \\
+	Avg weighted clustering coefficient         & 0.002          & 0.001          & \textless0.001 & 0.001          & 0.006          \\
+	Intersection count                          & 12582          & 26054          & 751            & 4593           & 307848         \\
+	Avg degree centrality                       & 0.001          & 0.001          & \textless0.001 & 0.001          & 0.007          \\
+	Edge density (km/km2)                       & 13.455         & 2.137          & 7.961          & 13.352         & 21.233         \\
+	Avg edge length (m)                         & 158.588        & 17.653         & 117.341        & 157.332        & 223.08         \\
+	Total edge length (km)                      & 6353           & 12625          & 427            & 2393           & 1.42E+08       \\
+	Proportion of dead-ends                     & 0.213          & 0.055          & 0.077          & 0.207          & 0.416          \\
+	Proportion of 3-way intersections           & 0.593          & 0.046          & 0.444          & 0.591          & 0.778          \\
+	Proportion of 4-way intersections           & 0.187          & 0.063          & 0.054          & 0.178          & 0.422          \\
+	Intersection density (per km2)              & 26.469         & 6.256          & 12.469         & 26.029         & 49.423         \\
+	Average node degree                         & 5.153          & 0.302          & 4.307          & 5.143          & 6.056          \\
+	m                                           & 40890          & 83678          & 2516           & 14955          & 981646         \\
+	n                                           & 16032          & 32585          & 874            & 5830           & 373309         \\
+	Node density (per km2)                      & 33.628         & 7.641          & 17.675         & 33.071         & 61.655         \\
+	Max PageRank value                          & 0.001          & 0.001          & \textless0.001 & 0.001          & 0.003          \\
+	Min PageRank value                          & \textless0.001 & \textless0.001 & \textless0.001 & \textless0.001 & \textless0.001 \\
+	Self-loop proportion                        & 0.008          & 0.008          & \textless0.001 & 0.006          & 0.071          \\
+	Street density (km/km2)                     & 7.262          & 1.221          & 4.217          & 7.171          & 11.797         \\
+	Average street segment length (m)           & 161.331        & 17.765         & 119.573        & 160.288        & 225.92         \\
+	Total street length (km)                    & 3480           & 7026           & 222            & 1269           & 79046          \\
+	Street segment count                        & 22011          & 45725          & 1281           & 7868           & 533757         \\
+	Average streets per node                    & 2.764          & 0.162          & 2.223          & 2.77           & 3.217          \\
+	\bottomrule
+\end{tabular}
+\end{table}
+
+\begin{figure}
+\includegraphics[width=0.5\textwidth]{media/fig02.png}
 \caption{Distribution of node types in 9 urbanized areas, with number of streets emanating from the node on the $x$-axis and proportion of nodes of this type on the $y$-axis }
 \label{fig02}
-\end{figure*}
+\end{figure}
 
@@
 
-\begin{table*}[hp]
+\newgeometry{margin=1in}
+\begin{landscape}
+\begin{table}
+\centering
 \caption{Selected measures of the 30 largest (by land area) urbanized areas' street networks.}
 \label{table02}
-\csvautotabular{media/table02.csv}
-\end{table*}
+\begin{tabular}{p{2.5cm} p{1.5cm} p{1.5cm} p{1.5cm} p{1.55cm} p{1.5cm} p{1.5cm} p{1.5cm} p{1.5cm} p{1.5cm} p{1.5cm}}
+\hline
+urban area core city & land area km2 & avg circuity & avg clustering coefficient & dead-end ratio & 3-way ratio & 4-way ratio & intersect density/ km2 & street density km/km2 & avg street length (m) & avg streets/ node \\
+\hline
+New York             & 8937          & 1.06         & 0.04                       & 0.18           & 0.62        & 0.2         & 34.44                 & 8.84                  & 148                   & 2.86             \\
+Atlanta              & 6850          & 1.1          & 0.04                       & 0.32           & 0.58        & 0.09        & 18.39                 & 6.16                  & 186                   & 2.45             \\
+Chicago              & 6325          & 1.07         & 0.04                       & 0.17           & 0.57        & 0.25        & 27.05                 & 7.77                  & 163                   & 2.92             \\
+Philadelphia         & 5132          & 1.08         & 0.05                       & 0.17           & 0.63        & 0.2         & 26.65                 & 7.3                   & 159                   & 2.87             \\
+Boston               & 4852          & 1.09         & 0.05                       & 0.2            & 0.68        & 0.11        & 24.23                 & 6.44                  & 154                   & 2.71             \\
+Dallas               & 4612          & 1.07         & 0.05                       & 0.15           & 0.61        & 0.23        & 34.16                 & 9.16                  & 156                   & 2.95             \\
+Los Angeles          & 4497          & 1.06         & 0.03                       & 0.21           & 0.56        & 0.22        & 39.45                 & 10.59                 & 151                   & 2.82             \\
+Houston              & 4303          & 1.08         & 0.04                       & 0.2            & 0.57        & 0.22        & 33.49                 & 8.62                  & 145                   & 2.83             \\
+Detroit              & 3461          & 1.07         & 0.04                       & 0.15           & 0.63        & 0.22        & 31.1                  & 8.56                  & 159                   & 2.93             \\
+Washington           & 3424          & 1.09         & 0.04                       & 0.26           & 0.56        & 0.17        & 31.22                 & 8.26                  & 146                   & 2.66             \\
+Miami                & 3204          & 1.1          & 0.05                       & 0.17           & 0.59        & 0.23        & 40.54                 & 10.61                 & 149                   & 2.89             \\
+Phoenix              & 2968          & 1.09         & 0.05                       & 0.2            & 0.62        & 0.17        & 35.31                 & 9.1                   & 150                   & 2.77             \\
+Minneapolis          & 2647          & 1.08         & 0.05                       & 0.19           & 0.57        & 0.23        & 29.54                 & 8.68                  & 167                   & 2.84             \\
+Seattle              & 2617          & 1.07         & 0.03                       & 0.3            & 0.54        & 0.16        & 31.57                 & 8.2                   & 143                   & 2.57             \\
+Tampa                & 2479          & 1.1          & 0.05                       & 0.2            & 0.58        & 0.21        & 31.35                 & 8.46                  & 153                   & 2.83             \\
+St.
+Louis            & 2392          & 1.1          & 0.04                       & 0.22           & 0.62        & 0.15        & 29.68                 & 8.16                  & 154                   & 2.73             \\
+Pittsburgh           & 2345          & 1.09         & 0.04                       & 0.23           & 0.6         & 0.16        & 23.57                 & 6.71                  & 165                   & 2.72             \\
+San Juan             & 2245          & 1.11         & 0.02                       & 0.36           & 0.56        & 0.08        & 26.57                 & 6.43                  & 131                   & 2.36             \\
+Cincinnati           & 2040          & 1.07         & 0.03                       & 0.31           & 0.54        & 0.14        & 17.96                 & 6.1                   & 186                   & 2.51             \\
+Cleveland            & 2004          & 1.07         & 0.04                       & 0.19           & 0.66        & 0.14        & 19.13                 & 6.51                  & 198                   & 2.76             \\
+Charlotte            & 1920          & 1.08         & 0.04                       & 0.3            & 0.57        & 0.11        & 21                    & 6.43                  & 170                   & 2.51             \\
+San Diego            & 1897          & 1.08         & 0.03                       & 0.28           & 0.54        & 0.17        & 28.89                 & 8.32                  & 159                   & 2.62             \\
+Baltimore            & 1857          & 1.09         & 0.04                       & 0.23           & 0.59        & 0.17        & 27.72                 & 7.56                  & 152                   & 2.72             \\
+Indianapolis         & 1828          & 1.08         & 0.05                       & 0.23           & 0.59        & 0.17        & 27.62                 & 7.63                  & 157                   & 2.7              \\
+Kansas City          & 1756          & 1.06         & 0.04                       & 0.21           & 0.58        & 0.2         & 32.09                 & 8.57                  & 152                   & 2.79             \\
+Denver               & 1729          & 1.07         & 0.05                       & 0.2            & 0.57        & 0.22        & 40.6                  & 9.84                  & 138                   & 2.84             \\
+Orlando              & 1548          & 1.11         & 0.06                       & 0.2            & 0.61        & 0.18        & 26.3                  & 7.44                  & 163                   & 2.79             \\
+San Antonio          & 1547          & 1.07         & 0.05                       & 0.17           & 0.6         & 0.21        & 28.33                 & 7.91                  & 162                   & 2.87             \\
+Nashville            & 1460          & 1.08         & 0.03                       & 0.27           & 0.59        & 0.14        & 19.08                 & 6.1                   & 181                   & 2.6              \\
+Milwaukee            & 1413          & 1.06         & 0.06                       & 0.14           & 0.55        & 0.3         & 28.27                 & 7.81                  & 157                   & 3.03 \\
+\bottomrule
+\end{tabular}
+\end{table}
+\end{landscape}
+\restoregeometry
 
@@
 
-\begin{table*}[hp]
+\begin{table}
+\centering
 \caption{Selected summary stats for every incorporated city and town in the US.}
 \label{table03}
-\csvautotabular{media/table03.csv}
-\end{table*}
-
-\begin{figure*}[h]
-	\includegraphics[width=1\textwidth]{media/fig03.png}
-	\caption{The linear relationship between total street length and number of nodes in the street networks of 6,857 US neighborhoods (top) and 19,655 US cities and towns (bottom).}
-	\label{fig03}
-\end{figure*}
-
-\begin{figure*}[h]
-	\includegraphics[width=1\textwidth]{media/fig04.png}
-	\caption{Left: contiguous US states by median of mean streets per node in municipal street networks, from lowest (pale yellow) to highest (dark red).
-Right: contiguous US states by median of mean street segment length in municipal street networks, from longest (pale yellow) to shortest (dark red).}
-	\label{fig04}
-\end{figure*}
+\begin{tabular}{llllll}
+	\toprule
+	measure                                     & mean    & std dev & min            & median         & max       \\
+	\midrule
+	Area (km2)                                  & 16.703  & 107.499 & 0.039          & 3.918          & 7434.258  \\
+	Avg of the avg neighborhood degree          & 2.94    & 0.297   & 0.4            & 2.953          & 3.735     \\
+	Avg of the avg weighted neighborhood degree & 0.033   & 0.141   & \textless0.001 & 0.029          & 9.357     \\
+	Avg circuity                                & 1.067   & 0.159   & 1              & 1.055          & 20.452    \\
+	Avg clustering coefficient                  & 0.048   & 0.041   & \textless0.001 & 0.04           & 1         \\
+	Avg weighted clustering coefficient         & 0.01    & 0.018   & \textless0.001 & 0.005          & 0.524     \\
+	Intersection count                          & 324     & 1266    & 0              & 83             & 62996     \\
+	Avg degree centrality                       & 0.093   & 0.136   & \textless0.001 & 0.052          & 2.667     \\
+	Edge density (km/km2)                       & 12.654  & 6.705   & 0.006          & 11.814         & 58.603    \\
+	Avg edge length (m)                         & 161.184 & 80.769  & 25.822         & 144.447        & 3036.957  \\
+	Total edge length (km)                      & 159.067 & 578.521 & 0.052          & 40.986         & 24728.326 \\
+	Proportion of dead-ends                     & 0.192   & 0.093   & \textless0.001 & 0.184          & 1         \\
+	Proportion of 3-way intersections           & 0.572   & 0.11    & \textless0.001 & 0.579          & 1         \\
+	Proportion of 4-way intersections           & 0.237   & 0.129   & \textless0.001 & 0.217          & 1         \\
+	Intersection density (per km2)              & 29.363  & 21.607  & \textless0.001 & 24.719         & 259.647   \\
+	Average node degree                         & 5.251   & 0.668   & 0.8            & 5.268          & 7.166     \\
+	m                                           & 1046    & 3924    & 2              & 275            & 176161    \\
+	n                                           & 401     & 1516    & 2              & 103            & 71993     \\
+	Node density (per km2)                      & 35.449  & 24.409  & 0.047          & 30.718         & 296.74    \\
+	Max PageRank value                          & 0.034   & 0.046   & \textless0.001 & 0.021          & 0.87      \\
+	Min PageRank value                          & 0.005   & 0.018   & \textless0.001 & 0.002          & 0.5       \\
+	Self-loop proportion                        & 0.005   & 0.015   & \textless0.001 & \textless0.001 & 1         \\
+	Street density (km/km2)                     & 6.528   & 3.435   & 0.003          & 6.109          & 29.302    \\
+	Average street segment length (m)           & 162.408 & 81.035  & 25.822         & 145.479        & 3036.957  \\
+	Total street length (km)                    & 86.096  & 331.048 & 0.026          & 21.005         & 15348.008 \\
+	Street segment count                        & 558     & 2208    & 1              & 140            & 107393    \\
+	Average streets per node                    & 2.851   & 0.282   & 1              & 2.852          & 4         \\
+	\bottomrule
+\end{tabular}
+\end{table}
 
@@
 
-\begin{table*}[hp]
+\begin{figure}
+	\includegraphics[width=0.7\textwidth]{media/fig03.png}
+	\caption{The linear relationship between total street length and number of nodes in the street networks of 6,857 US neighborhoods (top) and 19,655 US cities and towns (bottom).}
+	\label{fig03}
+\end{figure}
+
+\begin{figure}
+	\includegraphics[width=0.85\textwidth]{media/fig04.png}
+	\caption{Left: contiguous US states by median of mean streets per node in municipal street networks, from lowest (pale yellow) to highest (dark red).
+Right: contiguous US states by median of mean street segment length in municipal street networks, from longest (pale yellow) to shortest (dark red).}
+	\label{fig04}
+\end{figure}
+
+\newgeometry{margin=2cm}
+\begin{table}
+\centering
 \caption{Median values, aggregated by state plus DC, of selected measures of the municipal-scale street networks for every city and town in the US.}
 \label{table04}
-\csvautotabular{media/table04.csv}
-\end{table*}
+\begin{tabular}{p{1in}p{1in}p{1in}p{1in}p{1in}}
+	\toprule
+	State & Intersect density (per km2) & Avg streets per node & Avg circuity & Avg street segment length \\
+	\midrule
+	AK    & 1.28                        & 2.43                 & 1.1          & 223.5                     \\
+	AL    & 9.7                         & 2.64                 & 1.07         & 190.81                    \\
+	AR    & 15.75                       & 2.78                 & 1.06         & 166.32                    \\
+	AZ    & 12.45                       & 2.77                 & 1.08         & 171.8                     \\
+	CA    & 32.58                       & 2.74                 & 1.07         & 143.79                    \\
+	CO    & 29.26                       & 2.88                 & 1.06         & 136.68                    \\
+	CT    & 28.05                       & 2.7                  & 1.07         & 165.87                    \\
+	DC    & 58.91                       & 3.26                 & 1.04         & 122.23                    \\
+	DE    & 25.3                        & 2.8                  & 1.06         & 127.8                     \\
+	FL    & 26.26                       & 2.87                 & 1.07         & 150.75                    \\
+	GA    & 15.25                       & 2.78                 & 1.07         & 177.5                     \\
+	HI    & 8                           & 2.42                 & 1.07         & 177.93                    \\
+	IA    & 24.08                       & 3.02                 & 1.04         & 129.36                    \\
+	ID    & 33.85                       & 2.91                 & 1.06         & 132.08                    \\
+	IL    & 29.02                       & 2.93                 & 1.05         & 137.77                    \\
+	IN    & 35.25                       & 2.93                 & 1.05         & 125.72                    \\
+	KS    & 43.94                       & 3.14                 & 1.04         & 124.39                    \\
+	KY    & 25.12                       & 2.68                 & 1.07         & 151.28                    \\
+	LA    & 17.14                       & 2.79                 & 1.06         & 162.62                    \\
+	MA    & 32.33                       & 2.76                 & 1.07         & 135.98                    \\
+	MD    & 28.67                       & 2.79                 & 1.07         & 133.69                    \\
+	ME    & 7.69                        & 2.67                 & 1.07         & 198.93                    \\
+	MI    & 20.93                       & 2.9                  & 1.05         & 153.5                     \\
+	MN    & 18.96                       & 2.87                 & 1.06         & 152.92                    \\
+	MO    & 29.87                       & 2.89                 & 1.06         & 138.29                    \\
+	MS    & 14.76                       & 2.75                 & 1.06         & 174.86                    \\
+	MT    & 38.94                       & 3.11                 & 1.04         & 126.89                    \\
+	NC    & 19.28                       & 2.65                 & 1.06         & 166.69                    \\
+	ND    & 34.28                       & 3.07                 & 1.04         & 123.93                    \\
+	NE    & 45.89                       & 3.16                 & 1.04         & 119.79                    \\
+	NH    & 12.22                       & 2.69                 & 1.1          & 175.88                    \\
+	NJ    & 44.98                       & 2.88                 & 1.04         & 130.79                    \\
+	NM    & 18.5                        & 2.93                 & 1.05         & 152.02                    \\
+	NV    & 13.86                       & 2.77                 & 1.07         & 147.35                    \\
+	NY    & 21.89                       & 2.75                 & 1.06         & 156.88                    \\
+	OH    & 25.23                       & 2.8                  & 1.05         & 142.08                    \\
+	OK    & 28.22                       & 3.03                 & 1.05         & 139.5                     \\
+	OR    & 35.08                       & 2.69                 & 1.06         & 121.18                    \\
+	PA    & 35.69                       & 2.87                 & 1.05         & 128.34                    \\
+	RI    & 56.23                       & 2.86                 & 1.05         & 110.35                    \\
+	SC    & 18.76                       & 2.81                 & 1.06         & 169.21                    \\
+	SD    & 32.01                       & 3.12                 & 1.04         & 130.75                    \\
+	TN    & 13.62                       & 2.71                 & 1.07         & 192.83                    \\
+	TX    & 23.85                       & 2.92                 & 1.05         & 160.44                    \\
+	UT    & 12.58                       & 2.71                 & 1.06         & 191.04                    \\
+	VA    & 25.18                       & 2.63                 & 1.08         & 145.65                    \\
+	VT    & 18.91                       & 2.55                 & 1.08         & 145.18                    \\
+	WA    & 28.71                       & 2.75                 & 1.06         & 134.02                    \\
+	WI    & 17.87                       & 2.81                 & 1.06         & 156.19                    \\
+	WV    & 28.45                       & 2.67                 & 1.08         & 136.57                    \\
+	WY    & 23.48                       & 2.92                 & 1.06         & 143.63                    \\
+	\bottomrule
+\end{tabular}
+\end{table}
+\restoregeometry
 
@@
 
-\begin{table*}[hp]
+\begin{table}
+\centering
 \caption{Selected summary stats for all the neighborhood-scale street networks.}
 \label{table05}
-\csvautotabular{media/table05.csv}
-\end{table*}
+\begin{tabular}{llllll}
+	\toprule
+	& mean    & std dev       & min            & median         & max      \\
+	\midrule
+	Area (km2)                                  & 5.322   & 15.463  & 0.008          & 1.738          & 323.306  \\
+	Avg of the avg neighborhood degree          & 2.598   & 0.436   & \textless0.001 & 2.67           & 3.632    \\
+	Avg of the avg weighted neighborhood degree & 0.031   & 0.041   & \textless0.001 & 0.029          & 2.991    \\
+	Avg circuity                                & 1.08    & 0.411   & 1              & 1.044          & 24.29    \\
+	Avg clustering coefficient                  & 0.044   & 0.055   & \textless0.001 & 0.034          & 1        \\
+	Avg weighted clustering coefficient         & 0.01    & 0.027   & \textless0.001 & 0.005          & 0.799    \\
+	Intersection count                          & 173     & 379     & 0              & 76             & 8371     \\
+	Avg degree centrality                       & 0.13    & 0.27    & 0.001          & 0.054          & 4        \\
+	Edge density (km/km2)                       & 17.569  & 7.095   & 0.025          & 18.152         & 59.939   \\
+	Avg edge length (m)                         & 142.279 & 59.182  & 8.447          & 133.848        & 2231.331 \\
+	Total edge length (km)                      & 71.369  & 166.566 & 0.017          & 29.88          & 3563.409 \\
+	Proportion of dead-ends                     & 0.17    & 0.131   & \textless0.001 & 0.145          & 1        \\
+	Proportion of 3-way intersections           & 0.559   & 0.146   & \textless0.001 & 0.574          & 1        \\
+	Proportion of 4-way intersections           & 0.275   & 0.176   & \textless0.001 & 0.234          & 1        \\
+	Intersection density (per km2)              & 49.497  & 28.33   & \textless0.001 & 46.43          & 444.355  \\
+	Average node degree                         & 4.675   & 0.836   & 0.545          & 4.736          & 7.283    \\
+	m                                           & 5201    & 1185    & 1              & 217            & 27289    \\
+	n                                           & 208     & 459     & 2              & 90             & 9327     \\
+	Node density (per km2)                      & 58.677  & 31.802  & 0.063          & 55.626         & 499.9    \\
+	Max PageRank value                          & 0.055   & 0.086   & \textless0.001 & 0.026          & 0.889    \\
+	Min PageRank value                          & 0.01    & 0.041   & \textless0.001 & 0.002          & 0.5      \\
+	Self-loop proportion                        & 0.007   & 0.034   & \textless0.001 & \textless0.001 & 1        \\
+	Street density (km/km2)                     & 9.744   & 4.085   & 0.013          & 9.882          & 33.737   \\
+	Average street segment length (m)           & 143.664 & 60.023  & 7.376          & 134.877        & 2231.331 \\
+	Total street length (km)                    & 40.049  & 93.987  & 0.009          & 16.248         & 1960.643 \\
+	Street segment count                        & 288     & 656     & 1              & 119            & 14754    \\
+	Average streets per node                    & 2.925   & 0.408   & 1              & 2.944          & 4.026    \\
+	\bottomrule
+\end{tabular}
+\end{table}
 
@@
 
-\begin{figure*}[h]
+\begin{figure}
 	\includegraphics[width=1\textwidth]{media/fig05.png}
@@
 	\label{fig05}
-\end{figure*}
+\end{figure}
 

### 2017-03-26 latex/paper.tex
update intro and methods

@@
 \begin{abstract}
-OpenStreetMap presents an emerging source of worldwide geospatial urban data useful to researchers.
+
+OpenStreetMap is an emerging source of worldwide geospatial data useful to researchers.
 This paper uses the OSMnx software to automatically download and analyze 27,000 street networks from OpenStreetMap at metropolitan, municipal, and neighborhood scales---that is, every urbanized area, incorporated place, and Zillow neighborhood in the US.
@@
 This paper addresses past limitations by conducting an analysis of street networks with an extremely large sample sizes, with clearly defined network definitions and extents for reproducibility, and using non-planar, directed graphs.
-Its analysis emphasizes urban street networks measures relevant to urban design and morphology, including structure, connectedness, centrality, and robustness.
+Its analysis emphasizes urban street network measures relevant to urban design and morphology, including structure, connectedness, centrality, and robustness.
 In total we cross-sectionally analyze the street networks of 497 urbanized areas, 19,655 cities and towns, and 6,857 neighborhoods.
 These sample sizes are larger than those in any previous study and illustrate the use of OSMnx and OpenStreetMap as a new urban form research platform.
+
 \end{abstract}
@@
 \section{Introduction}
+
 On May 20, 1862, Abraham Lincoln signed the Homestead Act into law, making land across the United States Midwest and Great Plains available for free to all applicants \cite{porterfield_homestead_2005}.
@@
 
-A cross-sectional analysis of American urban form through its street networks at metropolitan, municipal, and neighborhood scales can reveal these artifacts and histories across the nation.
-Network analysis is a natural approach to the study of cities as growing, complex systems that self-organize to fill space with a capillary circulation network \cite{masucci_random_2009}.
-The empirical literature on street network analysis is growing ever richer, but suffers from some limitations---discussed in \cite{boeing_osmnx:_2017} and summarized here.
-First, sample sizes tend to be fairly small due to data availability, gathering, and processing constraints.
-Second, reproducibility is difficult when the dozens of decisions that go into analysis---such as spatial extents, topological simplification and correction, definitions of nodes and edges, etc.---are ad hoc or only partly reported.
+A cross-sectional analysis of American urban form can reveal these artifacts and histories through street networks at metropolitan, municipal, and neighborhood scales.
+Network analysis is a natural approach to the study of cities as complex systems \cite{masucci_random_2009}.
+The empirical literature on street network analysis is growing ever richer, but suffers from some limitations---discussed in detail in \cite{boeing_osmnx:_2017} and summarized here.
+First, sample sizes have tended to be fairly small due to data availability, gathering, and processing constraints.
+Second, reproducibility has been difficult when the dozens of decisions that go into analysis---such as spatial extents, topological simplification and correction, definitions of nodes and edges, etc.---are ad hoc or only partly reported.
 Third, and related to the first two, studies frequently oversimplify to planar or undirected primal graphs for tractability, or use dual graphs despite the loss of geographic and metric information.
 
-This paper addresses these limitations by conducting an analysis of street networks at multiple scales, with large sample sizes, with clearly defined network definitions and extents for reproducibility, and using non-planar, directed graphs.
-In particular, it examines urban street networks---represented here as primal, non-planar, weighted multidigraphs with possible self-loops---focusing on structure, connectedness, centrality, and resilience.
-Most studies in the street network literature that conduct topological and/or metric analysis tend to have sample sizes ranging around 5 to 50 networks.
-This paper instead conducts a large analysis of over 27,000 urban street networks at multiple overlapping scales across the United States.
-Namely, it examines the street networks of every US incorporated city and town, urbanized area, and Zillow-defined neighborhood.
-In total, this study presented uses OSMnx \cite{boeing_osmnx:_2017} to download, construct, and analyze 497 urbanized areas' street networks, 19,655 cities' and towns' street networks, and 6,857 neighborhoods' street networks.
-It uses these street networks to conduct four analyses: at the metropolitan scale, at the municipal scale, at the neighborhood scale, and with a case study looking deeper at the neighborhood-scale street networks in the city of San Francisco.
+This paper addresses these limitations by conducting an urban morphological analysis of street networks at multiple scales, with large sample sizes, with clearly defined network definitions and extents for reproducibility, and using non-planar, directed graphs.
+In particular, it examines urban street network---represented here as primal, non-planar, weighted multidigraphs with possible self-loops---focusing on structure, connectedness, centrality, and resilience.
+Most studies in the street network literature that conduct topological and/or metric analysis tend to have sample sizes ranging around 10 to 50 networks.
+This study instead analyzes over 27,000 urban street networks at multiple overlapping scales across the United States.
+Namely, it examines the street networks of every incorporated city and town (n=19,655), census urbanized area (n=497), and Zillow-defined neighborhood (n=6,857) in the US.
+To do so, it uses OSMnx \cite{boeing_osmnx:_2017}---a new street network research toolkit---to download, construct, and analyze these street networks at metropolitan, municipal, and neighborhood scales.
 
+The purpose of this paper is threefold.
+First, it demonstrates a new methodology for easily and consistently acquiring, constructing, and analyzing large samples of street networks as nonplanar directed graphs.
+Second, it presents empirical findings of descriptive urban morphology for the street networks of every city, urbanized area, and Zillow neighborhood in the US.
+Third, it investigates with large sample sizes some claims previously made in the research literature with small sample sizes.
 This paper is organized as follows.
-In the next section, it presents the data sources, tools, and methods used to collect, construct, and analyze these street networks.
-Then it presents findings of the analyses at three spatial scales: metropolitan, municipal, and neighborhood.
-Finally, it concludes with a discussion of these findings and their implications for street network analysis, urban form complexity, and city planning.
+In the next section, it discusses the data sources, tools, and methods used to collect, construct, and analyze these street networks.
+Next, it presents findings of the analyses at metropolitan, municipal, and neighborhood spatial scales.
+Finally, it concludes with a discussion of these findings and their implications for street network analysis, urban morphology, and city planning.
 
 \section{Methodology}
+
+A network (also called a \emph{graph} in mathematics) comprises a set of nodes connected to one another by a set of edges.
+Street networks can be conceptualized as primal, directed, nonplanar graphs.
+A primal street network represents intersections as nodes and street segments as edges.
+A directed network has directed edges: that is, edge $uv$ points one-way from node $u$ to node $v$, but there need not exist a reciprocal edge $vu$.
+A planar network can be represented in two dimensions with its edges intersecting only at nodes.
+Most street networks are nonplanar---due to grade-separated expressways, overpasses, bridges, and tunnels---but most quantitative studies of urban street networks represent them as planar (e.g., Buhl et al.
+2006; Cardillo et al.
+2006; Barthélemy and Flammini 2008; Masucci et al.
+2009; Strano et al.
+2013) for tractability because bridges and tunnels are reasonably uncommon in some cities.
+Planarity may reasonably model the street networks of old European town centers, but poorly models the street network in more modern cities like Los Angeles with their grade-separated expressways, bridges, and tunnels.
+Thus, a planar simplification could yield inaccurate analyses.
+
 This study uses OSMnx to download, construct, correct, analyze, and visualize street networks at metropolitan, municipal, and neighborhood scales \cite{boeing_osmnx:_2017}.
+OSMnx is a Python-based research tool that easily downloads OpenStreetMap data for any place in the world and constructs it into a spatially-embedded graph-theoretic object for analysis and visualization.
+OpenStreetMap is a collaborative worldwide mapping project that makes its spatial data available via various APIs.
+These data are of high quality and compare well to CIA World Factbook estimates and US Census TIGER/Line data (Haklay 2010; Over et al.
+2010; Zielstra and Hochmair 2011; Maron 2015).
+In 2007, OpenStreetMap imported the 2005 TIGER/Line roads data, and since then many community-led corrections and improvements have been made (Willis 2008).
+Many of these additions go beyond TIGER/Line's scope, including passageways between buildings, footpaths through parks, bike paths, and detailed feature attributes, such as finer-grained street classifiers, speed limits, etc.
+
 To define the study sites and their spatial boundaries, this study uses three sets of geometries.
@@
 Finally, it automatically calculates metric and topological measures for each.
-However, in a planar graph, topological and metric properties are somewhat interrelated \cite{masucci_random_2009}.
+Such measures extend the toolkit commonly used by urban morphologists (Talen 2003).
+Brief descriptions of these measures are discussed here, but extended definitions and algorithms can be found in Newman (2010) and Barthélemy (2011).
 
@@
 In total, this study cross-sectionally analyzes 27,009 networks: 497 urbanized areas' street networks, 19,655 cities' and towns' street networks, and 6,857 neighborhoods' street networks.
-These sample sizes are larger than those in any previous study.
+These sample sizes are larger than those in any previous similar study.
 In the following sections, we present the findings of these analyses at the metropolitan scale, the municipal scale, the neighborhood scale, and through a case study looking deeper at the neighborhood-scale street networks in the city of San Francisco.
@@
 \begin{figure}
-\includegraphics[width=0.5\textwidth]{media/fig02.png}
+\includegraphics[width=0.75\textwidth]{media/fig02.png}
 \caption{Distribution of node types in 9 urbanized areas, with number of streets emanating from the node on the $x$-axis and proportion of nodes of this type on the $y$-axis }
@@
 The grid-like San Angelo, TX urbanized area has the most streets per node (3.2) on average, and (outside of Puerto Rico, which contains the seven lowest urbanized areas) the sprawling, disconnected Lexington Park, MD urbanized area has the fewest (2.2).
-These two urban areas fit the trend seen in the spatial distribution across the US in Figure \ref{fig01}: urbanized areas in the great plains and Midwest have particularly high numbers of streets per node on average, indicating more grid-like, connected networks.
+These two urbanized areas fit the trend seen in the spatial distribution across the US in Figure \ref{fig01}: urbanized areas in the great plains and Midwest have particularly high numbers of streets per node on average, indicating more grid-like, connected networks.
 Cities in the southern and western US tend to have fewer streets per node, reflecting more dead-ends and a disconnected network.
@@
 In Figure \ref{fig02}, urban Atlanta and Chattanooga have very high proportions of dead-ends, each over 30\% of all nodes, and very few 4-way intersections, indicating a disconnected street pattern.
-The urban areas of Phoenix, Boston, Detroit, and Chattanooga have particularly high proportions of 3-way intersections, each over 60\%, indicating a prevalence of T-intersections.
+The urbanized areas of Phoenix, Boston, Detroit, and Chattanooga have particularly high proportions of 3-way intersections, each over 60\%, indicating a prevalence of T-intersections.
 Conversely, Chicago, New Orleans, Duluth, and Lubbock have high proportions of 4-way intersections, indicating more grid-like connected networks.
@@
 Due to the substantial variation in urbanized area size, from 25 to 9,000 km\textsuperscript{2}, the preceding analysis covers a wide swath of metropolitan place types.
-To better compare apples-to-apples, Table \ref{table02} focuses on the 30 largest urban areas cross-sectionally to examine how their metric and topological measures compare.
+To better compare apples-to-apples, Table \ref{table02} focuses on the 30 largest urbanized areas cross-sectionally to examine how their metric and topological measures compare.
 This provides more consistent spatial scales and extents, while offering a window into the similarities and differences in the built forms of America's largest urban agglomerations.
 
-Among these urban areas, Milwaukee has the least circuitous network (6\% more circuitous than straight-line edges would be), and Orlando has the most (12\%).
+Among these urbanized areas, Milwaukee has the least circuitous network (6\% more circuitous than straight-line edges would be), and Orlando has the most (12\%).
 San Juan and Atlanta have the fewest streets per node on average (2.36 and 2.45, respectively), while Milwaukee has the most (3.03).

### 2017-03-28 latex/paper.tex
update citations

@@
 \usepackage{pdflscape}
+\usepackage[USenglish]{babel}
+\usepackage{csquotes}
 \usepackage[numbers]{natbib}
@@
 
-\bibliographystyle{apalike}
+\bibliographystyle{chicago}
 
@@
 
-OpenStreetMap is an emerging source of worldwide geospatial data useful to researchers.
-This paper uses the OSMnx software to automatically download and analyze 27,000 street networks from OpenStreetMap at metropolitan, municipal, and neighborhood scales---that is, every urbanized area, incorporated place, and Zillow neighborhood in the US.
+OpenStreetMap offers an underexplored source of worldwide geospatial data useful to urban researchers.
+This paper uses the OSMnx software to automatically download and analyze over 27,000 street networks from OpenStreetMap at metropolitan, municipal, and neighborhood scales---namely, every urbanized area, incorporated city and town, and Zillow neighborhood in the US.
 In turn it presents wide-ranging empirical findings on US urban form and street network characteristics.
-This paper addresses past limitations by conducting an analysis of street networks with an extremely large sample sizes, with clearly defined network definitions and extents for reproducibility, and using non-planar, directed graphs.
-Its analysis emphasizes urban street network measures relevant to urban design and morphology, including structure, connectedness, centrality, and robustness.
-In total we cross-sectionally analyze the street networks of 497 urbanized areas, 19,655 cities and towns, and 6,857 neighborhoods.
+This study conducts its analysis of with extremely large sample sizes, with clearly defined network definitions and extents for reproducibility, and using nonplanar, directed graphs.
+This analysis emphasizes urban street network measures relevant to urban design and morphology including structure, connectedness, centrality, and robustness.
+In total it cross-sectionally analyzes the street networks of 497 urbanized areas, 19,655 cities and towns, and 6,857 neighborhoods.
 These sample sizes are larger than those in any previous study and illustrate the use of OSMnx and OpenStreetMap as a new urban form research platform.
@@
 
-On May 20, 1862, Abraham Lincoln signed the Homestead Act into law, making land across the United States Midwest and Great Plains available for free to all applicants \cite{porterfield_homestead_2005}.
+On May 20, 1862, Abraham Lincoln signed the Homestead Act into law, making land across the United States' Midwest and Great Plains available for free to all applicants \cite{porterfield_homestead_2005}.
 Under its auspices over the next 70 years, the federal government distributed 270 million acres of public land (10\% of the entire US landmass) to private owners in the form of 1.6 million homesteads \cite{lee_kansas_1979, sherraden_inclusion_2005}.
@@
 The empirical literature on street network analysis is growing ever richer, but suffers from some limitations---discussed in detail in \cite{boeing_osmnx:_2017} and summarized here.
-First, sample sizes have tended to be fairly small due to data availability, gathering, and processing constraints.
-Second, reproducibility has been difficult when the dozens of decisions that go into analysis---such as spatial extents, topological simplification and correction, definitions of nodes and edges, etc.---are ad hoc or only partly reported.
-Third, and related to the first two, studies frequently oversimplify to planar or undirected primal graphs for tractability, or use dual graphs despite the loss of geographic and metric information.
+First, sample sizes have tended to be fairly small due to data availability, gathering, and processing constraints: most studies in the street network literature that conduct topological and/or metric analysis tend to have sample sizes ranging around 10 to 50 networks \cite{buhl_topological_2006, cardillo_structural_2006, barthelemy_modeling_2008, strano_urban_2013}, which may limit the generalizability and interpretability of findings.
+Second, reproducibility has been difficult when the dozens of decisions that go into analysis---such as spatial extents, topological simplification and correction, definitions of nodes and edges, etc.---are ad hoc or only partly reported \citep[e.g.,][]{porta_network_2006, strano_urban_2013}.
+Third, and related to the first two, studies frequently oversimplify to planar or undirected primal graphs for tractability \cite[e.g.,][]{buhl_topological_2006, cardillo_structural_2006, barthelemy_modeling_2008, masucci_random_2009}, or use dual graphs despite the loss of geographic and metric information \cite{batty_network_2005, jiang_integration_2002, ratti_space_2004, crucitti_centrality_2006, crucitti_centrality_2006-1}.
 
-This paper addresses these limitations by conducting an urban morphological analysis of street networks at multiple scales, with large sample sizes, with clearly defined network definitions and extents for reproducibility, and using non-planar, directed graphs.
-In particular, it examines urban street network---represented here as primal, non-planar, weighted multidigraphs with possible self-loops---focusing on structure, connectedness, centrality, and resilience.
-Most studies in the street network literature that conduct topological and/or metric analysis tend to have sample sizes ranging around 10 to 50 networks.
-This study instead analyzes over 27,000 urban street networks at multiple overlapping scales across the United States.
-Namely, it examines the street networks of every incorporated city and town (n=19,655), census urbanized area (n=497), and Zillow-defined neighborhood (n=6,857) in the US.
-To do so, it uses OSMnx \cite{boeing_osmnx:_2017}---a new street network research toolkit---to download, construct, and analyze these street networks at metropolitan, municipal, and neighborhood scales.
+This study addresses these limitations by conducting an urban morphological analysis of street networks at multiple scales, with large sample sizes, with clearly defined network definitions and extents for reproducibility, and using nonplanar, directed graphs.
+In particular, it examines urban street networks---represented here as primal, nonplanar, weighted multidigraphs with possible self-loops---focusing on structure, connectedness, centrality, and resilience.
+This study analyzes over 27,000 urban street networks at multiple overlapping scales across the US.
+Namely, it examines the street networks of every incorporated city and town (population size $N=19655$), census urbanized area ($N=497$), and Zillow-defined neighborhood ($N=6857$) in the US.
+To do so, it uses OSMnx---a new street network research toolkit \cite{boeing_osmnx:_2017}---to download, construct, and analyze these street networks at metropolitan, municipal, and neighborhood scales.
 
 The purpose of this paper is threefold.
-First, it demonstrates a new methodology for easily and consistently acquiring, constructing, and analyzing large samples of street networks as nonplanar directed graphs.
+First, it describes and demonstrates a new methodology for easily and consistently acquiring, constructing, and analyzing large samples of street networks as nonplanar directed graphs.
 Second, it presents empirical findings of descriptive urban morphology for the street networks of every city, urbanized area, and Zillow neighborhood in the US.
@@
 
-\section{Methodology}
+\section{Data and Methods}
 
@@
 A directed network has directed edges: that is, edge $uv$ points one-way from node $u$ to node $v$, but there need not exist a reciprocal edge $vu$.
-A planar network can be represented in two dimensions with its edges intersecting only at nodes.
-Most street networks are nonplanar---due to grade-separated expressways, overpasses, bridges, and tunnels---but most quantitative studies of urban street networks represent them as planar (e.g., Buhl et al.
-2006; Cardillo et al.
-2006; Barthélemy and Flammini 2008; Masucci et al.
-2009; Strano et al.
-2013) for tractability because bridges and tunnels are reasonably uncommon in some cities.
-Planarity may reasonably model the street networks of old European town centers, but poorly models the street network in more modern cities like Los Angeles with their grade-separated expressways, bridges, and tunnels.
-Thus, a planar simplification could yield inaccurate analyses.
+A planar network can be represented in two dimensions with its edges intersecting only at nodes \cite{viana_simplicity_2013, fischer_spatial_2014}.
+Most street networks are nonplanar---due to grade-separated expressways, overpasses, bridges, tunnels, etc.---but most quantitative studies of urban street networks represent them as planar \cite{buhl_topological_2006, cardillo_structural_2006, barthelemy_modeling_2008, masucci_random_2009, strano_urban_2013} for tractability because bridges and tunnels are reasonably uncommon in some cities.
+Planar graphs may reasonably model the street networks of old European town centers, but poorly model the street networks in modern autocentric cities like Los Angeles or Tokyo with many grade-separated expressways, bridges, and underpasses.
 
-This study uses OSMnx to download, construct, correct, analyze, and visualize street networks at metropolitan, municipal, and neighborhood scales \cite{boeing_osmnx:_2017}.
-OSMnx is a Python-based research tool that easily downloads OpenStreetMap data for any place in the world and constructs it into a spatially-embedded graph-theoretic object for analysis and visualization.
-OpenStreetMap is a collaborative worldwide mapping project that makes its spatial data available via various APIs.
-These data are of high quality and compare well to CIA World Factbook estimates and US Census TIGER/Line data (Haklay 2010; Over et al.
-2010; Zielstra and Hochmair 2011; Maron 2015).
-In 2007, OpenStreetMap imported the 2005 TIGER/Line roads data, and since then many community-led corrections and improvements have been made (Willis 2008).
-Many of these additions go beyond TIGER/Line's scope, including passageways between buildings, footpaths through parks, bike paths, and detailed feature attributes, such as finer-grained street classifiers, speed limits, etc.
+\subsection{Study Sites and Data Acquisition}
+
+This study uses OSMnx to download, construct, correct, analyze, and visualize street networks at metropolitan, municipal, and neighborhood scales.
+OSMnx is a Python-based research tool that easily downloads OpenStreetMap data for any place in the world and constructs it into a spatially-embedded graph-theoretic object for analysis and visualization \cite{boeing_osmnx:_2017}.
+OpenStreetMap is a collaborative worldwide mapping project that makes its spatial data available via various APIs \cite{corcoran_analysing_2013, jokar_arsanjani_openstreetmap_2015}.
+Its data are of high quality and compare favorably to CIA World Factbook estimates and US Census TIGER/Line data \cite{haklay_how_2010, over_generating_2010, zielstra_comparative_2011, maron_how_2015, wu_improving_2005, frizzelle_importance_2009}.
+In 2007, OpenStreetMap imported the TIGER/Line roads (2005 vintage), and since then many community-led corrections and improvements have been made \cite{willis_openstreetmap_2008}.
+Many of these additions go beyond TIGER/Line's scope, including passageways between buildings, footpaths through parks, bike routes, and detailed feature attributes such as finer-grained street classifiers, speed limits, etc.
 
 To define the study sites and their spatial boundaries, this study uses three sets of geometries.
-The first defines the metropolitan-scale study sites using the 2016 national TIGER/Line shapefile of US census bureau urban areas, released 19 August 2016.
-Census-defined urban areas comprise a set of census tracts that meet a minimum density threshold \cite{u.s._census_bureau_2010_2010}.
-In particular, we retain only the urbanized areas subset in this data set (i.e., areas with greater than 50,000 population), discarding the smaller urban clusters subset.
-The second set of geometries defines the municipal-scale study sites using 51 TIGER/Line shapefiles (again, 2016) of US census bureau places for all 50 states plus the District of Columbia.
-In particular, we discard the subset of census-designated places in this data set, instead retaining all the incorporated cities and towns in the United States.
+The first defines the metropolitan-scale study sites using the 2016 national TIGER/Line shapefile of US census bureau urban areas.
+Each census-defined urban areas comprises a set of tracts that meet a minimum density threshold \cite{u.s._census_bureau_2010_2010}.
+We retain only the \emph{urbanized areas} subset of these data (i.e., areas with greater than 50,000 population), discarding the small \emph{urban clusters} subset.
+The second set of geometries defines our municipal-scale study sites using 51 separate TIGER/Line shapefiles (again, 2016) of US census bureau \emph{places} within all 50 states plus the District of Columbia.
+We discard the subset of \emph{census-designated places} (i.e., small unincorporated communities) in these data, while retaining every incorporated city and town in the US.
+The third set of geometries defines the neighborhood-scale study sites using 42 separate Zillow neighborhood shapefiles.
+These shapefiles contain neighborhood boundaries for major cities in 41 states plus the District of Columbia.
+This is a fairly new data set comprising nearly 7,000 neighborhoods in large American cities, but as \citet{schernthanner_spatial_2016} point out, Zillow does not publish the methodology it uses to construct these boundaries.
+However, despite its newness it already has a track record in the academic literature: \citet{besbris_effect_2015} used Zillow boundaries to examine neighborhood stigma and \citet{albrecht_indicator_2014} used them to study neighborhood-level poverty in New York.
 
-The third and final set of geometries defines the neighborhood-scale study sites using 42 Zillow neighborhood shapefiles.
-These shapefiles contain the geometries of neighborhoods in 41 states plus the District of Columbia.
-This is a fairly new data set comprising nearly 7,000 neighborhoods in large American cities, but as \citet{schernthanner_spatial_2016} point out, Zillow does not publish the methodology it uses to construct these geometries.
-However, despite its newness it already has a track record in the academic literature.
-For instance, \citet{besbris_effect_2015} used Zillow boundaries to examine neighborhood stigma, and \citet{albrecht_indicator_2014} used these data to study neighborhood-level poverty in New York.
+For all of these geometries, we use OSMnx to download the (drivable, public) street network contained within the geometry, a process described in detail in \cite{boeing_osmnx:_2017} and summarized here.
+First OSMnx buffers each geometry by 0.5 km, then downloads the OpenStreetMap \enquote{nodes} and \enquote{ways} within this buffered geometry.
+Next it constructs a street network from these data, corrects the topology, calculates street counts per node, then truncates the network to the original, desired polygon (ensuring that intersections are not considered cul-de-sacs simply because an incident edge connects to a node outside the desired polygon).
+OSMnx saves each of these networks to disk as GraphML and shapefiles.
+Finally, it calculates metric and topological measures for each network.
+Such measures extend the toolkit commonly used in urban form studies \cite{talen_measuring_2003, ewing_travel_2010}.
 
-For all of these geometries, we use OSMnx to download the (drivable, public) street network contained within the geometry.
-First it buffers each geometry by 0.5 km, then downloads the street network within this buffered geometry.
-Next it constructs a street network from this data, corrects the topology, calculates street counts per node (this ensures that intersections are not considered cul-de-sacs simply because an incident edge connects to a node outside the desired polygon), then truncates the network to the original, desired polygon.
-OSMnx saves each of these networks to disk as GraphML and shapefiles.
-Finally, it automatically calculates metric and topological measures for each.
-Such measures extend the toolkit commonly used by urban morphologists (Talen 2003).
-Brief descriptions of these measures are discussed here, but extended definitions and algorithms can be found in Newman (2010) and Barthélemy (2011).
+\subsection{Street Network Measures}
 
+Brief descriptions of these measures are discussed here, but extended technical definitions and algorithms can be found in e.g.
+\cite{trudeau_introduction_1994, albert_statistical_2002, dorogovtsev_evolution_2002, brandes_network_2005, costa_characterization_2007, newman_structure_2003, newman_networks:_2010, barthelemy_spatial_2011, cranmer_navigating_2017}.
 Average street length, the mean edge length in the undirected representation of the graph, serves as a linear proxy for block size and indicates how fine-grained or coarse-grained the network is.
@@
 Connectivity measures the minimum number of nodes or edges that must be removed from a connected graph to disconnect it.
-The average node connectivity of a network---the mean number of internally node-disjoint paths between each pair of nodes in the graph---represents the expected number of nodes that must be removed to disconnect a randomly selected pair of non-adjacent nodes \cite{beineke_average_2002}.
+The average node connectivity of a network---the mean number of internally node-disjoint paths between each pair of nodes in the graph---represents the expected number of nodes that must be removed to disconnect a randomly selected pair of non-adjacent nodes \cite{beineke_average_2002,dankelmann_bounds_2003}.
 
-The clustering coefficient of a node is the ratio of the number of edges between its neighbors to the maximum possible number of edges that could exist between these neighbors.
+The clustering coefficient of a node is the ratio of the number of edges between its neighbors to the maximum possible number of edges that could exist between these neighbors \cite{jiang_topological_2004, opsahl_clustering_2009}.
 The weighted clustering coefficient weights this ratio by edge length and the average clustering coefficient is the mean of the clustering coefficients of all the nodes in the network.
-Centrality indicates the importance of nodes in a network.
+Centrality indicates the importance of nodes in a network \cite{huang_trajgraph:_2016, zhong_revealing_2017}.
 Betweenness centrality evaluates the number of shortest paths that pass through each node \cite{barthelemy_betweenness_2004}.
@@
 Closeness centrality ranks nodes as more central if they are on average closer to all other nodes.
-Finally, PageRank ranks nodes based on the structure of incoming links and the rank of the source node.
+Finally, PageRank ranks nodes based on the structure of incoming links and the rank of the source node \cite{brin_anatomy_1998, jiang_predicting_2008, agryzkov_algorithm_2012, chin_geographically_2015, gleich_pagerank_2015}.
 
@@
 
-\section{Analysis of Metropolitan-Scale Street Networks}
+\section{Results}
+
+\subsection{Analysis of Metropolitan-Scale Street Networks}
 
@@
 	\midrule
-	Area (km2)                                  & 460.657        & 858.125        & 25.685         & 184.898        & 8937.429       \\
+	Area (km\textsuperscript{2})                                  & 460.657        & 858.125        & 25.685         & 184.898        & 8937.429       \\
 	Avg of the avg neighborhood degree          & 2.886          & 0.109          & 2.626          & 2.875          & 3.228          \\
@@
 	Avg degree centrality                       & 0.001          & 0.001          & \textless0.001 & 0.001          & 0.007          \\
-	Edge density (km/km2)                       & 13.455         & 2.137          & 7.961          & 13.352         & 21.233         \\
+	Edge density (km/km\textsuperscript{2})                       & 13.455         & 2.137          & 7.961          & 13.352         & 21.233         \\
 	Avg edge length (m)                         & 158.588        & 17.653         & 117.341        & 157.332        & 223.08         \\
@@
 	Proportion of 4-way intersections           & 0.187          & 0.063          & 0.054          & 0.178          & 0.422          \\
-	Intersection density (per km2)              & 26.469         & 6.256          & 12.469         & 26.029         & 49.423         \\
+	Intersection density (per km\textsuperscript{2})              & 26.469         & 6.256          & 12.469         & 26.029         & 49.423         \\
 	Average node degree                         & 5.153          & 0.302          & 4.307          & 5.143          & 6.056          \\
@@
 	n                                           & 16032          & 32585          & 874            & 5830           & 373309         \\
-	Node density (per km2)                      & 33.628         & 7.641          & 17.675         & 33.071         & 61.655         \\
+	Node density (per km\textsuperscript{2})                      & 33.628         & 7.641          & 17.675         & 33.071         & 61.655         \\
 	Max PageRank value                          & 0.001          & 0.001          & \textless0.001 & 0.001          & 0.003          \\
@@
 	Self-loop proportion                        & 0.008          & 0.008          & \textless0.001 & 0.006          & 0.071          \\
-	Street density (km/km2)                     & 7.262          & 1.221          & 4.217          & 7.171          & 11.797         \\
+	Street density (km/km\textsuperscript{2})                     & 7.262          & 1.221          & 4.217          & 7.171          & 11.797         \\
 	Average street segment length (m)           & 161.331        & 17.765         & 119.573        & 160.288        & 225.92         \\
@@
 \hline
-urban area core city & land area km2 & avg circuity & avg clustering coefficient & dead-end ratio & 3-way ratio & 4-way ratio & intersect density/ km2 & street density km/km2 & avg street length (m) & avg streets/ node \\
+urban area core city & land area km\textsuperscript{2} & avg circuity & avg clustering coefficient & dead-end ratio & 3-way ratio & 4-way ratio & intersect density/ km\textsuperscript{2} & street density km/km\textsuperscript{2} & avg street length (m) & avg streets/ node \\
 \hline
@@
 However, they aggregate multiple heterogenous neighborhoods and municipalities---the scales of human life, urban design projects, and planning jurisdiction---into single units of analysis.
-To disaggregate and analyze finer characteristics, the following sections examine municipal- and neighborhood-scale street networks across the United States.
+To disaggregate and analyze finer characteristics, the following sections examine municipal- and neighborhood-scale street networks across the US.
 
-\section{Analysis of Municipal-Scale Street Networks}
+\subsection{Analysis of Municipal-Scale Street Networks}
 
 Table \ref{table03} shows similarly great variation in street network characteristics across the entire data set of 19,655 cities and towns.
-Again, this data set comprises the street networks of every incorporated city and town in the United States.
+Again, this data set comprises the street networks of every incorporated city and town in the US.
 Following recent work by \citet{barthelemy_modeling_2008} and \citet{strano_urban_2013}, we examine the relationship between the total street length $L$ and the number of nodes $n$.
@@
 	\midrule
-	Area (km2)                                  & 16.703  & 107.499 & 0.039          & 3.918          & 7434.258  \\
+	Area (km\textsuperscript{2})                                  & 16.703  & 107.499 & 0.039          & 3.918          & 7434.258  \\
 	Avg of the avg neighborhood degree          & 2.94    & 0.297   & 0.4            & 2.953          & 3.735     \\
@@
 	Avg degree centrality                       & 0.093   & 0.136   & \textless0.001 & 0.052          & 2.667     \\
-	Edge density (km/km2)                       & 12.654  & 6.705   & 0.006          & 11.814         & 58.603    \\
+	Edge density (km/km\textsuperscript{2})                       & 12.654  & 6.705   & 0.006          & 11.814         & 58.603    \\
 	Avg edge length (m)                         & 161.184 & 80.769  & 25.822         & 144.447        & 3036.957  \\
@@
 	Proportion of 4-way intersections           & 0.237   & 0.129   & \textless0.001 & 0.217          & 1         \\
-	Intersection density (per km2)              & 29.363  & 21.607  & \textless0.001 & 24.719         & 259.647   \\
+	Intersection density (per km\textsuperscript{2})              & 29.363  & 21.607  & \textless0.001 & 24.719         & 259.647   \\
 	Average node degree                         & 5.251   & 0.668   & 0.8            & 5.268          & 7.166     \\
@@
 	n                                           & 401     & 1516    & 2              & 103            & 71993     \\
-	Node density (per km2)                      & 35.449  & 24.409  & 0.047          & 30.718         & 296.74    \\
+	Node density (per km\textsuperscript{2})                      & 35.449  & 24.409  & 0.047          & 30.718         & 296.74    \\
 	Max PageRank value                          & 0.034   & 0.046   & \textless0.001 & 0.021          & 0.87      \\
@@
 	Self-loop proportion                        & 0.005   & 0.015   & \textless0.001 & \textless0.001 & 1         \\
-	Street density (km/km2)                     & 6.528   & 3.435   & 0.003          & 6.109          & 29.302    \\
+	Street density (km/km\textsuperscript{2})                     & 6.528   & 3.435   & 0.003          & 6.109          & 29.302    \\
 	Average street segment length (m)           & 162.408 & 81.035  & 25.822         & 145.479        & 3036.957  \\
@@
 The latter results from the fact that most towns across the Great Plains exhibit minimal suburban sprawl.
-Thus, the municipal boundaries snugly embrace the grid-like street network, without extending to accommodate the vast peripheral belt of twentieth century sprawl, circuity, and ``loops and lollipops" \cite{southworth_streets_1997} that characterizes cities in California that were settled in the same era but later subjected to substantial suburbanization.
+Thus, the municipal boundaries snugly embrace the grid-like street network, without extending to accommodate the vast peripheral belt of twentieth century sprawl, circuity, and \enquote{loops and lollipops} \cite{southworth_streets_1997} that characterizes cities in California that were settled in the same era but later subjected to substantial suburbanization.
 
@@
 	\toprule
-	State & Intersect density (per km2) & Avg streets per node & Avg circuity & Avg street segment length \\
+	State & Intersect density (per km\textsuperscript{2}) & Avg streets per node & Avg circuity & Avg street segment length \\
 	\midrule
@@
 This indicates the most grid-like networks.
-If we measure density and connectedness in terms of intersection density---previously identified as an emergent property of complex network organization---at the city-scale and then aggregate these cities by state, we find Rhode Island, Nebraska, New Jersey, Kansas, and Montana have, in order, the highest medians.
+If we measure density and connectedness in terms of intersection density at the city-scale and then aggregate these cities by state, we find Rhode Island, Nebraska, New Jersey, Kansas, and Montana have, in order, the highest medians.
 We again see three Great Plains states near the top, alongside densely populated East Coast states.
@@
 
-\section{Analysis of Neighborhood-Scale Street Networks}
+\subsection{Analysis of Neighborhood-Scale Street Networks}
 
-Thus far, we have examined every urban street network in the United States at the metropolitan and municipal scales.
+Thus far, we have examined every urban street network in the US at the metropolitan and municipal scales.
 This analysis has focused on the complexity on the network in terms of density, resilience, and connectedness.
 While the metropolitan scale captures the emergent character of the wider region's complex system, and the municipal scale captures planning decisions made by a single city government, the neighborhood scale best represents the scale of individual urban design interventions into the urban form.
-Further, this scale more commonly reflects individual designs, eras, and paradigms in street network development than the ``many hands, many eras" evolution of form at larger scales.
+Further, this scale more commonly reflects individual designs, eras, and paradigms in street network development than the \enquote{many hands, many eras} evolution of form at larger scales.
 
@@
 	\midrule
-	Area (km2)                                  & 5.322   & 15.463  & 0.008          & 1.738          & 323.306  \\
+	Area (km\textsuperscript{2})                                  & 5.322   & 15.463  & 0.008          & 1.738          & 323.306  \\
 	Avg of the avg neighborhood degree          & 2.598   & 0.436   & \textless0.001 & 2.67           & 3.632    \\
@@
 	Avg degree centrality                       & 0.13    & 0.27    & 0.001          & 0.054          & 4        \\
-	Edge density (km/km2)                       & 17.569  & 7.095   & 0.025          & 18.152         & 59.939   \\
+	Edge density (km/km\textsuperscript{2})                       & 17.569  & 7.095   & 0.025          & 18.152         & 59.939   \\
 	Avg edge length (m)                         & 142.279 & 59.182  & 8.447          & 133.848        & 2231.331 \\
@@
 	Proportion of 4-way intersections           & 0.275   & 0.176   & \textless0.001 & 0.234          & 1        \\
-	Intersection density (per km2)              & 49.497  & 28.33   & \textless0.001 & 46.43          & 444.355  \\
+	Intersection density (per km\textsuperscript{2})              & 49.497  & 28.33   & \textless0.001 & 46.43          & 444.355  \\
 	Average node degree                         & 4.675   & 0.836   & 0.545          & 4.736          & 7.283    \\
@@
 	n                                           & 208     & 459     & 2              & 90             & 9327     \\
-	Node density (per km2)                      & 58.677  & 31.802  & 0.063          & 55.626         & 499.9    \\
+	Node density (per km\textsuperscript{2})                      & 58.677  & 31.802  & 0.063          & 55.626         & 499.9    \\
 	Max PageRank value                          & 0.055   & 0.086   & \textless0.001 & 0.026          & 0.889    \\
@@
 	Self-loop proportion                        & 0.007   & 0.034   & \textless0.001 & \textless0.001 & 1        \\
-	Street density (km/km2)                     & 9.744   & 4.085   & 0.013          & 9.882          & 33.737   \\
+	Street density (km/km\textsuperscript{2})                     & 9.744   & 4.085   & 0.013          & 9.882          & 33.737   \\
 	Average street segment length (m)           & 143.664 & 60.023  & 7.376          & 134.877        & 2231.331 \\
@@
 	\caption{Square-mile comparisons of central cities and their suburbs.
-Left: top, downtown Philadelphia; bottom, its suburb, King of Prussia.
-Middle: top, downtown Portland; bottom, its suburb, Beaverton.
-Right: top, downtown San Francisco; bottom, its suburb, Concord.}
+Left: top, downtown Philadelphia, PA; bottom, its suburb, King of Prussia.
+Middle: top, downtown Portland, OR; bottom, its suburb, Beaverton.
+Right: top, downtown San Francisco, CA; bottom, its suburb, Concord.}
 	\label{fig05}
@@
 In fact, the suburbs have more in common with one another---despite being hundreds or thousands of miles apart---than they do with their central city neighbors, suggesting that land use and an era's prevailing design paradigm is paramount to geographical localism and regional context.
-The top row of Figure \ref{fig05} represents an era of urban planning and development that preceded the automobile, while the bottom row reflects the exclusionary zoning and mid-late twentieth century era of automobility in residential suburb design---namely the ``loops and lollipops" and the ``lollipops on a stick" design patterns identified by \citet{southworth_streets_1997}.
+The top row of Figure \ref{fig05} represents an era of urban planning and development that preceded the automobile, while the bottom row reflects the exclusionary zoning and mid-late twentieth century era of automobility in residential suburb design---namely the \enquote{loops and lollipops} and the \enquote{lollipops on a stick} design patterns identified by \citet{southworth_streets_1997}.
 
@@
 
-In total, this paper analyzed 497 urbanized areas' street networks, 19,655 cities' and towns' street networks, and 6,857 neighborhoods' street networks.
-These sample sizes were larger than those in any previous study.
-It looked at both metric and topological measures of the structure and complexity of these networks---particularly focusing on density, connectedness, and resilience.
-These preliminary empirical findings demonstrate the use of OSMnx as a new piece of research infrastructure.
-They suggest to urban planners new methods for acquiring and analyzing street network data, including new methods for evaluating network resilience and resilience gains with betweenness centralities and average node connectivities.
-Finally, this study has made all of these network datasets---for 497 urbanized areas, 19,655 cities and towns, and 6,857 neighborhoods---along with all of their attribute data and complexity measures available in an online public repository for other researchers to study and repurpose.
+\section{Conclusion}
+
+This paper had three primary purposes.
+First, it examined urban morphology through metric and topological measures of these networks---particularly focusing on density, connectedness, and resilience.
+Second, these methods and findings demonstrate the use of OSMnx as a new street network research toolkit.
+They suggest to urban planners and scholars new methods for acquiring and analyzing street network data.
+Third, it investigated claims about the distribution of street segment lengths and the relationship between the total street length and the number of nodes in a network.
+
+This study has made all of these network datasets---for 497 urbanized areas, 19,655 cities and towns, and 6,857 neighborhoods---along with all of their attribute data and morphological measures available in an online public repository for other researchers to study and repurpose at [redacted for peer review].
 

### 2017-03-28 latex/paper.tex
clean up discussion and conclusion

@@
 \usepackage[affil-it]{authblk}
+\usepackage{setspace}
+
+\doublespacing
 
@@
 A planar network can be represented in two dimensions with its edges intersecting only at nodes \cite{viana_simplicity_2013, fischer_spatial_2014}.
-Most street networks are nonplanar---due to grade-separated expressways, overpasses, bridges, tunnels, etc.---but most quantitative studies of urban street networks represent them as planar \cite{buhl_topological_2006, cardillo_structural_2006, barthelemy_modeling_2008, masucci_random_2009, strano_urban_2013} for tractability because bridges and tunnels are reasonably uncommon in some cities.
+Most street networks are nonplanar---due to grade-separated expressways, overpasses, bridges, tunnels, etc.---but most quantitative studies of urban street networks represent them as planar \cite{buhl_topological_2006, cardillo_structural_2006, barthelemy_modeling_2008, masucci_random_2009, strano_urban_2013} for tractability because bridges and tunnels are uncommon in some cities.
 Planar graphs may reasonably model the street networks of old European town centers, but poorly model the street networks in modern autocentric cities like Los Angeles or Tokyo with many grade-separated expressways, bridges, and underpasses.
@@
 Connectivity measures the minimum number of nodes or edges that must be removed from a connected graph to disconnect it.
-The average node connectivity of a network---the mean number of internally node-disjoint paths between each pair of nodes in the graph---represents the expected number of nodes that must be removed to disconnect a randomly selected pair of non-adjacent nodes \cite{beineke_average_2002,dankelmann_bounds_2003}.
+The average node connectivity (ANC) of a network---the mean number of internally node-disjoint paths between each pair of nodes in the graph---represents the expected number of nodes that must be removed to disconnect a randomly selected pair of non-adjacent nodes \cite{beineke_average_2002,dankelmann_bounds_2003}.
 
@@
 Betweenness centrality evaluates the number of shortest paths that pass through each node \cite{barthelemy_betweenness_2004}.
-The maximum betweenness centrality in a network specifies the proportion of shortest paths that pass through the most important node.
-This is an indicator of resilience: networks with a high maximum betweenness centrality are more prone to failure or inefficiency should this single choke point fail.
+The maximum betweenness centrality (MBC) in a network specifies the proportion of shortest paths that pass through the most important node.
+This is an indicator of resilience: networks with a high MBC are more prone to failure or inefficiency should this single choke point fail.
 Closeness centrality ranks nodes as more central if they are on average closer to all other nodes.
@@
 Moreover, previous findings \citep[e.g.,][]{masucci_random_2009} suggest the distribution of street segment lengths in an urban street network follows a power-law distribution.
-However, we find that these networks instead generally follow a lognormal distribution, which makes theoretical sense as most street networks are not truly scale-free.
-For example, a typical street network might comprise very few very long street segments (e.g., 1 km), more medium-length segments (e.g., 250 m), and many short segments (e.g., 80 m), but very few very short segments (e.g., 10 m).
+However, we find that these networks instead generally follow a lognormal distribution.
+This makes theoretical sense as most street networks are not truly scale-free: for example, a typical street network might comprise very few very long street segments (e.g., 1 km), more medium-length segments (e.g., 250 m), many short segments (e.g., 80 m), but very few very short segments (e.g., 10 m).
 This theoretical illustration suggests the lognormal distribution that this study generally finds across municipal US street networks.
@@
 
-This study finds that such cities are not uncommon, particularly between the Mississippi River and the Rocky Mountains: the Great Plains states are characterized by a unique street network form that is both orthogonal and reasonably dense.
+This study finds that such cities are not uncommon in the US, particularly between the Mississippi River and the Rocky Mountains: the Great Plains states are characterized by a unique street network form that is both orthogonal and reasonably dense.
 The former is partly the result of topography (flat, plains terrain that allows idealized grids) and design history (platting and development during the late nineteenth century) that favor orthogonal grids, as discussed earlier.
@@
 
+Finally we briefly take a closer look at San Francisco, California's neighborhoods alone for a clear cross-sectional analysis with consistent geography to look across planning eras and design paradigms.
+Due to its highly connected orthogonal grid, the Outer Sunset neighborhood has the lowest MBC---only 9.6\% of all shortest paths pass through its most important node.
+By contrast, in Chinatown, 36\% of shortest paths pass through the most important node, and in Twin Peaks it is 37\%.
+In Chinatown, this is the result of the small neighborhood comprising only a few streets and that these streets are one-way, forcing paths through very few routing options.
+In Twin Peaks, this is the result of hilly terrain and a disconnected network forcing paths through a small set of chokepoints that link separate subsections of the network.
+If a large number of shortest paths rely on a single node, the network is more prone to failure or inefficiency given a single point of failure.
+
+Average node connectivity is another proxy for resilience.
+In San Francisco, Twin Peaks' network has the lowest ANC: on average only 1.05 nodes must be removed to disconnect a randomly selected pair of nodes.
+Outer Sunset has the highest ANC, 3.2.
+These concur with the above descriptions of these networks.
+However, some central-city orthogonal grid networks with many 4-way intersections---such as Downtown, Chinatown, and the Financial District---have surprisingly low ANC: 1.5, 1.3, and 1.6 respectively.
+These neighborhoods comprise primarily one-way streets.
+Although they have dense, highly connected networks, they can be easily disconnected given that traffic cannot flow bidirectionally.
+These three neighborhoods also experience the greatest increase in ANC if all the edges are made undirected: Chinatown's increases 87\%, Downtown's 80\%, and the Financial District's 75\%.
+By contrast, the street network of Outer Sunset sees only a 6\% increase due to it already comprising primarily bidirectional streets.
+Targeted conversion of one-way streets to bidirectional in Downtown, the Financial District, and Chinatown may yield substantial network resilience gains.
+
 \section{Discussion}
 
-A complex spatial network is a network, embedded in space, that has a nontrivial topology.
-In other words, its structure and organization is neither fully regular nor fully random.
-Of the network measures, this empirical study has presented preliminary findings that emphasize street network complexity in terms of density, resilience, and connectedness.
-Older, denser, and more self-organized networks, such as those at the heart of Boston or lower Manhattan are complex in terms of category I complexity.
-Sprawling, disconnected suburban neighborhoods rank low on all measures of complexity, with the exception that their high circuity can lend itself to disorder (i.e., category I).
-The orthogonal grid we see in the downtowns of Portland and San Francisco have high density (i.e., intersection and street densities), connectedness (i.e., average number of streets per node), and order (based on circuity and statistical dispersion of node types), but low resilience in the presence of one-way streets, measured by maximum betweenness centralities and average node connectivity increases when switching from one-way to bi-directional edges.
-These latter neighborhoods are complex in terms of category II and category III complexity.
-Their connectedness, density, and orderliness balances at the midpoint embraced by category II.
-However, their extreme regularity of block sizes, streets per node, and orthogonality best represents category III.
+These findings suggest the influence of planning eras, design paradigms, transportation technologies, topography, and economics on US street network density, resilience, and connectedness.
+Sprawling suburban neighborhoods rank low on density, connectedness, and order.
+The orthogonal grids we see in the downtowns of Portland and San Francisco have high density (i.e., intersection and street densities), connectedness (i.e., average number of streets per node), and order (based on circuity and statistical dispersion of node types), but low resilience in the presence of one-way streets, measured by MBC and ANC increases when switching from one-way to bidirectional edges.
 
-Another critical takeaway of this analysis is that scale matters.
+A critical takeaway is that \emph{scale} matters.
 The median average circuity is lower across the neighborhoods data set than across the municipal set, which in turn is lower than across the urbanized areas set.
@@
 
-The characteristics of an urban network for a city fundamentally depend on what city means: municipal boundaries, urbanized areas, or just certain neighborhoods?
-The first is a merely legal definition, but also captures the scope of city planning authority and decision-making for top-down interventions into a street network.
-The second captures a wider self-organized human system and its emergent built form, but tends to aggregate multiple non-homogenous built forms together into a single unit of analysis.
-The third captures the nature of the local built form and lived experience, but at the expense of a broad view of the urban system and metropolitan-scale trip-taking.
+The characteristics of city street network fundamentally depend on what \emph{city} means: municipal boundaries, urbanized areas, or certain central neighborhoods?
+The first is a legal/political definition, but captures the scope of city planning authority and decision-making for top-down interventions into human circulation.
+The second captures a wider self-organized human system and its emergent built form, but tends to aggregate multiple heterogeneous built forms together into a single unit of analysis.
+The third captures the nature of the local built form and lived experience, but at the expense of a broader view of the urban system and metropolitan-scale trip-taking.
 In short, multiple scales in concert provide planners and scholars a clearer view of the urban form and the topological and metric complexity of the street network than any single scale can.
 
-We find a strong linear relationship, invariant across scales, between total street length, L, and the number of nodes, n, in a network that contradicts some previous findings in the literature that relied on small sample sizes.
-We also find that most networks empirically demonstrate a lognormal distribution of street segment lengths, contradicting some previous findings in the literature, as discussed earlier.
-However, we believe our empirical finding makes more sense theoretically and is supported by the large-sample data at multiple scales.
-An obvious exception to lognormal distribution lies in those networks that exhibit substantial uniformity across the entire network.
-At the neighborhood scale, examples include downtown neighborhoods with similarly consistent orthogonal grids, such as that of Portland, Oregon.
-At the municipal scale, examples include towns in the Great Plains that have orthogonal grids with consistent block sizes, platted at one time, and never subjected to sprawl.
+This analysis finds a strong linear relationship, invariant across scales, between total street length and the number of nodes in a network.
+This differs from previous findings in the literature that relied on smaller sample sizes and examined European instead of US cities.
+We also find that most networks typically follow a lognormal distribution of street segment lengths.
+As discussed, this finding seems to make sense theoretically and is supported by these large-sample data at multiple scales, but obvious exceptions to lognormal distribution lie in those networks that exhibit substantial uniformity.
+At the neighborhood scale, examples include downtowns with consistent orthogonal grids, such as that of Portland, Oregon.
+At the municipal scale, examples include towns in the Great Plains that have orthogonal grids with consistent block sizes, platted at one time, and never subjected to expansion or sprawl.
 
-These findings are interesting for the practice and history of planning.
-The spatial signatures of the Homestead Act, successive land use regulations, urban design paradigms, and planning instruments remain clearly visible today in these cities' urban forms and street networks due to path dependence.
-When comparing the median municipal street networks of each state, Nebraska has the lowest circuity, the highest average number of streets per node, the second shortest average street segment length, and the second highest intersection density for similar reasons.
-These preliminary findings point to how street networks across the Great Plains developed all at once, but grew very little afterwards---unlike, for instance, most cities in California that were settled in the same era but later subjected to sprawl.
+These findings reveal urban form legacies of the practice and history of US planning.
+The spatial signatures of the Homestead Act, successive land use regulations, urban design paradigms, and planning instruments remain clearly etched in these cities' urban forms and street networks today.
+Accordingly when comparing median municipal street networks in each state, Nebraska has the lowest circuity, the highest average number of streets per node, the second shortest average street segment length, and the second highest intersection density.
+These findings illustrate how street networks across the Great Plains developed all at once, but grew very little afterwards---unlike, for instance, most cities in California that were settled in the same era but later subjected to sprawl.
 
-This finding suggests future research should incorporate a temporal analysis that this present study does not do with its cross-sectional data.
-This empirical analysis emphasized network structure.
-Expanding the study of complex urban form by examining the other types of complexity---and further linking structural complexity to the temporal complexity of dynamics and processes---lies ahead as critical future work.
+Future research should incorporate temporal analyses that goes beyond the present study's cross-sectional data.
+This empirical analysis emphasized network structure, but further linking structural complexity to the temporal complexity of city dynamics and processes lies ahead as critical work.
+As OSMnx can automatically calculate several dozen street network measures, future work can use dimensionality reduction to identify significant baskets of factors and cluster places into types.
+These variables can also be used as advanced urban form measures in hedonic regressions and accessibility studies.
 
@@
 Third, it investigated claims about the distribution of street segment lengths and the relationship between the total street length and the number of nodes in a network.
-
 This study has made all of these network datasets---for 497 urbanized areas, 19,655 cities and towns, and 6,857 neighborhoods---along with all of their attribute data and morphological measures available in an online public repository for other researchers to study and repurpose at [redacted for peer review].

### 2017-03-29 latex/paper.tex
text edits

@@
 \usepackage{pdflscape}
+\usepackage{setspace}
 \usepackage[USenglish]{babel}
@@
 \usepackage[affil-it]{authblk}
-\usepackage{setspace}
 
 \doublespacing
-
 \bibliographystyle{chicago}
@@
 OpenStreetMap offers an underexplored source of worldwide geospatial data useful to urban researchers.
-This paper uses the OSMnx software to automatically download and analyze over 27,000 street networks from OpenStreetMap at metropolitan, municipal, and neighborhood scales---namely, every urbanized area, incorporated city and town, and Zillow neighborhood in the US.
-In turn it presents wide-ranging empirical findings on US urban form and street network characteristics.
-This study conducts its analysis of with extremely large sample sizes, with clearly defined network definitions and extents for reproducibility, and using nonplanar, directed graphs.
-This analysis emphasizes urban street network measures relevant to urban design and morphology including structure, connectedness, centrality, and robustness.
-In total it cross-sectionally analyzes the street networks of 497 urbanized areas, 19,655 cities and towns, and 6,857 neighborhoods.
-These sample sizes are larger than those in any previous study and illustrate the use of OSMnx and OpenStreetMap as a new urban form research platform.
+This study uses the OSMnx software to automatically download and analyze over 27,000 street networks from OpenStreetMap at metropolitan, municipal, and neighborhood scales---namely, every US city and town ($N=19655$), census urbanized area ($N=497$), and Zillow-defined neighborhood ($N=6857$).
+It presents wide-ranging empirical findings on US urban form and street network characteristics, emphasizing measures relevant to urban design and morphology such as structure, connectedness, density, centrality, and resilience.
+In the past, street network data acquisition has been challenging and ad hoc, but this study illustrates the use of OSMnx and OpenStreetMap to conduct street network analysis with extremely large sample sizes, with clearly defined network definitions and extents for reproducibility, and using nonplanar, directed graphs.
+These street networks and their measures are now shared in a public repository for other researchers to use.
 
@@
 
-On May 20, 1862, Abraham Lincoln signed the Homestead Act into law, making land across the United States' Midwest and Great Plains available for free to all applicants \cite{porterfield_homestead_2005}.
-Under its auspices over the next 70 years, the federal government distributed 270 million acres of public land (10\% of the entire US landmass) to private owners in the form of 1.6 million homesteads \cite{lee_kansas_1979, sherraden_inclusion_2005}.
-Towns with gridiron street networks sprang up rapidly across the Great Plains and Midwest, due to both the prevailing urban design paradigm of the day and the standardized rectilinear town plats used repeatedly to lay out instant new cities \cite{southworth_streets_1997}.
+On May 20, 1862, Abraham Lincoln signed the Homestead Act into law, making land across the United States' Midwest and Great Plains available for free to applicants \cite{porterfield_homestead_2005}.
+Under its auspices over the next 70 years, the federal government distributed 10\% of the entire US landmass to private owners in the form of 1.6 million homesteads \cite{lee_kansas_1979, sherraden_inclusion_2005}.
+New towns with gridiron street networks sprang up rapidly across the Great Plains and Midwest, due to both the prevailing urban design paradigm of the day and the standardized rectilinear town plats used repeatedly to lay out instant new cities \cite{southworth_streets_1997}.
 Through path dependence, the spatial signatures of these land use laws, design paradigms, and planning instruments can still be seen today in these cities' urban forms and street networks.
@@
 Network analysis is a natural approach to the study of cities as complex systems \cite{masucci_random_2009}.
-The empirical literature on street network analysis is growing ever richer, but suffers from some limitations---discussed in detail in \cite{boeing_osmnx:_2017} and summarized here.
-First, sample sizes have tended to be fairly small due to data availability, gathering, and processing constraints: most studies in the street network literature that conduct topological and/or metric analysis tend to have sample sizes ranging around 10 to 50 networks \cite{buhl_topological_2006, cardillo_structural_2006, barthelemy_modeling_2008, strano_urban_2013}, which may limit the generalizability and interpretability of findings.
+The empirical literature on street networks is growing ever richer, but suffers from some limitations---discussed in detail in \cite{boeing_osmnx:_2017} and summarized here.
+First, sample sizes tend to be fairly small due to data availability, gathering, and processing constraints: most studies in this literature that conduct topological or metric analysis tend to have sample sizes ranging around 10 to 50 networks \cite{buhl_topological_2006, cardillo_structural_2006, barthelemy_modeling_2008, strano_urban_2013}, which may limit the generalizability and interpretability of findings.
 Second, reproducibility has been difficult when the dozens of decisions that go into analysis---such as spatial extents, topological simplification and correction, definitions of nodes and edges, etc.---are ad hoc or only partly reported \citep[e.g.,][]{porta_network_2006, strano_urban_2013}.
@@
 This study addresses these limitations by conducting an urban morphological analysis of street networks at multiple scales, with large sample sizes, with clearly defined network definitions and extents for reproducibility, and using nonplanar, directed graphs.
-In particular, it examines urban street networks---represented here as primal, nonplanar, weighted multidigraphs with possible self-loops---focusing on structure, connectedness, centrality, and resilience.
-This study analyzes over 27,000 urban street networks at multiple overlapping scales across the US.
+In particular, it examines urban street networks---represented as primal, nonplanar, weighted multidigraphs with possible self-loops---focusing on structure, connectedness, centrality, and resilience.
+This study analyzes 27,009 urban street networks at multiple overlapping scales across the US.
 Namely, it examines the street networks of every incorporated city and town (population size $N=19655$), census urbanized area ($N=497$), and Zillow-defined neighborhood ($N=6857$) in the US.
 To do so, it uses OSMnx---a new street network research toolkit \cite{boeing_osmnx:_2017}---to download, construct, and analyze these street networks at metropolitan, municipal, and neighborhood scales.
+These street networks and their measures data have been shared in a public repository for other researchers to use at [link redacted].
 
@@
 First, it describes and demonstrates a new methodology for easily and consistently acquiring, constructing, and analyzing large samples of street networks as nonplanar directed graphs.
-Second, it presents empirical findings of descriptive urban morphology for the street networks of every city, urbanized area, and Zillow neighborhood in the US.
+Second, it presents empirical findings of descriptive urban morphology for the street networks of every US city, urbanized area, and Zillow neighborhood.
 Third, it investigates with large sample sizes some claims previously made in the research literature with small sample sizes.
@@
 In the next section, it discusses the data sources, tools, and methods used to collect, construct, and analyze these street networks.
-Next, it presents findings of the analyses at metropolitan, municipal, and neighborhood spatial scales.
+Next, it presents findings of the analyses at metropolitan, municipal, and neighborhood scales.
 Finally, it concludes with a discussion of these findings and their implications for street network analysis, urban morphology, and city planning.
 
-\section{Data and Methods}
+\section{Methodology}
 
@@
 A directed network has directed edges: that is, edge $uv$ points one-way from node $u$ to node $v$, but there need not exist a reciprocal edge $vu$.
-A planar network can be represented in two dimensions with its edges intersecting only at nodes \cite{viana_simplicity_2013, fischer_spatial_2014}.
-Most street networks are nonplanar---due to grade-separated expressways, overpasses, bridges, tunnels, etc.---but most quantitative studies of urban street networks represent them as planar \cite{buhl_topological_2006, cardillo_structural_2006, barthelemy_modeling_2008, masucci_random_2009, strano_urban_2013} for tractability because bridges and tunnels are uncommon in some cities.
-Planar graphs may reasonably model the street networks of old European town centers, but poorly model the street networks in modern autocentric cities like Los Angeles or Tokyo with many grade-separated expressways, bridges, and underpasses.
+A planar network can be represented in two dimensions with its edges intersecting only at nodes \cite[e.g.,][]{viana_simplicity_2013, fischer_spatial_2014}.
+Most street networks are nonplanar---due to grade-separated expressways, overpasses, bridges, tunnels, etc.---but most quantitative studies of urban street networks represent them as planar \cite[e.g.,][]{buhl_topological_2006, cardillo_structural_2006, barthelemy_modeling_2008, masucci_random_2009, strano_urban_2013} for tractability because bridges and tunnels are uncommon in some cities.
+Planar graphs may reasonably model the street networks of old European town centers, but poorly model the street networks of modern autocentric cities like Los Angeles or Tokyo with many grade-separated expressways, bridges, and underpasses.
 
@@
 OpenStreetMap is a collaborative worldwide mapping project that makes its spatial data available via various APIs \cite{corcoran_analysing_2013, jokar_arsanjani_openstreetmap_2015}.
-Its data are of high quality and compare favorably to CIA World Factbook estimates and US Census TIGER/Line data \cite{haklay_how_2010, over_generating_2010, zielstra_comparative_2011, maron_how_2015, wu_improving_2005, frizzelle_importance_2009}.
-In 2007, OpenStreetMap imported the TIGER/Line roads (2005 vintage), and since then many community-led corrections and improvements have been made \cite{willis_openstreetmap_2008}.
+These data are of high quality and compare favorably to CIA World Factbook estimates and US Census TIGER/Line data \cite{haklay_how_2010, over_generating_2010, zielstra_comparative_2011, maron_how_2015, wu_improving_2005, frizzelle_importance_2009}.
+In 2007, OpenStreetMap imported the TIGER/Line roads (2005 vintage) and since then, many community-led corrections and improvements have been made \cite{willis_openstreetmap_2008}.
 Many of these additions go beyond TIGER/Line's scope, including passageways between buildings, footpaths through parks, bike routes, and detailed feature attributes such as finer-grained street classifiers, speed limits, etc.
@@
 To define the study sites and their spatial boundaries, this study uses three sets of geometries.
-The first defines the metropolitan-scale study sites using the 2016 national TIGER/Line shapefile of US census bureau urban areas.
+The first defines the metropolitan-scale study sites using the 2016 TIGER/Line shapefile of US census bureau urban areas.
 Each census-defined urban areas comprises a set of tracts that meet a minimum density threshold \cite{u.s._census_bureau_2010_2010}.
 We retain only the \emph{urbanized areas} subset of these data (i.e., areas with greater than 50,000 population), discarding the small \emph{urban clusters} subset.
-The second set of geometries defines our municipal-scale study sites using 51 separate TIGER/Line shapefiles (again, 2016) of US census bureau \emph{places} within all 50 states plus the District of Columbia.
-We discard the subset of \emph{census-designated places} (i.e., small unincorporated communities) in these data, while retaining every incorporated city and town in the US.
-The third set of geometries defines the neighborhood-scale study sites using 42 separate Zillow neighborhood shapefiles.
-These shapefiles contain neighborhood boundaries for major cities in 41 states plus the District of Columbia.
-This is a fairly new data set comprising nearly 7,000 neighborhoods in large American cities, but as \citet{schernthanner_spatial_2016} point out, Zillow does not publish the methodology it uses to construct these boundaries.
-However, despite its newness it already has a track record in the academic literature: \citet{besbris_effect_2015} used Zillow boundaries to examine neighborhood stigma and \citet{albrecht_indicator_2014} used them to study neighborhood-level poverty in New York.
+The second set of geometries defines our municipal-scale study sites using 51 separate TIGER/Line shapefiles (again, 2016) of US census bureau \emph{places} within all 50 states plus DC.
+We discard the subset of \emph{census-designated places} (i.e., small unincorporated communities) in these data, while retaining every US city and town.
+The third set of geometries defines the neighborhood-scale study sites using 42 separate shapefiles from Zillow, a real estate database company.
+These shapefiles contain neighborhood boundaries for major cities in 41 states plus DC.
+This fairly new data set comprises nearly 7,000 neighborhoods, but as \citet{schernthanner_spatial_2016} point out, Zillow does not publish the methodology used to construct these boundaries.
+However, despite its newness it already has a track record in the academic literature: \citet{besbris_effect_2015} use Zillow boundaries to examine neighborhood stigma and \citet{albrecht_indicator_2014} use them to study neighborhood-level poverty in New York.
 
-For all of these geometries, we use OSMnx to download the (drivable, public) street network contained within the geometry, a process described in detail in \cite{boeing_osmnx:_2017} and summarized here.
-First OSMnx buffers each geometry by 0.5 km, then downloads the OpenStreetMap \enquote{nodes} and \enquote{ways} within this buffered geometry.
-Next it constructs a street network from these data, corrects the topology, calculates street counts per node, then truncates the network to the original, desired polygon (ensuring that intersections are not considered cul-de-sacs simply because an incident edge connects to a node outside the desired polygon).
+For each of these geometries, we use OSMnx to download the (drivable, public) street network it contains, a process described in detail in \cite{boeing_osmnx:_2017} and summarized here.
+First OSMnx buffers each geometry by 0.5 km, then downloads the OpenStreetMap \enquote{nodes} and \enquote{ways} within this buffer.
+Next it constructs a street network from these data, corrects the topology, calculates street counts per node, then truncates the network to the original, desired polygon (ensuring that intersections are not considered dead-ends simply because an incident edge connects to a node outside the desired polygon).
 OSMnx saves each of these networks to disk as GraphML and shapefiles.
-Finally, it calculates metric and topological measures for each network.
+Finally, it calculates metric and topological measures for each network, described below.
 Such measures extend the toolkit commonly used in urban form studies \cite{talen_measuring_2003, ewing_travel_2010}.
@@
 
-Brief descriptions of these measures are discussed here, but extended technical definitions and algorithms can be found in e.g.
+Brief descriptions of these OSMnx-calculated measures are discussed here, but extended technical definitions and algorithms can be found in e.g.
 \cite{trudeau_introduction_1994, albert_statistical_2002, dorogovtsev_evolution_2002, brandes_network_2005, costa_characterization_2007, newman_structure_2003, newman_networks:_2010, barthelemy_spatial_2011, cranmer_navigating_2017}.
-Average street length, the mean edge length in the undirected representation of the graph, serves as a linear proxy for block size and indicates how fine-grained or coarse-grained the network is.
-Node density is the number of nodes divided by the area covered by the network.
-Intersection density is the node density of the set of nodes with more than one street emanating from them (thus excluding dead-ends).
-The edge density is the sum of all edge lengths divided by the area, and the physical street density is the sum of all edges in the undirected representation of the graph divided by the area.
+\emph{Average street length}, the mean edge length in the undirected representation of the graph, serves as a linear proxy for block size and indicates how fine-grained or coarse-grained the network is.
+\emph{Node density} is the number of nodes divided by the area covered by the network.
+\emph{Intersection density} is the density of the set of nodes with more than one street emanating from them (thus excluding dead-ends).
+\emph{Edge density} is the sum of all edge lengths divided by area, and the physical \emph{street density} is the sum of all edges in the undirected representation of the graph divided by area.
 These density measures all provide further indication of how fine-grained the network is.
-Finally, the average circuity represents the average ratio between an edge length and the straight-line distance between the two nodes it links.
+Finally, \emph{average circuity} represents the average ratio between an edge length and the straight-line distance between the two nodes it links.
 
-The average node degree, or mean number of edges incident to each node, quantifies how well the nodes are connected on average.
-Similarly, but more concretely, the average streets per node measures the mean number of physical streets (i.e., edges in the undirected representation of the graph) that emanate from each intersection and dead-end.
-This adapts the average node degree for physical form rather than directed circulation.
-The statistical and spatial distributions of number of streets per intersection characterize the type, prevalence, and dispersion of intersection connectedness in the network.
-Connectivity measures the minimum number of nodes or edges that must be removed from a connected graph to disconnect it.
-The average node connectivity (ANC) of a network---the mean number of internally node-disjoint paths between each pair of nodes in the graph---represents the expected number of nodes that must be removed to disconnect a randomly selected pair of non-adjacent nodes \cite{beineke_average_2002,dankelmann_bounds_2003}.
+The \emph{average node degree}, or mean number of edges incident to each node, quantifies how well the nodes are connected on average.
+Similarly, but more concretely, the \emph{average streets per node} measures the mean number of physical streets (i.e., edges in the undirected representation of the graph) that emanate from each intersection and dead-end.
+This adapts the average node degree for physical form rather than directed flow.
+The statistical and spatial distributions of number of streets per node characterize the type, prevalence, and dispersion of intersection connectedness in the network.
+\emph{Connectivity} measures the minimum number of nodes or edges that must be removed from a connected graph to disconnect it.
+The \emph{average node connectivity} (ANC) of a network---the mean number of internally node-disjoint paths between each pair of nodes in the graph---represents the expected number of nodes that must be removed to disconnect a randomly selected pair of non-adjacent nodes \cite{beineke_average_2002,dankelmann_bounds_2003}.
 
-The clustering coefficient of a node is the ratio of the number of edges between its neighbors to the maximum possible number of edges that could exist between these neighbors \cite{jiang_topological_2004, opsahl_clustering_2009}.
+The \emph{clustering coefficient} of a node is the ratio of the number of edges between its neighbors to the maximum possible number of edges that could exist between these neighbors \cite{jiang_topological_2004, opsahl_clustering_2009}.
 The weighted clustering coefficient weights this ratio by edge length and the average clustering coefficient is the mean of the clustering coefficients of all the nodes in the network.
 Centrality indicates the importance of nodes in a network \cite{huang_trajgraph:_2016, zhong_revealing_2017}.
-Betweenness centrality evaluates the number of shortest paths that pass through each node \cite{barthelemy_betweenness_2004}.
+\emph{Betweenness centrality} evaluates the number of shortest paths that pass through each node \cite{barthelemy_betweenness_2004}.
 The maximum betweenness centrality (MBC) in a network specifies the proportion of shortest paths that pass through the most important node.
 This is an indicator of resilience: networks with a high MBC are more prone to failure or inefficiency should this single choke point fail.
-Closeness centrality ranks nodes as more central if they are on average closer to all other nodes.
-Finally, PageRank ranks nodes based on the structure of incoming links and the rank of the source node \cite{brin_anatomy_1998, jiang_predicting_2008, agryzkov_algorithm_2012, chin_geographically_2015, gleich_pagerank_2015}.
+Finally, \emph{PageRank} ranks nodes based on the structure of incoming links and the rank of the source node \cite{brin_anatomy_1998, jiang_predicting_2008, agryzkov_algorithm_2012, chin_geographically_2015, gleich_pagerank_2015}.
 
 In total, this study cross-sectionally analyzes 27,009 networks: 497 urbanized areas' street networks, 19,655 cities' and towns' street networks, and 6,857 neighborhoods' street networks.
-These sample sizes are larger than those in any previous similar study.
-In the following sections, we present the findings of these analyses at the metropolitan scale, the municipal scale, the neighborhood scale, and through a case study looking deeper at the neighborhood-scale street networks in the city of San Francisco.
+These sample sizes are larger than those of any previous similar study.
+The following section presents the findings of these analyses at metropolitan, municipal, and neighborhood scales.
 
@@
 
-\subsection{Analysis of Metropolitan-Scale Street Networks}
+\subsection{Metropolitan-Scale Street Networks}
 
-Table \ref{table01} shows the considerable variation in street network characteristics across the entire data set of 497 urbanized areas.
-This is unsurprising: urbanized areas span a wide spectrum of sizes, from the Delano, CA Urbanized Area's 26 km\textsuperscript{2} to the New York--Newark, NY--NJ--CT Urbanized Area's 8,937 km\textsuperscript{2}---thus, density and count-based measures demonstrate substantial variance.
+Table \ref{table01} presents summary statistics for the entire data set of 497 urbanized areas.
+These urbanized areas span a wide range of sizes, from the Delano, CA Urbanized Area's 26 km\textsuperscript{2} to the New York--Newark, NY--NJ--CT Urbanized Area's 8,937 km\textsuperscript{2}---thus, density and count-based measures demonstrate substantial variance.
 Further, these urbanized areas span a wide spectrum of terrains, development eras and paradigms, and cultures.
@@
 Delano, CA's urbanized area---America's smallest---has 874 intersections and 222,328 meters of linear street (or 964 and 231,000 meters if including service roads).
-The typical (Table \ref{table01}, \emph{median}) American urbanized area is approximately 185 km\textsuperscript{2} in land area, has 5,830 intersections, and 1.3 million linear meters of street.
-Its street network is about 7.4\% more circuitous than straight-line, as-the-crow-flies edges between nodes would be.
+The typical American urbanized area is approximately 185 km\textsuperscript{2} in land area, has 5,830 intersections, and 1.3 million linear meters of street.
+Its street network is about 7.4\% more circuitous than straight-line as-the-crow-flies edges between nodes would be.
 The most circuitous network is 14\% more circuitous than straight-line would be, and least is only 2\%.
-Looking at network complexity in terms of density and connectivity, in the typical urbanized area, the average street segment length (a proxy for block size) is 160 meters.
+Looking density, grain, and connectivity in the typical urbanized area, the average street segment length (a proxy for block size) is 160 meters.
 The longest average street segment is the 226-meter average of urbanized Danbury, CT.
@@
 The typical urbanized area has 26 intersections per km\textsuperscript{2}.
-Both the densest and the sparsest are in the deep south: the sparsest is 12.5 (Gainesville, GA urbanized area) and the densest is 49.4 (New Orleans urbanized area).
-However, New Orleans is an anomaly in the deep south.
-Figure \ref{fig01} depicts the intersection density in each American urbanized area, from lowest density in dark red to highest density in light yellow.
-The map makes clear how the highest intersection densities are concentrated to the west of the Mississippi River.
-The lowest intersection densities are concentrated in a belt running from Louisiana, up through the Carolinas and Appalachians, and into New England.
+Both the densest and the sparsest are in the Deep South: the sparsest is 12.5 (Gainesville, GA urbanized area) and the densest is 49.4 (New Orleans urbanized area).
+However, New Orleans is an anomaly in the Deep South.
+Figure \ref{fig01} depicts the intersection density of each American urbanized area: the highest intersection densities concentrate west of the Mississippi River, while the lowest concentrate in a belt running from Louisiana, through the Carolinas and Appalachians, and into New England.
 In general, only the largest cities on the east coast (e.g., Boston, New York, Philadelphia, Washington) and Florida escape this trend.
 
-The distribution of node types (i.e., intersections and dead ends) provides a clear indicator of network connectedness.
-The typical urbanized area has 2.8 streets per intersection on average: lots of 3-way intersections, fewer cul-de-sacs, and even fewer 4-way intersections.
-The grid-like San Angelo, TX urbanized area has the most streets per node (3.2) on average, and (outside of Puerto Rico, which contains the seven lowest urbanized areas) the sprawling, disconnected Lexington Park, MD urbanized area has the fewest (2.2).
-These two urbanized areas fit the trend seen in the spatial distribution across the US in Figure \ref{fig01}: urbanized areas in the great plains and Midwest have particularly high numbers of streets per node on average, indicating more grid-like, connected networks.
+The distribution of node types (i.e., intersections and dead ends) provides an indicator of network connectedness.
+The typical urbanized area averages 2.8 streets per intersection: many 3-way intersections, fewer dead-ends, and even fewer 4-way intersections.
+The gridlike San Angelo, TX urbanized area has the most streets per node (3.2) on average, and (outside of Puerto Rico, which contains the seven lowest urbanized areas) the sprawling, disconnected Lexington Park, MD urbanized area has the fewest (2.2).
+These two urbanized areas fit the trend seen in the spatial distribution across the US in Figure \ref{fig01}: urbanized areas in the great plains and Midwest have particularly high numbers of streets per node on average, indicating more gridlike, connected networks.
 Cities in the southern and western US tend to have fewer streets per node, reflecting more dead-ends and a disconnected network.
@@
 In the typical urbanized area, 18\% of nodes are 4-way intersections, 59\% are 3-way intersections, and 21\% are dead-ends.
-However, this distribution varies somewhat between urbanized areas.
-Examining a small sample of 9 urbanized areas, chosen to maximize variance, reveals this in clearer detail.
-In Figure \ref{fig02}, urban Atlanta and Chattanooga have very high proportions of dead-ends, each over 30\% of all nodes, and very few 4-way intersections, indicating a disconnected street pattern.
+However, this distribution varies somewhat: examining a small sample of nine urbanized areas, chosen to maximize variance, reveals this in clearer detail.
+In Figure \ref{fig02}, urban Atlanta and Chattanooga have very high proportions of dead-ends---each over 30\% of all nodes---and very few 4-way intersections, indicating a disconnected street pattern.
 The urbanized areas of Phoenix, Boston, Detroit, and Chattanooga have particularly high proportions of 3-way intersections, each over 60\%, indicating a prevalence of T-intersections.
-Conversely, Chicago, New Orleans, Duluth, and Lubbock have high proportions of 4-way intersections, indicating more grid-like connected networks.
+Conversely, Chicago, New Orleans, Duluth, and Lubbock have high proportions of 4-way intersections, indicating more gridlike connected networks.
 But what is perhaps most notable about Figure \ref{fig02} is that these nine urbanized areas, despite being chosen to maximize variance, are overwhelmingly similar to each other.
-At the metropolitan scale, every large American urban agglomeration is characterized by a preponderance of 3-way intersections.
+Every large American urban agglomeration is characterized by a preponderance of 3-way intersections.
 
-The relationship between fine-grained networks and connectedness/grid-ness is, however, not clear-cut.
-Intersection density has only a weak, positive linear relationship with the proportion of 4-way intersections in the urbanized area ($r^{2}=0.17$).
-But the relationship between network circuity and grid-ness is somewhat clearer.
-Average circuity has a negative linear relationship with the proportion of 4-way intersections in the urbanized area ($r^{2}=0.43$).
+The relationship between fine-grained networks and connectedness/gridness is not, however, clear-cut: intersection density has only a weak, positive linear relationship with the proportion of 4-way intersections in the urbanized area ($r^{2}=0.17$).
+But the relationship between network circuity and gridness is somewhat clearer: average circuity has a negative linear relationship with the proportion of 4-way intersections ($r^{2}=0.43$).
 
@@
 
-Due to the substantial variation in urbanized area size, from 25 to 9,000 km\textsuperscript{2}, the preceding analysis covers a wide swath of metropolitan place types.
-To better compare apples-to-apples, Table \ref{table02} focuses on the 30 largest urbanized areas cross-sectionally to examine how their metric and topological measures compare.
-This provides more consistent spatial scales and extents, while offering a window into the similarities and differences in the built forms of America's largest urban agglomerations.
+Due to the substantial variation in urbanized area size, from 25 to 9,000 km\textsuperscript{2}, the preceding analysis covers a wide swath of metropolitan types.
+To better compare apples-to-apples, Table \ref{table02} focuses on the 30 largest urbanized areas cross-sectionally to examine their metric and topological measures.
+This provides more consistent spatial scales and extents, while offering a window into the similarities and differences in the built forms of America's largest agglomerations.
 
@@
 In other words, Cincinnati has a particularly coarse-grained network with few connections and paths.
-This can also be seen in the average street segment length, a proxy for block size: Cincinnati has the second highest (186 m), bested only by Cleveland (198 m).
+The average street segment length, a proxy for block size, also reflects this: Cincinnati has the second highest (186 m), bested only by Cleveland (198 m).
 In contrast, the two lowest are Denver's 138-meter average and San Juan's 131-meter average.
 
-These metropolitan-scale analyses consider trends in the built form at the scale of broad self-organized human systems and urbanized regions.
-However, they aggregate multiple heterogenous neighborhoods and municipalities---the scales of human life, urban design projects, and planning jurisdiction---into single units of analysis.
-To disaggregate and analyze finer characteristics, the following sections examine municipal- and neighborhood-scale street networks across the US.
+These metropolitan analyses consider trends in the built form at the scale of broad self-organized human systems and urbanized regions.
+However, they aggregate multiple heterogeneous municipalities and neighborhoods---the scales of human life, urban design projects, and planning jurisdiction---into single units of analysis.
+To disaggregate and analyze finer characteristics, the following subsections examine municipal- and neighborhood-scale street networks.
 
-\subsection{Analysis of Municipal-Scale Street Networks}
+\subsection{Municipal-Scale Street Networks}
 
-Table \ref{table03} shows similarly great variation in street network characteristics across the entire data set of 19,655 cities and towns.
-Again, this data set comprises the street networks of every incorporated city and town in the US.
+Table \ref{table03} presents summary statistics of street network characteristics across the entire data set of 19,655 cities and towns---every incorporated city and town in the US.
 Following recent work by \citet{barthelemy_modeling_2008} and \citet{strano_urban_2013}, we examine the relationship between the total street length $L$ and the number of nodes $n$.
-The former proposed a model of cities in which $L$ and $n$ scale nonlinearly as $n^{1/2}$, and the latter recently confirmed this empirically with a small sample of ten European cities' street networks.
-However, their small sample size may limit the generalizability and interpretability of their finding.
-To investigate this empirically, we examine the relationship between the total street length $L$ and the number of nodes $n$ for 19,655 US cities and towns and find a strong \emph{linear} relationship ($r^{2}=0.98$), as depicted in Figure \ref{fig03}, contradicting Strano et al.
+The former proposed a model of cities in which $L$ and $n$ scale nonlinearly as $n^{1/2}$, and the latter recently confirmed this empirically with a small sample of ten European cities.
+However, their small sample size may limit the generalizability of this finding.
+To investigate this empirically, we examine the relationship between $L$ and $n$ for 19,655 US cities and towns and instead find a strong \emph{linear} relationship ($r^{2}=0.98$), as depicted in Figure \ref{fig03}.
 We also find a similar linear relationship at the metropolitan and neighborhood scales.
@@
 
-Moreover, previous findings \citep[e.g.,][]{masucci_random_2009} suggest the distribution of street segment lengths in an urban street network follows a power-law distribution.
+Previous findings \citep[e.g.,][]{masucci_random_2009} also suggest street segment lengths in an urban network follow a power-law distribution.
 However, we find that these networks instead generally follow a lognormal distribution.
@@
 
-This study finds that such cities are not uncommon in the US, particularly between the Mississippi River and the Rocky Mountains: the Great Plains states are characterized by a unique street network form that is both orthogonal and reasonably dense.
-The former is partly the result of topography (flat, plains terrain that allows idealized grids) and design history (platting and development during the late nineteenth century) that favor orthogonal grids, as discussed earlier.
-The latter results from the fact that most towns across the Great Plains exhibit minimal suburban sprawl.
-Thus, the municipal boundaries snugly embrace the grid-like street network, without extending to accommodate the vast peripheral belt of twentieth century sprawl, circuity, and \enquote{loops and lollipops} \cite{southworth_streets_1997} that characterizes cities in California that were settled in the same era but later subjected to substantial suburbanization.
+We find that such cities are not uncommon in the US, particularly between the Mississippi River and the Rocky Mountains: the Great Plains states are characterized by a unique street network form that is both orthogonal and reasonably dense.
+The \emph{former} is partly the result of topography (flat terrain that allows idealized grids) and design history (rapid platting and development during the late nineteenth century) that favor orthogonal grids, as discussed earlier.
+The \emph{latter} results from the fact that most towns across the Great Plains exhibit minimal suburban sprawl.
+Thus, municipal boundaries snugly embrace the gridlike street network, without extending to accommodate a vast peripheral belt of twentieth century sprawl, circuity, and \enquote{loops and lollipops} \cite{southworth_streets_1997} that characterizes cities in California that were settled in the same era but later subjected to substantial suburbanization.
 
@@
 	\includegraphics[width=0.7\textwidth]{media/fig03.png}
-	\caption{The linear relationship between total street length and number of nodes in the street networks of 6,857 US neighborhoods (top) and 19,655 US cities and towns (bottom).}
+	\caption{The linear relationship between total street length and number of nodes in the street networks of 6,857 US neighborhoods (left) and 19,655 US cities and towns (right).}
 	\label{fig03}
@@
 For example, if we measure connectedness in terms of average number of streets per node at the city-scale and then aggregate these cities by state (Table \ref{table04}), we find Nebraska, Kansas, South Dakota, Montana, North Dakota, Oklahoma, and Iowa have, in order, the highest medians (Figure \ref{fig04}).
-This indicates the most grid-like networks.
-If we measure density and connectedness in terms of intersection density at the city-scale and then aggregate these cities by state, we find Rhode Island, Nebraska, New Jersey, Kansas, and Montana have, in order, the highest medians.
-We again see three Great Plains states near the top, alongside densely populated East Coast states.
-Nebraska also has the smallest block sizes (measured via the proxy of average street segment length) while the largest are concentrated in the deep South, upper New England, and Utah (Figure \ref{fig04}).
+This indicates the most gridlike networks.
+If we measure intersection density at the city-scale and then aggregate these cities by state, we find Rhode Island, Nebraska, New Jersey, Kansas, and Montana have, in order, the highest medians.
+We again see three Great Plains states near the top, alongside small densely populated East Coast states.
+Nebraska also has the smallest block sizes (measured via the proxy of average street segment length) while the largest concentrate in the Deep South, upper New England, and Utah (Figure \ref{fig04}).
 
-Municipal boundaries vary greatly in their extents around the built-up area.
-For example, while Rhode Island averages 56 intersections/km\textsuperscript{2} in its cities and towns, Alaska averages only 1.3.
-This is an artifact of Alaska's municipal boundaries often extending thousands of square kilometers beyond the actual built-up area.
+However, municipal boundaries vary greatly in their extents around the built-up area.
+While Rhode Island averages 56 intersections/km\textsuperscript{2} in its cities and towns, Alaska averages only 1.3, because the latter's municipal boundaries often extend thousands of km\textsuperscript{2} beyond the actual built-up area.
 In fact, Alaska has four cities (Anchorage, Juneau, Sitka, and Wrangell) with such large municipal extents that their land areas exceed that of the state of Rhode Island.
-These state-level aggregations of municipal-scale street network characteristics show clear variation across the country that reflect topography, economies, culture, planning paradigms, and settlement eras.
+These state-level aggregations of municipal street network characteristics show clear variation across the country that reflect topography, economies, culture, planning paradigms, and settlement eras.
 But they also aggregate and thus obfuscate the variation within each state and within each city.
@@
 
-\subsection{Analysis of Neighborhood-Scale Street Networks}
+\subsection{Neighborhood-Scale Street Networks}
 
-Thus far, we have examined every urban street network in the US at the metropolitan and municipal scales.
-This analysis has focused on the complexity on the network in terms of density, resilience, and connectedness.
-While the metropolitan scale captures the emergent character of the wider region's complex system, and the municipal scale captures planning decisions made by a single city government, the neighborhood scale best represents the scale of individual urban design interventions into the urban form.
+We have thus far examined every urban street network in the US at the metropolitan and municipal scales.
+While the metropolitan scale captures the emergent character of the wider region's complex system, and the municipal scale captures planning decisions made by a single city government, the neighborhood best represents the scale of individual urban design interventions into the urban form.
 Further, this scale more commonly reflects individual designs, eras, and paradigms in street network development than the \enquote{many hands, many eras} evolution of form at larger scales.
@@
 
-Table \ref{table05} presents summary statistics for this data set.
-Compared to the summary statistics presented at the metropolitan scale (Table \ref{table01}) and the municipal scale (Table \ref{table03}), here we see much greater variance.
-This is expected, given the smaller network sizes at the neighborhood scale.
+Table \ref{table05} presents summary statistics for these 6,857 neighborhoods.
+Compared to the metropolitan and municipal scales, we see much greater variance here, as expected, given the smaller network sizes at the neighborhood scale.
 A few neighborhoods have no intersections within their Zillow-defined boundaries, resulting in a minimum intersection density of 0 across the data set.
-Meanwhile, the small neighborhood of Cottages North in Davis, California has the highest intersection density in the country, 444/km\textsuperscript{2}, largely as an artifact of its small area as the denominator.
-
+Meanwhile, the small neighborhood of Cottages North in Davis, California has the highest intersection density in the country, 444/km\textsuperscript{2}, largely an artifact of its small area as the denominator.
 Nationwide, the typical neighborhood averages 2.9 streets per intersection, reflecting the prevalence of 3-way intersections in the US, discussed earlier.
-The median proportions of each node type are 14.5\% for cul-de-sacs, 57.4\% for 3-way intersections, and 23.4\% for 4-way intersections.
+The median proportions of each node type are 14.5\% for dead-ends, 57.4\% for 3-way intersections, and 23.4\% for 4-way intersections.
 The typical neighborhood averages 135-meter street segment lengths and 46.4 intersections per km\textsuperscript{2}.
-At the neighborhood scale (sample size of 6,857 in this analysis) we again find the same strong \emph{linear} relationship between total street length and the number of nodes in a network ($r^{2}=0.98$), as seen in Figure \ref{fig03}, contradicting the smaller-sample findings of \citet{strano_urban_2013}.
 
@@
 In this filtered set, the five neighborhoods with the highest intersection density are all in central Philadelphia.
-Central neighborhoods are common at the top of this list, including Point Breeze, Philadelphia; Central Boston; Central City, New Orleans; Downtown Tampa; and Downtown Portland (Figure \ref{fig05}).
+Central neighborhoods are common at the top of this list, including Point Breeze, Philadelphia; Central Boston; Central City, New Orleans; Downtown Tampa; and Downtown Portland.
 The three neighborhoods with the lowest intersection density are on the outskirts of Anchorage, Alaska.
-In this filtered set, the neighborhoods with the greatest average number of streets per node tend to be older neighborhoods with orthogonal grids, such as Virginia Park, Tampa; Outer Sunset, San Francisco; and New Orleans' French Quarter.
+In this filtered set, the greatest average numbers of streets per node tend to be in older neighborhoods with orthogonal grids, such as Virginia Park, Tampa; Outer Sunset, San Francisco; and New Orleans' French Quarter.
 The neighborhoods with the lowest tend to be sprawling and often hilly suburbs far from the urban core, such as Scholl Canyon in Glendale, California or Sonoma Ranch in San Antonio, Texas.
@@
 
-For comparison, Figure \ref{fig05} compares one square mile of the centers of Philadelphia, Portland, and San Francisco with one square mile of each of their suburbs.
+To illustrate these morphologies, Figure \ref{fig05} compares one square mile of the centers of Philadelphia, Portland, and San Francisco with one square mile of each of their suburbs.
 The connectedness and density of the central cities is clear, as is the disconnectivity of their suburbs.
 In fact, the suburbs have more in common with one another---despite being hundreds or thousands of miles apart---than they do with their central city neighbors, suggesting that land use and an era's prevailing design paradigm is paramount to geographical localism and regional context.
-The top row of Figure \ref{fig05} represents an era of urban planning and development that preceded the automobile, while the bottom row reflects the exclusionary zoning and mid-late twentieth century era of automobility in residential suburb design---namely the \enquote{loops and lollipops} and the \enquote{lollipops on a stick} design patterns identified by \citet{southworth_streets_1997}.
+The top row of Figure \ref{fig05} represents an era of planning and development that preceded the automobile, while the bottom row reflects the exclusionary zoning and mid--late twentieth century era of automobility in residential suburb design---namely the \enquote{loops and lollipops} and the \enquote{lollipops on a stick} design patterns identified by \citet{southworth_streets_1997}.
 
-Finally we briefly take a closer look at San Francisco, California's neighborhoods alone for a clear cross-sectional analysis with consistent geography to look across planning eras and design paradigms.
+Finally we briefly take a closer look at San Francisco, California's neighborhoods alone for a clear cross-sectional analysis with consistent geography to examine resilience through the MBC and ANC measures.
 Due to its highly connected orthogonal grid, the Outer Sunset neighborhood has the lowest MBC---only 9.6\% of all shortest paths pass through its most important node.
-By contrast, in Chinatown, 36\% of shortest paths pass through the most important node, and in Twin Peaks it is 37\%.
+By contrast, 36\% of Chinatown's shortest paths pass through its most important node, and in Twin Peaks it is 37\%.
 In Chinatown, this is the result of the small neighborhood comprising only a few streets and that these streets are one-way, forcing paths through very few routing options.
@@
 
-Average node connectivity is another proxy for resilience.
+ANC is another indicator of resilience.
 In San Francisco, Twin Peaks' network has the lowest ANC: on average only 1.05 nodes must be removed to disconnect a randomly selected pair of nodes.
-Outer Sunset has the highest ANC, 3.2.
+Outer Sunset has the highest ANC: 3.2.
 These concur with the above descriptions of these networks.
-However, some central-city orthogonal grid networks with many 4-way intersections---such as Downtown, Chinatown, and the Financial District---have surprisingly low ANC: 1.5, 1.3, and 1.6 respectively.
+However, some central San Francisco orthogonal grid networks with many 4-way intersections---such as Downtown, Chinatown, and the Financial District---have surprisingly low ANCs: 1.5, 1.3, and 1.6 respectively.
 These neighborhoods comprise primarily one-way streets.
 Although they have dense, highly connected networks, they can be easily disconnected given that traffic cannot flow bidirectionally.
-These three neighborhoods also experience the greatest increase in ANC if all the edges are made undirected: Chinatown's increases 87\%, Downtown's 80\%, and the Financial District's 75\%.
-By contrast, the street network of Outer Sunset sees only a 6\% increase due to it already comprising primarily bidirectional streets.
-Targeted conversion of one-way streets to bidirectional in Downtown, the Financial District, and Chinatown may yield substantial network resilience gains.
+These three neighborhoods also exhibit the greatest increase in ANC if all their edges are made undirected: Chinatown's increases 87\%, Downtown's 80\%, and the Financial District's 75\%.
+By contrast, Outer Sunset's street network sees only a 6\% increase due to it already comprising primarily bidirectional streets.
+Targeted conversion of one-way streets in networks like Downtown, the Financial District, and Chinatown may yield substantial resilience gains.
 
@@
 These findings suggest the influence of planning eras, design paradigms, transportation technologies, topography, and economics on US street network density, resilience, and connectedness.
-Sprawling suburban neighborhoods rank low on density, connectedness, and order.
+Overall, every large US metropolis is characterized by its preponderance of 3-way intersections.
+Sprawling suburban neighborhoods rank low on density and connectedness.
 The orthogonal grids we see in the downtowns of Portland and San Francisco have high density (i.e., intersection and street densities), connectedness (i.e., average number of streets per node), and order (based on circuity and statistical dispersion of node types), but low resilience in the presence of one-way streets, measured by MBC and ANC increases when switching from one-way to bidirectional edges.
@@
 
-The characteristics of city street network fundamentally depend on what \emph{city} means: municipal boundaries, urbanized areas, or certain central neighborhoods?
+The characteristics of city street networks fundamentally depend on what \emph{city} means: municipal boundaries, urbanized areas, or certain central neighborhoods?
 The first is a legal/political definition, but captures the scope of city planning authority and decision-making for top-down interventions into human circulation.

### 2017-12-14 draft/v1/latex/paper.tex
revision

--- 2017-03-29 latex/paper.tex
@@
-\documentclass{article}
+\documentclass[Afour,sageh,times]{sagej}
 
@@
 \usepackage{csquotes}
-\usepackage[numbers]{natbib}
-\usepackage[affil-it]{authblk}
+\usepackage{natbib}
+\usepackage{url}
 
+\usepackage{moreverb,url}
+\usepackage[colorlinks,bookmarksopen,bookmarksnumbered,citecolor=red,urlcolor=red]{hyperref}
+\newcommand\BibTeX{{\rmfamily B\kern-.05em \textsc{i\kern-.025em b}\kern-.08em
+		T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
+\def\volumeyear{2017}
+
+\renewcommand{\cite}{\citep}
 \doublespacing
-\bibliographystyle{chicago}
 
 \begin{document}
-\title{A Multi-Scale Analysis of 27,000 Urban Street Networks\footnote{This is a journal article manuscript currently under peer review.}}
-\author{Geoff Boeing\footnote{Email address: gboeing@berkeley.edu; corresponding author.}}
-\affil{Department of City and Regional Planning\\University of California, Berkeley}
-\date{\today}
+\title{A Multi-Scale Analysis of 27,000 Urban Street Networks}
+\author{Author Redacted}
+\date{March 2017}
 \maketitle
@@
 OpenStreetMap offers an underexplored source of worldwide geospatial data useful to urban researchers.
-This study uses the OSMnx software to automatically download and analyze over 27,000 street networks from OpenStreetMap at metropolitan, municipal, and neighborhood scales---namely, every US city and town ($N=19655$), census urbanized area ($N=497$), and Zillow-defined neighborhood ($N=6857$).
-It presents wide-ranging empirical findings on US urban form and street network characteristics, emphasizing measures relevant to urban design and morphology such as structure, connectedness, density, centrality, and resilience.
-In the past, street network data acquisition has been challenging and ad hoc, but this study illustrates the use of OSMnx and OpenStreetMap to conduct street network analysis with extremely large sample sizes, with clearly defined network definitions and extents for reproducibility, and using nonplanar, directed graphs.
-These street networks and their measures are now shared in a public repository for other researchers to use.
+This study uses the OSMnx software to automatically download and analyze 27,009 US street networks from OpenStreetMap at metropolitan, municipal, and neighborhood scales---namely, every US city and town ($N=19655$), census urbanized area ($N=497$), and Zillow-defined neighborhood ($N=6857$).
+It presents wide-ranging empirical findings on US urban form and street network characteristics, emphasizing measures relevant to graph theory, urban design, and morphology such as structure, connectedness, density, centrality, and resilience.
+In the past, street network data acquisition and processing has been challenging and ad hoc.
+This study illustrates the use of OSMnx and OpenStreetMap to consistently conduct street network analysis with extremely large sample sizes, with clearly defined network definitions and extents for reproducibility, and using nonplanar, directed graphs.
+These street networks and measures data sets have been shared in a public repository for other researchers to use.
 
@@
 Through path dependence, the spatial signatures of these land use laws, design paradigms, and planning instruments can still be seen today in these cities' urban forms and street networks.
+Cross-sectional analysis of American urban form can reveal these artifacts and histories through street networks at metropolitan, municipal, and neighborhood scales.
 
-A cross-sectional analysis of American urban form can reveal these artifacts and histories through street networks at metropolitan, municipal, and neighborhood scales.
 Network analysis is a natural approach to the study of cities as complex systems \cite{masucci_random_2009}.
-The empirical literature on street networks is growing ever richer, but suffers from some limitations---discussed in detail in \cite{boeing_osmnx:_2017} and summarized here.
-First, sample sizes tend to be fairly small due to data availability, gathering, and processing constraints: most studies in this literature that conduct topological or metric analysis tend to have sample sizes ranging around 10 to 50 networks \cite{buhl_topological_2006, cardillo_structural_2006, barthelemy_modeling_2008, strano_urban_2013}, which may limit the generalizability and interpretability of findings.
+The empirical literature on street networks is growing ever richer, but suffers from some limitations---discussed in detail in \cite{redacted_osmnx:_2017} and summarized here.
+First, sample sizes tend to be fairly small due to data availability, gathering, and processing constraints: most studies in this literature that conduct topological or metric analyses tend to have sample sizes ranging around 10 to 50 networks \cite{buhl_topological_2006, cardillo_structural_2006, barthelemy_modeling_2008, strano_urban_2013}, which may limit the generalizability and interpretability of findings.
 Second, reproducibility has been difficult when the dozens of decisions that go into analysis---such as spatial extents, topological simplification and correction, definitions of nodes and edges, etc.---are ad hoc or only partly reported \citep[e.g.,][]{porta_network_2006, strano_urban_2013}.
-Third, and related to the first two, studies frequently oversimplify to planar or undirected primal graphs for tractability \cite[e.g.,][]{buhl_topological_2006, cardillo_structural_2006, barthelemy_modeling_2008, masucci_random_2009}, or use dual graphs despite the loss of geographic and metric information \cite{batty_network_2005, jiang_integration_2002, ratti_space_2004, crucitti_centrality_2006, crucitti_centrality_2006-1}.
+Third, and related to the first two, studies frequently oversimplify to planar or undirected primal graphs for tractability \cite[e.g.,][]{buhl_topological_2006, cardillo_structural_2006, barthelemy_modeling_2008, masucci_random_2009}, or use dual graphs despite the loss of geographic, metric information \cite{batty_network_2005, jiang_integration_2002, ratti_space_2004, crucitti_centrality_2006, crucitti_centrality_2006-1}.
 
-This study addresses these limitations by conducting an urban morphological analysis of street networks at multiple scales, with large sample sizes, with clearly defined network definitions and extents for reproducibility, and using nonplanar, directed graphs.
-In particular, it examines urban street networks---represented as primal, nonplanar, weighted multidigraphs with possible self-loops---focusing on structure, connectedness, centrality, and resilience.
-This study analyzes 27,009 urban street networks at multiple overlapping scales across the US.
-Namely, it examines the street networks of every incorporated city and town (population size $N=19655$), census urbanized area ($N=497$), and Zillow-defined neighborhood ($N=6857$) in the US.
-To do so, it uses OSMnx---a new street network research toolkit \cite{boeing_osmnx:_2017}---to download, construct, and analyze these street networks at metropolitan, municipal, and neighborhood scales.
-These street networks and their measures data have been shared in a public repository for other researchers to use at [link redacted].
+This study addresses these limitations by conducting a morphological analysis of urban street networks at multiple scales, with large sample sizes, with clearly defined network definitions and extents for reproducibility, and using nonplanar, directed graphs.
+In particular, it examines 27,009 urban street networks---represented as primal, nonplanar, weighted multidigraphs with possible self-loops---at multiple overlapping scales across the US, focusing on structure, connectedness, centrality, and resilience.
+It examines the street networks of every incorporated city and town (population size $N=19655$), census urbanized area ($N=497$), and Zillow-defined neighborhood ($N=6857$) in the US.
+To do so, it uses OSMnx---a new street network research toolkit \cite{redacted_osmnx:_2017}---to download, construct, and analyze these street networks at metropolitan, municipal, and neighborhood scales.
+These street networks and measures data sets have been compiled and shared in a public repository for other researchers to repurpose at [redacted].
 
@@
 Second, it presents empirical findings of descriptive urban morphology for the street networks of every US city, urbanized area, and Zillow neighborhood.
-Third, it investigates with large sample sizes some claims previously made in the research literature with small sample sizes.
+Third, it investigates with large sample sizes some claims previously made in the research literature using small sample sizes.
 This paper is organized as follows.
 In the next section, it discusses the data sources, tools, and methods used to collect, construct, and analyze these street networks.
-Next, it presents findings of the analyses at metropolitan, municipal, and neighborhood scales.
+Then it presents findings of the analyses at metropolitan, municipal, and neighborhood scales.
 Finally, it concludes with a discussion of these findings and their implications for street network analysis, urban morphology, and city planning.
@@
 Street networks can be conceptualized as primal, directed, nonplanar graphs.
-A primal street network represents intersections as nodes and street segments as edges.
-A directed network has directed edges: that is, edge $uv$ points one-way from node $u$ to node $v$, but there need not exist a reciprocal edge $vu$.
-A planar network can be represented in two dimensions with its edges intersecting only at nodes \cite[e.g.,][]{viana_simplicity_2013, fischer_spatial_2014}.
+A \emph{primal} street network represents intersections as nodes and street segments as edges.
+A \emph{directed} network has directed edges: that is, edge $uv$ points one-way from node $u$ to node $v$, but there need not exist a reciprocal edge $vu$.
+A \emph{planar} network can be represented in two dimensions with its edges intersecting only at nodes \cite{viana_simplicity_2013, fischer_spatial_2014}.
 Most street networks are nonplanar---due to grade-separated expressways, overpasses, bridges, tunnels, etc.---but most quantitative studies of urban street networks represent them as planar \cite[e.g.,][]{buhl_topological_2006, cardillo_structural_2006, barthelemy_modeling_2008, masucci_random_2009, strano_urban_2013} for tractability because bridges and tunnels are uncommon in some cities.
@@
 This study uses OSMnx to download, construct, correct, analyze, and visualize street networks at metropolitan, municipal, and neighborhood scales.
-OSMnx is a Python-based research tool that easily downloads OpenStreetMap data for any place in the world and constructs it into a spatially-embedded graph-theoretic object for analysis and visualization \cite{boeing_osmnx:_2017}.
+OSMnx is a Python-based research tool that easily downloads OpenStreetMap data for any place name, address, or polygon in the world, then constructs it into a spatially-embedded graph-theoretic object for analysis and visualization \cite{redacted_osmnx:_2017}.
 OpenStreetMap is a collaborative worldwide mapping project that makes its spatial data available via various APIs \cite{corcoran_analysing_2013, jokar_arsanjani_openstreetmap_2015}.
@@
 
-To define the study sites and their spatial boundaries, this study uses three sets of geometries.
+To define the study sites and their spatial boundaries, we use three sets of geometries.
 The first defines the metropolitan-scale study sites using the 2016 TIGER/Line shapefile of US census bureau urban areas.
-Each census-defined urban areas comprises a set of tracts that meet a minimum density threshold \cite{u.s._census_bureau_2010_2010}.
+Each census-defined urban area comprises a set of tracts that meet a minimum density threshold \cite{u.s._census_bureau_2010_2010}.
 We retain only the \emph{urbanized areas} subset of these data (i.e., areas with greater than 50,000 population), discarding the small \emph{urban clusters} subset.
@@
 
-For each of these geometries, we use OSMnx to download the (drivable, public) street network it contains, a process described in detail in \cite{boeing_osmnx:_2017} and summarized here.
+For each of these geometries, we use OSMnx to download the (drivable, public) street network within it, a process described in detail in \cite{redacted_osmnx:_2017} and summarized here.
 First OSMnx buffers each geometry by 0.5 km, then downloads the OpenStreetMap \enquote{nodes} and \enquote{ways} within this buffer.
-Next it constructs a street network from these data, corrects the topology, calculates street counts per node, then truncates the network to the original, desired polygon (ensuring that intersections are not considered dead-ends simply because an incident edge connects to a node outside the desired polygon).
+Next it constructs a street network from these data, corrects the topology, calculates street counts per node, then truncates the network to the original, desired polygon.
 OSMnx saves each of these networks to disk as GraphML and shapefiles.
-Finally, it calculates metric and topological measures for each network, described below.
+Finally, it calculates metric and topological measures for each network, summarized below.
 Such measures extend the toolkit commonly used in urban form studies \cite{talen_measuring_2003, ewing_travel_2010}.
@@
 \emph{Edge density} is the sum of all edge lengths divided by area, and the physical \emph{street density} is the sum of all edges in the undirected representation of the graph divided by area.
-These density measures all provide further indication of how fine-grained the network is.
-Finally, \emph{average circuity} represents the average ratio between an edge length and the straight-line distance between the two nodes it links.
+These density measures suggest how fine-grained the network is.
+Finally, \emph{average circuity} represents the average ratio between each edge length and the straight-line distance between the two nodes it links.
 
-The \emph{average node degree}, or mean number of edges incident to each node, quantifies how well the nodes are connected on average.
+The \emph{average node degree}, or mean number of edges incident to each node, quantifies how well the nodes are connected.
 Similarly, but more concretely, the \emph{average streets per node} measures the mean number of physical streets (i.e., edges in the undirected representation of the graph) that emanate from each intersection and dead-end.
@@
 The statistical and spatial distributions of number of streets per node characterize the type, prevalence, and dispersion of intersection connectedness in the network.
-\emph{Connectivity} measures the minimum number of nodes or edges that must be removed from a connected graph to disconnect it.
-The \emph{average node connectivity} (ANC) of a network---the mean number of internally node-disjoint paths between each pair of nodes in the graph---represents the expected number of nodes that must be removed to disconnect a randomly selected pair of non-adjacent nodes \cite{beineke_average_2002,dankelmann_bounds_2003}.
+\emph{Connectivity} measures the minimum number of nodes or edges that must be removed from a connected network to disconnect it.
+The \emph{average node connectivity} (ANC) of a network---the mean number of internally node-disjoint paths between each pair of nodes---represents the expected number of nodes that must be removed to disconnect a randomly selected pair of non-adjacent nodes \cite{beineke_average_2002,dankelmann_bounds_2003}.
 
 The \emph{clustering coefficient} of a node is the ratio of the number of edges between its neighbors to the maximum possible number of edges that could exist between these neighbors \cite{jiang_topological_2004, opsahl_clustering_2009}.
-The weighted clustering coefficient weights this ratio by edge length and the average clustering coefficient is the mean of the clustering coefficients of all the nodes in the network.
-Centrality indicates the importance of nodes in a network \cite{huang_trajgraph:_2016, zhong_revealing_2017}.
-\emph{Betweenness centrality} evaluates the number of shortest paths that pass through each node \cite{barthelemy_betweenness_2004}.
+The weighted clustering coefficient weights this ratio by edge length and the average clustering coefficient is the mean of the clustering coefficients of all the network's nodes.
+\emph{Betweenness centrality} indicates the importance of nodes in a network by evaluating the number of shortest paths that pass through each node \cite{barthelemy_betweenness_2004,huang_trajgraph:_2016, zhong_revealing_2017}.
 The maximum betweenness centrality (MBC) in a network specifies the proportion of shortest paths that pass through the most important node.
@@
 
-\begin{figure}
-\includegraphics[width=1\textwidth]{media/fig01.png}
-\caption{Intersection density per urbanized area (left) and average streets per node per urbanized area (right) from lowest (pale yellow) to highest (dark red), in the contiguous US}
-\label{fig01}
-\end{figure}
-
 Nevertheless, looking across the data set provides a sense of the breadth of American metropolitan street networks.
@@
 Its street network is about 7.4\% more circuitous than straight-line as-the-crow-flies edges between nodes would be.
-The most circuitous network is 14\% more circuitous than straight-line would be, and least is only 2\%.
-Looking density, grain, and connectivity in the typical urbanized area, the average street segment length (a proxy for block size) is 160 meters.
+The most circuitous network is 14\% more circuitous than straight-line would be, and the least is only 2\%.
+Looking at density, grain, and connectivity in the typical urbanized area, the average street segment length (a proxy for block size) is 160 meters.
 The longest average street segment is the 226-meter average of urbanized Danbury, CT.
-Puerto Rican cities hold the top four positions for shortest average street segment length, but among the 50 states plus DC, the shortest average street segment is the 125.3-meter average of urbanized Tracy, CA, indicating a much finer street network.
-The urbanized area of Portland, Oregon, with its famously compact walkable blocks, ranks second at 125.5 meters on average.
+Puerto Rican cities hold the top four positions for shortest average street segment length, but among the 50 states plus DC, the shortest average street segment is the 125.3-meter average of urbanized Tracy, CA, indicating a fine-grained street network.
+The urbanized area of Portland, Oregon, with its famously compact walkable blocks, ranks second at 125.5 meters.
 
-\begin{table}
+\begin{table*}
 \centering
@@
 \end{tabular}
-\end{table}
-
-\begin{figure}
-\includegraphics[width=0.75\textwidth]{media/fig02.png}
-\caption{Distribution of node types in 9 urbanized areas, with number of streets emanating from the node on the $x$-axis and proportion of nodes of this type on the $y$-axis }
-\label{fig02}
-\end{figure}
+\end{table*}
 
@@
 
+\begin{figure*}
+	\includegraphics[width=1\textwidth]{fig01.png}
+	\caption{Intersection density per urbanized area (left) and average streets per node per urbanized area (right) from lowest (pale yellow) to highest (dark red), in the contiguous US}
+	\label{fig01}
+\end{figure*}
+
+\begin{figure*}
+	\includegraphics[width=0.7\textwidth]{fig02.png}
+	\caption{Distribution of node types in 9 urbanized areas, with number of streets emanating from the node on the $x$-axis and proportion of nodes of this type on the $y$-axis }
+	\label{fig02}
+\end{figure*}
+
 The distribution of node types (i.e., intersections and dead ends) provides an indicator of network connectedness.
@@
 The gridlike San Angelo, TX urbanized area has the most streets per node (3.2) on average, and (outside of Puerto Rico, which contains the seven lowest urbanized areas) the sprawling, disconnected Lexington Park, MD urbanized area has the fewest (2.2).
-These two urbanized areas fit the trend seen in the spatial distribution across the US in Figure \ref{fig01}: urbanized areas in the great plains and Midwest have particularly high numbers of streets per node on average, indicating more gridlike, connected networks.
+These fit the trend seen in the spatial distribution across the US in Figure \ref{fig01}: urbanized areas in the Great Plains and Midwest have particularly high numbers of streets per node on average, indicating more gridlike, connected networks.
 Cities in the southern and western US tend to have fewer streets per node, reflecting more dead-ends and a disconnected network.
@@
 \begin{landscape}
-\begin{table}
+\begin{table*}
 \centering
@@
 \end{tabular}
-\end{table}
+\end{table*}
 \end{landscape}
@@
 
-\begin{table}
+\begin{table*}
 \centering
@@
 \end{tabular}
-\end{table}
+\end{table*}
 
@@
 The \emph{latter} results from the fact that most towns across the Great Plains exhibit minimal suburban sprawl.
-Thus, municipal boundaries snugly embrace the gridlike street network, without extending to accommodate a vast peripheral belt of twentieth century sprawl, circuity, and \enquote{loops and lollipops} \cite{southworth_streets_1997} that characterizes cities in California that were settled in the same era but later subjected to substantial suburbanization.
+Thus, municipal boundaries snugly embrace the gridlike street network, without extending to accommodate a vast peripheral belt of twentieth century sprawl, circuity, and \enquote{loops and lollipops} \cite{southworth_streets_1997} that characterizes cities in e.g.
+California that were settled in the same era but later subjected to substantial suburbanization.
 
-\begin{figure}
-	\includegraphics[width=0.7\textwidth]{media/fig03.png}
+\begin{figure*}
+	\includegraphics[width=0.7\textwidth]{fig03.png}
 	\caption{The linear relationship between total street length and number of nodes in the street networks of 6,857 US neighborhoods (left) and 19,655 US cities and towns (right).}
 	\label{fig03}
-\end{figure}
+\end{figure*}
 
-\begin{figure}
-	\includegraphics[width=0.85\textwidth]{media/fig04.png}
+\begin{figure*}
+	\includegraphics[width=0.85\textwidth]{fig04.png}
 	\caption{Left: contiguous US states by median of mean streets per node in municipal street networks, from lowest (pale yellow) to highest (dark red).
@@
 	\label{fig04}
-\end{figure}
+\end{figure*}
 
 \newgeometry{margin=2cm}
-\begin{table}
+\begin{table*}
 \centering
@@
 \end{tabular}
-\end{table}
+\end{table*}
 \restoregeometry
@@
 The three neighborhoods with the lowest intersection density are on the outskirts of Anchorage, Alaska.
-In this filtered set, the greatest average numbers of streets per node tend to be in older neighborhoods with orthogonal grids, such as Virginia Park, Tampa; Outer Sunset, San Francisco; and New Orleans' French Quarter.
+In the filtered set, the greatest average numbers of streets per node tend to be in older neighborhoods with orthogonal grids, such as Virginia Park, Tampa; Outer Sunset, San Francisco; and New Orleans' French Quarter.
 The neighborhoods with the lowest tend to be sprawling and often hilly suburbs far from the urban core, such as Scholl Canyon in Glendale, California or Sonoma Ranch in San Antonio, Texas.
 
-\begin{figure}
-	\includegraphics[width=1\textwidth]{media/fig05.png}
+\begin{figure*}
+	\includegraphics[width=1\textwidth]{fig05.png}
 	\caption{Square-mile comparisons of central cities and their suburbs.
@@
 	\label{fig05}
-\end{figure}
+\end{figure*}
 
 To illustrate these morphologies, Figure \ref{fig05} compares one square mile of the centers of Philadelphia, Portland, and San Francisco with one square mile of each of their suburbs.
-The connectedness and density of the central cities is clear, as is the disconnectivity of their suburbs.
+The connectedness and fine-grain of the central cities are clear, as are the disconnectivity and coarse-grain of their suburbs.
 In fact, the suburbs have more in common with one another---despite being hundreds or thousands of miles apart---than they do with their central city neighbors, suggesting that land use and an era's prevailing design paradigm is paramount to geographical localism and regional context.
-The top row of Figure \ref{fig05} represents an era of planning and development that preceded the automobile, while the bottom row reflects the exclusionary zoning and mid--late twentieth century era of automobility in residential suburb design---namely the \enquote{loops and lollipops} and the \enquote{lollipops on a stick} design patterns identified by \citet{southworth_streets_1997}.
+The top row of Figure \ref{fig05} represents an era of planning and development that preceded the automobile, while the bottom row reflects the exclusionary zoning and mid--late twentieth century era of automobility in residential suburb design---namely the \enquote{loops and lollipops} and the \enquote{lollipops on a stick} design patterns \cite{southworth_streets_1997}.
 
@@
 By contrast, 36\% of Chinatown's shortest paths pass through its most important node, and in Twin Peaks it is 37\%.
-In Chinatown, this is the result of the small neighborhood comprising only a few streets and that these streets are one-way, forcing paths through very few routing options.
+In Chinatown, this is the result of the small neighborhood comprising only a few streets---and that these streets are one-way, forcing paths through very few routing options.
 In Twin Peaks, this is the result of hilly terrain and a disconnected network forcing paths through a small set of chokepoints that link separate subsections of the network.
@@
 This paper had three primary purposes.
-First, it examined urban morphology through metric and topological measures of these networks---particularly focusing on density, connectedness, and resilience.
-Second, these methods and findings demonstrate the use of OSMnx as a new street network research toolkit.
-They suggest to urban planners and scholars new methods for acquiring and analyzing street network data.
+First, it examined urban morphology through metric and topological measures of 27,009 US street networks---particularly focusing on density, connectedness, and resilience.
+Second, these methods and findings demonstrate the use of OSMnx as a new street network research toolkit, suggesting to urban planners and scholars new methods for acquiring and analyzing data.
 Third, it investigated claims about the distribution of street segment lengths and the relationship between the total street length and the number of nodes in a network.
-This study has made all of these network datasets---for 497 urbanized areas, 19,655 cities and towns, and 6,857 neighborhoods---along with all of their attribute data and morphological measures available in an online public repository for other researchers to study and repurpose at [redacted for peer review].
+This study has made all of these network datasets---for 497 urbanized areas, 19,655 cities and towns, and 6,857 neighborhoods---along with all of their attribute data and morphological measures available in an online public repository for other researchers to study and repurpose at [redacted].
 
-\bibliography{references}
+\begin{thebibliography}{54}
+	\providecommand{\natexlab}[1]{#1}
+	\providecommand{\url}[1]{\texttt{#1}}
+	\expandafter\ifx\csname urlstyle\endcsname\relax
+	\providecommand{\doi}[1]{doi: #1}\else
+	\providecommand{\doi}{doi: \begingroup \urlstyle{rm}\Url}\fi
+
+	\bibitem[Agryzkov et~al.(2012)Agryzkov, Oliver, Tortosa, and
+	Vicent]{agryzkov_algorithm_2012}
+	T.~Agryzkov, J.~L. Oliver, L.~Tortosa, and J.~F. Vicent.
+	\newblock An algorithm for ranking the nodes of an urban network based on the
+	concept of {PageRank} vector.
+	\newblock \emph{Applied Mathematics and Computation}, 219\penalty0
+	(4):\penalty0 2186--2193, 2012.
+	\newblock ISSN 00963003.
+	\newblock \doi{10.1016/j.amc.2012.08.064}.
+
+	\bibitem[Albert and Barabási(2002)]{albert_statistical_2002}
+	R.~Albert and A.-L. Barabási.
+	\newblock Statistical mechanics of complex networks.
+	\newblock \emph{Reviews of Modern Physics}, 74\penalty0 (1):\penalty0 47, 2002.
+	\newblock ISSN 0034-6861.
+	\newblock \doi{10.1103/RevModPhys.74.47}.
+
+	\bibitem[Albrecht and Abramovitz(2014)]{albrecht_indicator_2014}
+	J.~Albrecht and M.~Abramovitz.
+	\newblock Indicator {Analysis} for {Unpacking} {Poverty} in {New} {York}
+	{City}.
+	\newblock Technical report, CUNY, New York, NY, 2014.
+
+	\bibitem[Barthelemy(2004)]{barthelemy_betweenness_2004}
+	M.~Barthelemy.
+	\newblock Betweenness centrality in large complex networks.
+	\newblock \emph{The European Physical Journal B: Condensed Matter and Complex
+		Systems}, 38\penalty0 (2):\penalty0 163--168, 2004.
+	\newblock ISSN 1434-6028.
+	\newblock \doi{10.1140/epjb/e2004-00111-4}.
+
+	\bibitem[Barthelemy(2011)]{barthelemy_spatial_2011}
+	M.~Barthelemy.
+	\newblock Spatial networks.
+	\newblock \emph{Physics Reports}, 499\penalty0 (1-3):\penalty0 1--101, 2011.
+	\newblock ISSN 03701573.
+	\newblock \doi{10.1016/j.physrep.2010.11.002}.
+
+	\bibitem[Barthelemy and Flammini(2008)]{barthelemy_modeling_2008}
+	M.~Barthelemy and A.~Flammini.
+	\newblock Modeling {Urban} {Street} {Patterns}.
+	\newblock \emph{Physical Review Letters}, 100\penalty0 (13), 2008.
+	\newblock ISSN 0031-9007.
+	\newblock \doi{10.1103/PhysRevLett.100.138702}.
+
+	\bibitem[Batty(2005)]{batty_network_2005}
+	M.~Batty.
+	\newblock Network geography: {Relations}, interactions, scaling and spatial
+	processes in {GIS}.
+	\newblock In D.~J. Unwin and P.~Fisher, editors, \emph{Re-presenting
+		{Geographical} {Information} {Systems}}, pages 149--170. John Wiley \& Sons,
+	Chichester, England, 2005.
+
+	\bibitem[Beineke et~al.(2002)Beineke, Oellermann, and
+	Pippert]{beineke_average_2002}
+	L.~W. Beineke, O.~R. Oellermann, and R.~E. Pippert.
+	\newblock The {Average} {Connectivity} of a {Graph}.
+	\newblock \emph{Discrete Mathematics}, 252\penalty0 (1):\penalty0 31--45, 2002.
+	\newblock ISSN 0012-365X.
+	\newblock \doi{10.1016/S0012-365X(01)00180-7}.
+
+	\bibitem[Besbris et~al.(2015)Besbris, Faber, Rich, and
+	Sharkey]{besbris_effect_2015}
+	M.~Besbris, J.~W. Faber, P.~Rich, and P.~Sharkey.
+	\newblock Effect of neighborhood stigma on economic transactions.
+	\newblock \emph{Proceedings of the National Academy of Sciences}, 112\penalty0
+	(16):\penalty0 4994--4998, 2015.
+	\newblock ISSN 0027-8424.
+	\newblock \doi{10.1073/pnas.1414139112}.
+
+	\bibitem[Brandes and Erlebach(2005)]{brandes_network_2005}
+	U.~Brandes and T.~Erlebach, editors.
+	\newblock \emph{Network analysis: methodological foundations}.
+	\newblock Number 3418 in Lecture {Notes} in {Computer} {Science}. Springer,
+	Berlin, Germany, 2005.
+	\newblock ISBN 978-3-540-24979-5.
+
+	\bibitem[Brin and Page(1998)]{brin_anatomy_1998}
+	S.~Brin and L.~Page.
+	\newblock The anatomy of a large-scale hypertextual web search engine.
+	\newblock \emph{Computer Networks and ISDN Systems: Proceedings of the Seventh
+		International World Wide Web Conference}, 30\penalty0 (1-7):\penalty0
+	107--117, 1998.
+	\newblock ISSN 1389-1286.
+	\newblock \doi{10.1016/S0169-7552(98)00110-X}.
+
+	\bibitem[Buhl et~al.(2006)Buhl, Gautrais, Reeves, Solé, Valverde, Kuntz, and
+	Theraulaz]{buhl_topological_2006}
+	J.~Buhl, J.~Gautrais, N.~Reeves, R.~V. Solé, S.~Valverde, P.~Kuntz, and
+	G.~Theraulaz.
+	\newblock Topological patterns in street networks of self-organized urban
+	settlements.
+	\newblock \emph{The European Physical Journal B: Condensed Matter and Complex
+		Systems}, 49\penalty0 (4):\penalty0 513--522, 2006.
+	\newblock ISSN 1434-6028.
+	\newblock \doi{10.1140/epjb/e2006-00085-1}.
+
+	\bibitem[Cardillo et~al.(2006)Cardillo, Scellato, Latora, and
+	Porta]{cardillo_structural_2006}
+	A.~Cardillo, S.~Scellato, V.~Latora, and S.~Porta.
+	\newblock Structural properties of planar graphs of urban street patterns.
+	\newblock \emph{Physical Review E}, 73\penalty0 (6), 2006.
+	\newblock ISSN 1539-3755.
+	\newblock \doi{10.1103/PhysRevE.73.066107}.
+
+	\bibitem[Chin and Wen(2015)]{chin_geographically_2015}
+	W.-C.-B. Chin and T.-H. Wen.
+	\newblock Geographically {Modified} {PageRank} {Algorithms}: {Identifying} the
+	{Spatial} {Concentration} of {Human} {Movement} in a {Geospatial} {Network}.
+	\newblock \emph{PLoS ONE}, 10\penalty0 (10):\penalty0 e0139509, 2015.
+	\newblock ISSN 1932-6203.
+	\newblock \doi{10.1371/journal.pone.0139509}.
+
+	\bibitem[Corcoran et~al.(2013)Corcoran, Mooney, and
+	Bertolotto]{corcoran_analysing_2013}
+	P.~Corcoran, P.~Mooney, and M.~Bertolotto.
+	\newblock Analysing the growth of {OpenStreetMap} networks.
+	\newblock \emph{Spatial Statistics}, 3:\penalty0 21--32, 2013.
+	\newblock ISSN 22116753.
+	\newblock \doi{10.1016/j.spasta.2013.01.002}.
+
+	\bibitem[Costa et~al.(2007)Costa, Rodrigues, Travieso, and
+	Villas~Boas]{costa_characterization_2007}
+	L.~d.~F. Costa, F.~A. Rodrigues, G.~Travieso, and P.~R. Villas~Boas.
+	\newblock Characterization of complex networks: {A} survey of measurements.
+	\newblock \emph{Advances in Physics}, 56\penalty0 (1):\penalty0 167--242, 2007.
+	\newblock ISSN 0001-8732, 1460-6976.
+	\newblock \doi{10.1080/00018730601170527}.
+
+	\bibitem[Cranmer et~al.(2017)Cranmer, Leifeld, McClurg, and
+	Rolfe]{cranmer_navigating_2017}
+	S.~J. Cranmer, P.~Leifeld, S.~D. McClurg, and M.~Rolfe.
+	\newblock Navigating the range of statistical tools for inferential network
+	analysis.
+	\newblock \emph{American Journal of Political Science}, 61\penalty0
+	(1):\penalty0 237--251, 2017.
+	\newblock \doi{10.1111/ajps.12263}.
+
+	\bibitem[Crucitti et~al.(2006{\natexlab{a}})Crucitti, Latora, and
+	Porta]{crucitti_centrality_2006}
+	P.~Crucitti, V.~Latora, and S.~Porta.
+	\newblock Centrality measures in spatial networks of urban streets.
+	\newblock \emph{Physical Review E}, 73\penalty0 (3):\penalty0 036125,
+	2006{\natexlab{a}}.
+	\newblock ISSN 1539-3755.
+	\newblock \doi{10.1103/PhysRevE.73.036125}.
+
+	\bibitem[Crucitti et~al.(2006{\natexlab{b}})Crucitti, Latora, and
+	Porta]{crucitti_centrality_2006-1}
+	P.~Crucitti, V.~Latora, and S.~Porta.
+	\newblock Centrality in networks of urban streets.
+	\newblock \emph{Chaos: An Interdisciplinary Journal of Nonlinear Science},
+	16\penalty0 (1):\penalty0 015113, 2006{\natexlab{b}}.
+	\newblock ISSN 10541500.
+	\newblock \doi{10.1063/1.2150162}.
+
+	\bibitem[Dankelmann and Oellermann(2003)]{dankelmann_bounds_2003}
+	P.~Dankelmann and O.~R. Oellermann.
+	\newblock Bounds on the average connectivity of a graph.
+	\newblock \emph{Discrete Applied Mathematics}, 129\penalty0 (2):\penalty0
+	305--318, 2003.
+	\newblock ISSN 0166-218X.
+	\newblock \doi{10.1016/S0166-218X(02)00572-3}.
+
+	\bibitem[Dorogovtsev and Mendes(2002)]{dorogovtsev_evolution_2002}
+	S.~Dorogovtsev and J.~Mendes.
+	\newblock Evolution of networks.
+	\newblock \emph{Advances in Physics}, 51\penalty0 (4):\penalty0 1079--1187,
+	2002.
+	\newblock \doi{10.1080 /0001873011011251}.
+
+	\bibitem[Ewing and Cervero(2010)]{ewing_travel_2010}
+	R.~Ewing and R.~Cervero.
+	\newblock Travel and the {Built} {Environment}: {A} {Meta}-{Analysis}.
+	\newblock \emph{Journal of the American Planning Association}, 76\penalty0
+	(3):\penalty0 265--294, 2010.
+	\newblock ISSN 0194-4363.
+	\newblock \doi{10.1080/01944361003766766}.
+
+	\bibitem[Frizzelle et~al.(2009)Frizzelle, Evenson, Rodriguez, and
+	Laraia]{frizzelle_importance_2009}
+	B.~Frizzelle, K.~Evenson, D.~Rodriguez, and B.~Laraia.
+	\newblock The importance of accurate road data for spatial applications in
+	public health: customizing a road network.
+	\newblock \emph{International Journal of Health Geographics}, 8\penalty0 (24),
+	2009.
+	\newblock ISSN 1476-072X.
+	\newblock \doi{10.1186/1476-072X-8-24}.
+
+	\bibitem[Gleich(2015)]{gleich_pagerank_2015}
+	D.~F. Gleich.
+	\newblock {PageRank} {Beyond} the {Web}.
+	\newblock \emph{SIAM Review}, 57\penalty0 (3):\penalty0 321--363, 2015.
+	\newblock ISSN 0036-1445.
+	\newblock \doi{10.1137/140976649}.
+
+	\bibitem[Haklay(2010)]{haklay_how_2010}
+	M.~Haklay.
+	\newblock How {Good} is {Volunteered} {Geographical} {Information}? {A}
+	{Comparative} {Study} of {OpenStreetMap} and {Ordnance} {Survey} {Datasets}.
+	\newblock \emph{Environment and Planning B: Planning and Design}, 37\penalty0
+	(4):\penalty0 682--703, 2010.
+	\newblock ISSN 1472-3417.
+	\newblock \doi{10.1068/b35097}.
+
+	\bibitem[Huang et~al.(2016)Huang, Zhao, Ma, Yang, Ye, and
+	Zhang]{huang_trajgraph:_2016}
+	X.~Huang, Y.~Zhao, C.~Ma, J.~Yang, X.~Ye, and C.~Zhang.
+	\newblock {TrajGraph}: {A} {Graph}-{Based} {Visual} {Analytics} {Approach} to
+	{Studying} {Urban} {Network} {Centralities} {Using} {Taxi} {Trajectory}
+	{Data}.
+	\newblock \emph{IEEE Transactions on Visualization and Computer Graphics},
+	22\penalty0 (1):\penalty0 160--169, 2016.
+	\newblock ISSN 1077-2626.
+	\newblock \doi{10.1109/TVCG.2015.2467771}.
+
+	\bibitem[Jiang(2008)]{jiang_predicting_2008}
+	B.~Jiang.
+	\newblock Predicting human movement using {PageRank} in an urban environment.
+	\newblock 2008.
+
+	\bibitem[Jiang and Claramunt(2002)]{jiang_integration_2002}
+	B.~Jiang and C.~Claramunt.
+	\newblock Integration of space syntax into {GIS}: new perspectives for urban
+	morphology.
+	\newblock \emph{Transactions in GIS}, 6\penalty0 (3):\penalty0 295--309, 2002.
+	\newblock ISSN 1467-9671.
+	\newblock \doi{10.1111/1467-9671.00112}.
+
+	\bibitem[Jiang and Claramunt(2004)]{jiang_topological_2004}
+	B.~Jiang and C.~Claramunt.
+	\newblock Topological {Analysis} of {Urban} {Street} {Networks}.
+	\newblock \emph{Environment and Planning B: Planning and Design}, 31\penalty0
+	(1):\penalty0 151--162, 2004.
+	\newblock ISSN 0265-8135.
+	\newblock \doi{10.1068/b306}.
+
+	\bibitem[Jokar~Arsanjani et~al.(2015)Jokar~Arsanjani, Zipf, Mooney, and
+	Helbich]{jokar_arsanjani_openstreetmap_2015}
+	J.~Jokar~Arsanjani, A.~Zipf, P.~Mooney, and M.~Helbich, editors.
+	\newblock \emph{{OpenStreetMap} in {GIScience}}.
+	\newblock Lecture {Notes} in {Geoinformation} and {Cartography}. Springer
+	International, Cham, Switzerland, 2015.
+	\newblock ISBN 978-3-319-14279-1.
+
+	\bibitem[Lee(1979)]{lee_kansas_1979}
+	L.~B. Lee.
+	\newblock \emph{Kansas and the {Homestead} act, 1862-1905}.
+	\newblock The management of public lands in the {United} {States}. Ayer Company
+	Publishers, North Stratford, NH, 1979.
+	\newblock ISBN 978-0-405-11341-3.
+
+	\bibitem[Maron(2015)]{maron_how_2015}
+	M.~Maron.
+	\newblock How complete is {OpenStreetMap}?
+	\newblock \emph{Mapbox}, Nov. 2015.
+	\newblock URL \url{https://www.mapbox.com/blog/how-complete-is-openstreetmap/}.
+
+	\bibitem[Masucci et~al.(2009)Masucci, Smith, Crooks, and
+	Batty]{masucci_random_2009}
+	A.~P. Masucci, D.~Smith, A.~Crooks, and M.~Batty.
+	\newblock Random planar graphs and the {London} street network.
+	\newblock \emph{The European Physical Journal B: Condensed Matter and Complex
+		Systems}, 71\penalty0 (2):\penalty0 259--271, 2009.
+	\newblock ISSN 1434-6028.
+	\newblock \doi{10.1140/epjb/e2009-00290-4}.
+
+	\bibitem[Newman(2003)]{newman_structure_2003}
+	M.~E.~J. Newman.
+	\newblock The {Structure} and {Function} of {Complex} {Networks}.
+	\newblock \emph{SIAM Review}, 45\penalty0 (2):\penalty0 167--256, 2003.
+	\newblock ISSN 0036-1445.
+	\newblock \doi{10.1137/S003614450342480}.
+
+	\bibitem[Newman(2010)]{newman_networks:_2010}
+	M.~E.~J. Newman.
+	\newblock \emph{Networks: {An} {Introduction}}.
+	\newblock Oxford University Press, Oxford, England, 2010.
+	\newblock ISBN 978-0-19-920665-0.
+
+	\bibitem[Opsahl and Panzarasa(2009)]{opsahl_clustering_2009}
+	T.~Opsahl and P.~Panzarasa.
+	\newblock Clustering in weighted networks.
+	\newblock \emph{Social Networks}, 31\penalty0 (2):\penalty0 155--163, 2009.
+	\newblock ISSN 03788733.
+	\newblock \doi{10.1016/j.socnet.2009.02.002}.
+
+	\bibitem[O'Sullivan(2014)]{fischer_spatial_2014}
+	D.~O'Sullivan.
+	\newblock Spatial {Network} {Analysis}.
+	\newblock In M.~M. Fischer and P.~Nijkamp, editors, \emph{Handbook of
+		{Regional} {Science}}, pages 1253--1273. Springer-Verlag, Berlin, Germany,
+	2014.
+	\newblock ISBN 978-3-642-23429-3.
+
+	\bibitem[Over et~al.(2010)Over, Schilling, Neubauer, and
+	Zipf]{over_generating_2010}
+	M.~Over, A.~Schilling, S.~Neubauer, and A.~Zipf.
+	\newblock Generating web-based 3d {City} {Models} from {OpenStreetMap}: {The}
+	current situation in {Germany}.
+	\newblock \emph{Computers, Environment and Urban Systems}, 34\penalty0
+	(6):\penalty0 496--507, 2010.
+	\newblock ISSN 01989715.
+	\newblock \doi{10.1016/j.compenvurbsys.2010.05.001}.
+
+	\bibitem[Porta et~al.(2006)Porta, Crucitti, and Latora]{porta_network_2006}
+	S.~Porta, P.~Crucitti, and V.~Latora.
+	\newblock The network analysis of urban streets: {A} dual approach.
+	\newblock \emph{Physica A: Statistical Mechanics and its Applications},
+	369\penalty0 (2):\penalty0 853--866, 2006.
+	\newblock ISSN 03784371.
+	\newblock \doi{10.1016/j.physa.2005.12.063}.
+
+	\bibitem[Porterfield(2005)]{porterfield_homestead_2005}
+	J.~Porterfield.
+	\newblock \emph{The {Homestead} {Act} of 1862: {A} {Primary} {Source} {History}
+		of the {Settlement} of the {American} {Heartland} in the {Late} 19th
+		{Century}}.
+	\newblock Rosen Publishing Group, New York, NY, 1st edition, 2005.
+	\newblock ISBN 978-1-4042-0178-1.
+
+	\bibitem[Ratti(2004)]{ratti_space_2004}
+	C.~Ratti.
+	\newblock Space syntax: some inconsistencies.
+	\newblock \emph{Environment and Planning B: Planning and Design}, 31\penalty0
+	(4):\penalty0 487--499, 2004.
+	\newblock ISSN 1472-3417.
+	\newblock \doi{10.1068/b3019}.
+
+	\bibitem[Redacted(2017)]{redacted_osmnx:_2017}
+	R.~Redacted.
+	\newblock Title redacted.
+	\newblock \penalty0 (Redacted), 2017.
+	\newblock \doi{Redacted}.
+
+	\bibitem[Schernthanner et~al.(2016)Schernthanner, Asche, Gonschorek, and
+	Scheele]{schernthanner_spatial_2016}
+	H.~Schernthanner, H.~Asche, J.~Gonschorek, and L.~Scheele.
+	\newblock Spatial {Modeling} and {Geovisualization} of {Rental} {Prices} for
+	{Real} {Estate} {Portals}.
+	\newblock In O.~Gervasi, B.~Murgante, S.~Misra, A.~M. A.~C. Rocha, C.~M. Torre,
+	D.~Taniar, B.~O. Apduhan, E.~Stankova, and S.~Wang, editors,
+	\emph{Computational {Science} and {Its} {Applications} - {ICCSA} 2016},
+	number 9788 in Lecture {Notes} in {Computer} {Science}, pages 120--133.
+	Springer International, Cham, Switzerland, 2016.
+	\newblock ISBN 978-3-319-42110-0.
+	\newblock DOI: 10.1007/978-3-319-42111-7\_11.
+
+	\bibitem[Sherraden(2005)]{sherraden_inclusion_2005}
+	M.~Sherraden.
+	\newblock \emph{Inclusion in the {American} {Dream}: {Assets}, {Poverty}, and
+		{Public} {Policy}}.
+	\newblock Oxford University Press, Oxford, England, 2005.
+	\newblock ISBN 978-0-19-534709-8.
+
+	\bibitem[Southworth and Ben-Joseph(1997)]{southworth_streets_1997}
+	M.~Southworth and E.~Ben-Joseph.
+	\newblock \emph{Streets and the {Shaping} of {Towns} and {Cities}}.
+	\newblock McGraw-Hill, New York, NY, 1997.
+	\newblock ISBN 978-1-55963-916-3.
+
+	\bibitem[Strano et~al.(2013)Strano, Viana, da~Fontoura~Costa, Cardillo, Porta,
+	and Latora]{strano_urban_2013}
+	E.~Strano, M.~Viana, L.~da~Fontoura~Costa, A.~Cardillo, S.~Porta, and
+	V.~Latora.
+	\newblock Urban {Street} {Networks}, a {Comparative} {Analysis} of {Ten}
+	{European} {Cities}.
+	\newblock \emph{Environment and Planning B: Planning and Design}, 40\penalty0
+	(6):\penalty0 1071--1086, 2013.
+	\newblock ISSN 1472-3417.
+	\newblock \doi{10.1068/b38216}.
+
+	\bibitem[Talen(2003)]{talen_measuring_2003}
+	E.~Talen.
+	\newblock Measuring {Urbanism}: {Issues} in {Smart} {Growth} {Research}.
+	\newblock \emph{Journal of Urban Design}, 8\penalty0 (3):\penalty0 195--215,
+	2003.
+	\newblock ISSN 1357-4809.
+	\newblock \doi{10.1080/1357480032000155141}.
+
+	\bibitem[Trudeau(1994)]{trudeau_introduction_1994}
+	R.~J. Trudeau.
+	\newblock \emph{Introduction to {Graph} {Theory}}.
+	\newblock Dover Publications, New York, NY, 2nd edition, 1994.
+	\newblock ISBN 978-0-486-67870-2.
+
+	\bibitem[{U.S. Census Bureau}(2010)]{u.s._census_bureau_2010_2010}
+	{U.S. Census Bureau}.
+	\newblock 2010 {Census} {Urban} and {Rural} {Classification} and {Urban} {Area}
+	{Criteria}, 2010.
+	\newblock URL
+	\url{https://www.census.gov/geo/reference/ua/urban-rural-2010.html}.
+
+	\bibitem[Viana et~al.(2013)Viana, Strano, Bordin, and
+	Barthelemy]{viana_simplicity_2013}
+	M.~P. Viana, E.~Strano, P.~Bordin, and M.~Barthelemy.
+	\newblock The simplicity of planar networks.
+	\newblock \emph{Scientific Reports}, 3\penalty0 (3495):\penalty0 1--6, 2013.
+	\newblock ISSN 2045-2322.
+	\newblock \doi{10.1038/srep03495}.
+
+	\bibitem[Willis(2008)]{willis_openstreetmap_2008}
+	N.~Willis.
+	\newblock {OpenStreetMap} project completes import of {United} {States} {TIGER}
+	data.
+	\newblock \emph{Linux.com}, Jan. 2008.
+	\newblock URL
+	\url{https://www.linux.com/news/openstreetmap-project-completes-import-united-states-tiger-data}.
+
+	\bibitem[Wu et~al.(2005)Wu, Funk, Lurmann, and Winer]{wu_improving_2005}
+	J.~Wu, T.~H. Funk, F.~W. Lurmann, and A.~M. Winer.
+	\newblock Improving spatial accuracy of roadway networks and geocoded
+	addresses.
+	\newblock \emph{Transactions in GIS}, 9\penalty0 (4):\penalty0 585--601, 2005.
+	\newblock ISSN 1467-9671.
+	\newblock \doi{10.1111/j.1467-9671.2005.00236.x}.
+
+	\bibitem[Zhong et~al.(2017)Zhong, Schläpfer, Arisona, Batty, Ratti, and
+	Schmitt]{zhong_revealing_2017}
+	C.~Zhong, M.~Schläpfer, S.~M. Arisona, M.~Batty, C.~Ratti, and G.~Schmitt.
+	\newblock Revealing centrality in the spatial structure of cities from human
+	activity patterns.
+	\newblock \emph{Urban Studies}, 54\penalty0 (2):\penalty0 437--455, 2017.
+	\newblock ISSN 0042-0980.
+	\newblock \doi{10.1177/0042098015601599}.
+
+	\bibitem[Zielstra and Hochmair(2011)]{zielstra_comparative_2011}
+	D.~Zielstra and H.~Hochmair.
+	\newblock Comparative {Study} of {Pedestrian} {Accessibility} to {Transit}
+	{Stations} {Using} {Free} and {Proprietary} {Network} {Data}.
+	\newblock \emph{Transportation Research Record: Journal of the Transportation
+		Research Board}, 2217:\penalty0 145--152, 2011.
+	\newblock ISSN 0361-1981.
+	\newblock \doi{10.3141/2217-18}.
+
+\end{thebibliography}
 

### 2017-12-14 draft/v2/latex/paper.tex
revision

--- 2017-12-14 draft/v1/latex/paper.tex
@@
 
-\usepackage{moreverb,url}
+\usepackage{moreverb}
 \usepackage[colorlinks,bookmarksopen,bookmarksnumbered,citecolor=red,urlcolor=red]{hyperref}
@@
 OpenStreetMap offers an underexplored source of worldwide geospatial data useful to urban researchers.
-This study uses the OSMnx software to automatically download and analyze 27,009 US street networks from OpenStreetMap at metropolitan, municipal, and neighborhood scales---namely, every US city and town ($N=19655$), census urbanized area ($N=497$), and Zillow-defined neighborhood ($N=6857$).
+This study uses the OSMnx software to automatically download and analyze 27,000 US street networks from OpenStreetMap at metropolitan, municipal, and neighborhood scales---namely, every US city and town ($N=19655$), census urbanized area ($N=497$), and Zillow-defined neighborhood ($N=6857$).
 It presents wide-ranging empirical findings on US urban form and street network characteristics, emphasizing measures relevant to graph theory, urban design, and morphology such as structure, connectedness, density, centrality, and resilience.
@@
 This study addresses these limitations by conducting a morphological analysis of urban street networks at multiple scales, with large sample sizes, with clearly defined network definitions and extents for reproducibility, and using nonplanar, directed graphs.
-In particular, it examines 27,009 urban street networks---represented as primal, nonplanar, weighted multidigraphs with possible self-loops---at multiple overlapping scales across the US, focusing on structure, connectedness, centrality, and resilience.
+In particular, it examines 27,000 urban street networks---represented as primal, nonplanar, weighted multidigraphs with possible self-loops---at multiple overlapping scales across the US, focusing on structure, connectedness, centrality, and resilience.
 It examines the street networks of every incorporated city and town (population size $N=19655$), census urbanized area ($N=497$), and Zillow-defined neighborhood ($N=6857$) in the US.
@@
 These density measures suggest how fine-grained the network is.
-Finally, \emph{average circuity} represents the average ratio between each edge length and the straight-line distance between the two nodes it links.
+Finally, \emph{average circuity} represents the average ratio between each edge length and the straight-line distance between the two nodes it links (cf.
+[redacted]).
 
@@
 
-Table \ref{table01} presents summary statistics for the entire data set of 497 urbanized areas.
+Table \ref{tab:measures_urban_areas} presents summary statistics for the entire data set of 497 urbanized areas.
 These urbanized areas span a wide range of sizes, from the Delano, CA Urbanized Area's 26 km\textsuperscript{2} to the New York--Newark, NY--NJ--CT Urbanized Area's 8,937 km\textsuperscript{2}---thus, density and count-based measures demonstrate substantial variance.
 Further, these urbanized areas span a wide spectrum of terrains, development eras and paradigms, and cultures.
+
+\begin{table*}
+\centering
+\caption{Central tendency and statistical dispersion for selected measures of all US urbanized areas' street networks: $\mu$ is the mean, $\sigma$ is the standard deviation, and $D$ is the dispersion index $\frac{\sigma ^ 2}{\mu}$.}
+\label{tab:measures_urban_areas}
+\begin{tabular}{ l r r r r r r }
+	\toprule
+	measure                                          & $\mu$          & $\sigma$       & min            & median         & max            & $D$            \\
+	\midrule
+	Area (km\textsuperscript{2})                     & 460.657        & 858.125        & 25.685         & 184.898        & 8937.429       & 1598.539       \\
+	Avg of the avg neighborhood degree               & 2.886          & 0.109          & 2.626          & 2.875          & 3.228          & 0.004          \\
+	Avg of the avg weighted neighborhood degree      & 0.032          & 0.018          & 0.021          & 0.030          & 0.321          & 0.011          \\
+	Avg circuity                                     & 1.076          & 0.019          & 1.023          & 1.074          & 1.140          & \textless0.001 \\
+	Avg clustering coefficient                       & 0.042          & 0.009          & 0.015          & 0.042          & 0.071          & 0.002          \\
+	Avg weighted clustering coefficient              & 0.002          & 0.001          & \textless0.001 & 0.001          & 0.006          & \textless0.001 \\
+	Intersection count                               & 12582          & 26054          & 751            & 4593           & 307848         & 53949.814      \\
+	Avg degree centrality                            & 0.001          & 0.001          & \textless0.001 & 0.001          & 0.007          & 0.001          \\
+	Edge density (km/km\textsuperscript{2})          & 13.455         & 2.137          & 7.961          & 13.352         & 21.233         & 0.340          \\
+	Avg edge length (m)                              & 158.588        & 17.653         & 117.341        & 157.332        & 223.080        & 1.965          \\
+	Total edge length (km)                           & 6353           & 12625          & 427            & 2393           & 1.42e8         & 25089.459      \\
+	Proportion of dead-ends                          & 0.213          & 0.055          & 0.077          & 0.207          & 0.416          & 0.014          \\
+	Proportion of 3-way intersections                & 0.593          & 0.046          & 0.444          & 0.591          & 0.778          & 0.004          \\
+	Proportion of 4-way intersections                & 0.187          & 0.063          & 0.054          & 0.178          & 0.422          & 0.021          \\
+	Intersection density (per km\textsuperscript{2}) & 26.469         & 6.256          & 12.469         & 26.029         & 49.423         & 1.478          \\
+	Average node degree                              & 5.153          & 0.302          & 4.307          & 5.143          & 6.056          & 0.018          \\
+	$m$                                              & 40890          & 83678          & 2516           & 14955          & 981646         & 171238.406     \\
+	$n$                                              & 16032          & 32585          & 874            & 5830           & 373309         & 66229.939      \\
+	Node density (per km\textsuperscript{2})         & 33.628         & 7.641          & 17.675         & 33.071         & 61.655         & 1.736          \\
+	Max PageRank value                               & 0.001          & 0.001          & \textless0.001 & 0.001          & 0.003          & \textless0.001 \\
+	Min PageRank value                               & \textless0.001 & \textless0.001 & \textless0.001 & \textless0.001 & \textless0.001 & \textless0.001 \\
+	Self-loop proportion                             & 0.008          & 0.008          & \textless0.001 & 0.006          & 0.071          & 0.008          \\
+	Street density (km/km\textsuperscript{2})        & 7.262          & 1.221          & 4.217          & 7.171          & 11.797         & 0.205          \\
+	Average street segment length (m)                & 161.331        & 17.765         & 119.573        & 160.288        & 225.920        & 1.956          \\
+	Total street length (km)                         & 3480           & 7026           & 222            & 1269           & 79046          & 14185.880      \\
+	Street segment count                             & 22011          & 45725          & 1281           & 7868           & 533757         & 94987.570      \\
+	Average streets per node                         & 2.764          & 0.162          & 2.223          & 2.770          & 3.217          & 0.010          \\
+	\bottomrule
+\end{tabular}
+\end{table*}
 
@@
 
-\begin{table*}
-\centering
-\caption{Central tendency and statistical dispersion for selected measures of all US urbanized areas' street networks.}
-\label{table01}
-\begin{tabular}{llllll}
-	\toprule
-	measure                                     & mean           & std dev        & min            & median         & max            \\
-	\midrule
-	Area (km\textsuperscript{2})                                  & 460.657        & 858.125        & 25.685         & 184.898        & 8937.429       \\
-	Avg of the avg neighborhood degree          & 2.886          & 0.109          & 2.626          & 2.875          & 3.228          \\
-	Avg of the avg weighted neighborhood degree & 0.032          & 0.018          & 0.021          & 0.03           & 0.321          \\
-	Avg circuity                                & 1.076          & 0.019          & 1.023          & 1.074          & 1.14           \\
-	Avg clustering coefficient                  & 0.042          & 0.009          & 0.015          & 0.042          & 0.071          \\
-	Avg weighted clustering coefficient         & 0.002          & 0.001          & \textless0.001 & 0.001          & 0.006          \\
-	Intersection count                          & 12582          & 26054          & 751            & 4593           & 307848         \\
-	Avg degree centrality                       & 0.001          & 0.001          & \textless0.001 & 0.001          & 0.007          \\
-	Edge density (km/km\textsuperscript{2})                       & 13.455         & 2.137          & 7.961          & 13.352         & 21.233         \\
-	Avg edge length (m)                         & 158.588        & 17.653         & 117.341        & 157.332        & 223.08         \\
-	Total edge length (km)                      & 6353           & 12625          & 427            & 2393           & 1.42E+08       \\
-	Proportion of dead-ends                     & 0.213          & 0.055          & 0.077          & 0.207          & 0.416          \\
-	Proportion of 3-way intersections           & 0.593          & 0.046          & 0.444          & 0.591          & 0.778          \\
-	Proportion of 4-way intersections           & 0.187          & 0.063          & 0.054          & 0.178          & 0.422          \\
-	Intersection density (per km\textsuperscript{2})              & 26.469         & 6.256          & 12.469         & 26.029         & 49.423         \\
-	Average node degree                         & 5.153          & 0.302          & 4.307          & 5.143          & 6.056          \\
-	m                                           & 40890          & 83678          & 2516           & 14955          & 981646         \\
-	n                                           & 16032          & 32585          & 874            & 5830           & 373309         \\
-	Node density (per km\textsuperscript{2})                      & 33.628         & 7.641          & 17.675         & 33.071         & 61.655         \\
-	Max PageRank value                          & 0.001          & 0.001          & \textless0.001 & 0.001          & 0.003          \\
-	Min PageRank value                          & \textless0.001 & \textless0.001 & \textless0.001 & \textless0.001 & \textless0.001 \\
-	Self-loop proportion                        & 0.008          & 0.008          & \textless0.001 & 0.006          & 0.071          \\
-	Street density (km/km\textsuperscript{2})                     & 7.262          & 1.221          & 4.217          & 7.171          & 11.797         \\
-	Average street segment length (m)           & 161.331        & 17.765         & 119.573        & 160.288        & 225.92         \\
-	Total street length (km)                    & 3480           & 7026           & 222            & 1269           & 79046          \\
-	Street segment count                        & 22011          & 45725          & 1281           & 7868           & 533757         \\
-	Average streets per node                    & 2.764          & 0.162          & 2.223          & 2.77           & 3.217          \\
-	\bottomrule
-\end{tabular}
-\end{table*}
-
 The typical urbanized area has 26 intersections per km\textsuperscript{2}.
@@
 However, New Orleans is an anomaly in the Deep South.
-Figure \ref{fig01} depicts the intersection density of each American urbanized area: the highest intersection densities concentrate west of the Mississippi River, while the lowest concentrate in a belt running from Louisiana, through the Carolinas and Appalachians, and into New England.
+Figure \ref{fig:fig01} depicts the intersection density of each American urbanized area: the highest intersection densities concentrate west of the Mississippi River, while the lowest concentrate in a belt running from Louisiana, through the Carolinas and Appalachians, and into New England.
 In general, only the largest cities on the east coast (e.g., Boston, New York, Philadelphia, Washington) and Florida escape this trend.
@@
 	\caption{Intersection density per urbanized area (left) and average streets per node per urbanized area (right) from lowest (pale yellow) to highest (dark red), in the contiguous US}
-	\label{fig01}
+	\label{fig:fig01}
 \end{figure*}
@@
 	\caption{Distribution of node types in 9 urbanized areas, with number of streets emanating from the node on the $x$-axis and proportion of nodes of this type on the $y$-axis }
-	\label{fig02}
+	\label{fig:fig02}
 \end{figure*}
@@
 The gridlike San Angelo, TX urbanized area has the most streets per node (3.2) on average, and (outside of Puerto Rico, which contains the seven lowest urbanized areas) the sprawling, disconnected Lexington Park, MD urbanized area has the fewest (2.2).
-These fit the trend seen in the spatial distribution across the US in Figure \ref{fig01}: urbanized areas in the Great Plains and Midwest have particularly high numbers of streets per node on average, indicating more gridlike, connected networks.
+These fit the trend seen in the spatial distribution across the US in Figure \ref{fig:fig01}: urbanized areas in the Great Plains and Midwest have particularly high numbers of streets per node on average, indicating more gridlike, connected networks.
 Cities in the southern and western US tend to have fewer streets per node, reflecting more dead-ends and a disconnected network.
@@
 However, this distribution varies somewhat: examining a small sample of nine urbanized areas, chosen to maximize variance, reveals this in clearer detail.
-In Figure \ref{fig02}, urban Atlanta and Chattanooga have very high proportions of dead-ends---each over 30\% of all nodes---and very few 4-way intersections, indicating a disconnected street pattern.
+In Figure \ref{fig:fig02}, urban Atlanta and Chattanooga have very high proportions of dead-ends---each over 30\% of all nodes---and very few 4-way intersections, indicating a disconnected street pattern.
 The urbanized areas of Phoenix, Boston, Detroit, and Chattanooga have particularly high proportions of 3-way intersections, each over 60\%, indicating a prevalence of T-intersections.
 Conversely, Chicago, New Orleans, Duluth, and Lubbock have high proportions of 4-way intersections, indicating more gridlike connected networks.
-But what is perhaps most notable about Figure \ref{fig02} is that these nine urbanized areas, despite being chosen to maximize variance, are overwhelmingly similar to each other.
+But what is perhaps most notable about Figure \ref{fig:fig02} is that these nine urbanized areas, despite being chosen to maximize variance, are overwhelmingly similar to each other.
 Every large American urban agglomeration is characterized by a preponderance of 3-way intersections.
@@
 \caption{Selected measures of the 30 largest (by land area) urbanized areas' street networks.}
-\label{table02}
+\label{tab:largest_urbanized_areas}
 \begin{tabular}{p{2.5cm} p{1.5cm} p{1.5cm} p{1.5cm} p{1.55cm} p{1.5cm} p{1.5cm} p{1.5cm} p{1.5cm} p{1.5cm} p{1.5cm}}
@@
 
+The dispersion index $D$ in Table \ref{tab:measures_urban_areas} demonstrates the heterogeneity of each indicator across the data set.
+Several \enquote{families} of indicators can be discerned by their heterogeneity.
+For instance, counts and totals such as $m$, $n$, and the total street length are extremely heterogeneous.
+Densities and average distances such as intersection density and the average street segment length exhibit only moderate heterogeneity.
+Finally, several topological measures such as the average clustering coefficient and PageRank are extremely homogeneous.
 Due to the substantial variation in urbanized area size, from 25 to 9,000 km\textsuperscript{2}, the preceding analysis covers a wide swath of metropolitan types.
-To better compare apples-to-apples, Table \ref{table02} focuses on the 30 largest urbanized areas cross-sectionally to examine their metric and topological measures.
+To better compare apples-to-apples, Table \ref{tab:largest_urbanized_areas} focuses on the 30 largest urbanized areas cross-sectionally to examine their metric and topological measures.
 This provides more consistent spatial scales and extents, while offering a window into the similarities and differences in the built forms of America's largest agglomerations.
@@
 
-Table \ref{table03} presents summary statistics of street network characteristics across the entire data set of 19,655 cities and towns---every incorporated city and town in the US.
-Following recent work by \citet{barthelemy_modeling_2008} and \citet{strano_urban_2013}, we examine the relationship between the total street length $L$ and the number of nodes $n$.
-The former proposed a model of cities in which $L$ and $n$ scale nonlinearly as $n^{1/2}$, and the latter recently confirmed this empirically with a small sample of ten European cities.
-However, their small sample size may limit the generalizability of this finding.
-To investigate this empirically, we examine the relationship between $L$ and $n$ for 19,655 US cities and towns and instead find a strong \emph{linear} relationship ($r^{2}=0.98$), as depicted in Figure \ref{fig03}.
+Table \ref{tab:measures_cities} presents summary statistics of street network characteristics across the entire data set of 19,655 cities and towns---every incorporated city and town in the US.
+Following recent work by \citet{barthelemy_modeling_2008} and \citet{strano_urban_2013}, we examine the relationship between the total street length $L$ and the number of nodes $n$ across different cities.
+The former proposed a model of city network evolution in which $L$ and $n$ scale nonlinearly as $n^{1/2}$, and the latter suggested that this relationship applies cross-sectionally, using a small empirical sample of ten European cities.
+However, the latter's small sample size may limit the generalizability of this finding.
+To investigate this empirically, we examine the relationship between $L$ and $n$ for 19,655 US cities and towns and instead find a strong \emph{linear} relationship ($r^{2}=0.98$), as depicted in Figure \ref{fig:fig03}.
 We also find a similar linear relationship at the metropolitan and neighborhood scales.
@@
 \centering
-\caption{Selected summary stats for every incorporated city and town in the US.}
-\label{table03}
-\begin{tabular}{llllll}
+\caption{Central tendency and statistical dispersion for selected measures of all incorporated cities and towns in the US: $\mu$ is the mean, $\sigma$ is the standard deviation, and $D$ is the dispersion index $\frac{\sigma ^ 2}{\mu}$.}
+\label{tab:measures_cities}
+\begin{tabular}{ l r r r r r r }
 	\toprule
-	measure                                     & mean    & std dev & min            & median         & max       \\
+	measure                                          & $\mu$   & $\sigma$& min            & median         & max       & $D$      \\
 	\midrule
-	Area (km\textsuperscript{2})                                  & 16.703  & 107.499 & 0.039          & 3.918          & 7434.258  \\
-	Avg of the avg neighborhood degree          & 2.94    & 0.297   & 0.4            & 2.953          & 3.735     \\
-	Avg of the avg weighted neighborhood degree & 0.033   & 0.141   & \textless0.001 & 0.029          & 9.357     \\
-	Avg circuity                                & 1.067   & 0.159   & 1              & 1.055          & 20.452    \\
-	Avg clustering coefficient                  & 0.048   & 0.041   & \textless0.001 & 0.04           & 1         \\
-	Avg weighted clustering coefficient         & 0.01    & 0.018   & \textless0.001 & 0.005          & 0.524     \\
-	Intersection count                          & 324     & 1266    & 0              & 83             & 62996     \\
-	Avg degree centrality                       & 0.093   & 0.136   & \textless0.001 & 0.052          & 2.667     \\
-	Edge density (km/km\textsuperscript{2})                       & 12.654  & 6.705   & 0.006          & 11.814         & 58.603    \\
-	Avg edge length (m)                         & 161.184 & 80.769  & 25.822         & 144.447        & 3036.957  \\
-	Total edge length (km)                      & 159.067 & 578.521 & 0.052          & 40.986         & 24728.326 \\
-	Proportion of dead-ends                     & 0.192   & 0.093   & \textless0.001 & 0.184          & 1         \\
-	Proportion of 3-way intersections           & 0.572   & 0.11    & \textless0.001 & 0.579          & 1         \\
-	Proportion of 4-way intersections           & 0.237   & 0.129   & \textless0.001 & 0.217          & 1         \\
-	Intersection density (per km\textsuperscript{2})              & 29.363  & 21.607  & \textless0.001 & 24.719         & 259.647   \\
-	Average node degree                         & 5.251   & 0.668   & 0.8            & 5.268          & 7.166     \\
-	m                                           & 1046    & 3924    & 2              & 275            & 176161    \\
-	n                                           & 401     & 1516    & 2              & 103            & 71993     \\
-	Node density (per km\textsuperscript{2})                      & 35.449  & 24.409  & 0.047          & 30.718         & 296.74    \\
-	Max PageRank value                          & 0.034   & 0.046   & \textless0.001 & 0.021          & 0.87      \\
-	Min PageRank value                          & 0.005   & 0.018   & \textless0.001 & 0.002          & 0.5       \\
-	Self-loop proportion                        & 0.005   & 0.015   & \textless0.001 & \textless0.001 & 1         \\
-	Street density (km/km\textsuperscript{2})                     & 6.528   & 3.435   & 0.003          & 6.109          & 29.302    \\
-	Average street segment length (m)           & 162.408 & 81.035  & 25.822         & 145.479        & 3036.957  \\
-	Total street length (km)                    & 86.096  & 331.048 & 0.026          & 21.005         & 15348.008 \\
-	Street segment count                        & 558     & 2208    & 1              & 140            & 107393    \\
-	Average streets per node                    & 2.851   & 0.282   & 1              & 2.852          & 4         \\
+	Area (km\textsuperscript{2})                     & 16.703  & 107.499 & 0.039          & 3.918          & 7434.258  & 691.860  \\
+	Avg of the avg neighborhood degree               & 2.940   & 0.297   & 0.400          & 2.953          & 3.735     & 0.030    \\
+	Avg of the avg weighted neighborhood degree      & 0.033   & 0.141   & \textless0.001 & 0.029          & 9.357     & 0.607    \\
+	Avg circuity                                     & 1.067   & 0.159   & 1.000          & 1.055          & 20.452    & 0.024    \\
+	Avg clustering coefficient                       & 0.048   & 0.041   & \textless0.001 & 0.04           & 1.000     & 0.035    \\
+	Avg weighted clustering coefficient              & 0.010   & 0.018   & \textless0.001 & 0.005          & 0.524     & 0.033    \\
+	Intersection count                               & 324     & 1266    & 0              & 83             & 62996     & 4951.293 \\
+	Avg degree centrality                            & 0.093   & 0.136   & \textless0.001 & 0.052          & 2.667     & 0.199    \\
+	Edge density (km/km\textsuperscript{2})          & 12.654  & 6.705   & 0.006          & 11.814         & 58.603    & 3.553    \\
+	Avg edge length (m)                              & 161.184 & 80.769  & 25.822         & 144.447        & 3036.957  & 40.473   \\
+	Total edge length (km)                           & 159.067 & 578.521 & 0.052          & 40.986         & 24728.326 & 2104.061 \\
+	Proportion of dead-ends                          & 0.192   & 0.093   & \textless0.001 & 0.184          & 1.000     & 0.045    \\
+	Proportion of 3-way intersections                & 0.572   & 0.11    & \textless0.001 & 0.579          & 1.000     & 0.021    \\
+	Proportion of 4-way intersections                & 0.237   & 0.129   & \textless0.001 & 0.217          & 1.000     & 0.070    \\
+	Intersection density (per km\textsuperscript{2}) & 29.363  & 21.607  & \textless0.001 & 24.719         & 259.647   & 15.900   \\
+	Average node degree                              & 5.251   & 0.668   & 0.800          & 5.268          & 7.166     & 0.085    \\
+	$m$                                              & 1046    & 3924    & 2              & 275            & 176161    & 14714.556\\
+	$n$                                              & 401     & 1516    & 2              & 103            & 71993     & 5734.363 \\
+	Node density (per km\textsuperscript{2})         & 35.449  & 24.409  & 0.047          & 30.718         & 296.740   & 16.807   \\
+	Max PageRank value                               & 0.034   & 0.046   & \textless0.001 & 0.021          & 0.870     & 0.062    \\
+	Min PageRank value                               & 0.005   & 0.018   & \textless0.001 & 0.002          & 0.500     & 0.060    \\
+	Self-loop proportion                             & 0.005   & 0.015   & \textless0.001 & \textless0.001 & 1.000     & 0.042    \\
+	Street density (km/km\textsuperscript{2})        & 6.528   & 3.435   & 0.003          & 6.109          & 29.302    & 1.807    \\
+	Average street segment length (m)                & 162.408 & 81.035  & 25.822         & 145.479        & 3036.957  & 40.433   \\
+	Total street length (km)                         & 86.096  & 331.048 & 0.026          & 21.005         & 15348.008 & 1272.917 \\
+	Street segment count                             & 558     & 2208    & 1.000          & 140            & 107393    & 8745.983 \\
+	Average streets per node                         & 2.851   & 0.282   & 1.000          & 2.852          & 4.000     & 0.028    \\
 	\bottomrule
@@
 Previous findings \citep[e.g.,][]{masucci_random_2009} also suggest street segment lengths in an urban network follow a power-law distribution.
-However, we find that these networks instead generally follow a lognormal distribution.
+However, we find that these networks instead generally follow lognormal-style right-skewed distributions.
 This makes theoretical sense as most street networks are not truly scale-free: for example, a typical street network might comprise very few very long street segments (e.g., 1 km), more medium-length segments (e.g., 250 m), many short segments (e.g., 80 m), but very few very short segments (e.g., 10 m).
-This theoretical illustration suggests the lognormal distribution that this study generally finds across municipal US street networks.
-An exception, of course, lies in consistently-sized orthogonal grids filling a city's incorporated spatial extents.
+To test this in a statistically rigorous way, we fit a set of candidate distributions to the street segment lengths of each US city/town.
+The candidate distributions comprise the lognormal, Gumbel, gamma, exponentiated Weibull, Fréchet, power-law, uniform, and exponential distributions.
+We then assess each fit with a Kolmogorov-Smirnov test, which tells us how closely each fitted distribution matches the observed data.
+Power-law distributions provide the best fit for fewer than 0.7\% of these cities.
+In contrast, the exponentiated Weibull distribution provides the best fit 56\% of the time, followed by the Gumbel (19\%), gamma (11\%), and lognormal (7\%) distributions.
+
+An exception to this general pattern, of course, lies in consistently-sized orthogonal grids filling a city's incorporated spatial extents.
 Such distributions are extremely peaked around a single value: the linear length of a grid block.
-
 We find that such cities are not uncommon in the US, particularly between the Mississippi River and the Rocky Mountains: the Great Plains states are characterized by a unique street network form that is both orthogonal and reasonably dense.
@@
 \begin{figure*}
-	\includegraphics[width=0.7\textwidth]{fig03.png}
+	\includegraphics[width=\textwidth]{fig03.png}
 	\caption{The linear relationship between total street length and number of nodes in the street networks of 6,857 US neighborhoods (left) and 19,655 US cities and towns (right).}
-	\label{fig03}
+	\label{fig:fig03}
 \end{figure*}
@@
 \begin{figure*}
-	\includegraphics[width=0.85\textwidth]{fig04.png}
+	\includegraphics[width=\textwidth]{fig04.png}
 	\caption{Left: contiguous US states by median of mean streets per node in municipal street networks, from lowest (pale yellow) to highest (dark red).
 Right: contiguous US states by median of mean street segment length in municipal street networks, from longest (pale yellow) to shortest (dark red).}
-	\label{fig04}
+	\label{fig:fig04}
 \end{figure*}
@@
 \caption{Median values, aggregated by state plus DC, of selected measures of the municipal-scale street networks for every city and town in the US.}
-\label{table04}
-\begin{tabular}{p{1in}p{1in}p{1in}p{1in}p{1in}}
+\label{tab:medians_state_aggregate}
+\begin{tabular}{ l r r r r }
 	\toprule
@@
 	\midrule
-	AK    & 1.28                        & 2.43                 & 1.1          & 223.5                     \\
-	AL    & 9.7                         & 2.64                 & 1.07         & 190.81                    \\
+	AK    &  1.28                       & 2.43                 & 1.10         & 223.50                    \\
+	AL    &  9.70                       & 2.64                 & 1.07         & 190.81                    \\
 	AR    & 15.75                       & 2.78                 & 1.06         & 166.32                    \\
-	AZ    & 12.45                       & 2.77                 & 1.08         & 171.8                     \\
+	AZ    & 12.45                       & 2.77                 & 1.08         & 171.80                    \\
 	CA    & 32.58                       & 2.74                 & 1.07         & 143.79                    \\
 	CO    & 29.26                       & 2.88                 & 1.06         & 136.68                    \\
-	CT    & 28.05                       & 2.7                  & 1.07         & 165.87                    \\
+	CT    & 28.05                       & 2.70                 & 1.07         & 165.87                    \\
 	DC    & 58.91                       & 3.26                 & 1.04         & 122.23                    \\
-	DE    & 25.3                        & 2.8                  & 1.06         & 127.8                     \\
+	DE    & 25.30                       & 2.80                 & 1.06         & 127.80                    \\
 	FL    & 26.26                       & 2.87                 & 1.07         & 150.75                    \\
-	GA    & 15.25                       & 2.78                 & 1.07         & 177.5                     \\
-	HI    & 8                           & 2.42                 & 1.07         & 177.93                    \\
+	GA    & 15.25                       & 2.78                 & 1.07         & 177.50                    \\
+	HI    &  8.00                       & 2.42                 & 1.07         & 177.93                    \\
 	IA    & 24.08                       & 3.02                 & 1.04         & 129.36                    \\
@@
 	ME    &  7.69                       & 2.67                 & 1.07         & 198.93                    \\
-	MI    & 20.93                       & 2.9                  & 1.05         & 153.5                     \\
+	MI    & 20.93                       & 2.90                 & 1.05         & 153.50                    \\
 	MN    & 18.96                       & 2.87                 & 1.06         & 152.92                    \\
@@
 	NE    & 45.89                       & 3.16                 & 1.04         & 119.79                    \\
-	NH    & 12.22                       & 2.69                 & 1.1          & 175.88                    \\
+	NH    & 12.22                       & 2.69                 & 1.10         & 175.88                    \\
 	NJ    & 44.98                       & 2.88                 & 1.04         & 130.79                    \\
-	NM    & 18.5                        & 2.93                 & 1.05         & 152.02                    \\
+	NM    & 18.50                       & 2.93                 & 1.05         & 152.02                    \\
 	NV    & 13.86                       & 2.77                 & 1.07         & 147.35                    \\
 	NY    & 21.89                       & 2.75                 & 1.06         & 156.88                    \\
-	OH    & 25.23                       & 2.8                  & 1.05         & 142.08                    \\
-	OK    & 28.22                       & 3.03                 & 1.05         & 139.5                     \\
+	OH    & 25.23                       & 2.80                 & 1.05         & 142.08                    \\
+	OK    & 28.22                       & 3.03                 & 1.05         & 139.50                    \\
 	OR    & 35.08                       & 2.69                 & 1.06         & 121.18                    \\
@@
 
-For example, if we measure connectedness in terms of average number of streets per node at the city-scale and then aggregate these cities by state (Table \ref{table04}), we find Nebraska, Kansas, South Dakota, Montana, North Dakota, Oklahoma, and Iowa have, in order, the highest medians (Figure \ref{fig04}).
+For example, if we measure connectedness in terms of average number of streets per node at the city-scale and then aggregate these cities by state (Table \ref{tab:medians_state_aggregate}), we find Nebraska, Kansas, South Dakota, Montana, North Dakota, Oklahoma, and Iowa have, in order, the highest medians (Figure \ref{fig:fig04}).
 This indicates the most gridlike networks.
@@
 We again see three Great Plains states near the top, alongside small densely populated East Coast states.
-Nebraska also has the smallest block sizes (measured via the proxy of average street segment length) while the largest concentrate in the Deep South, upper New England, and Utah (Figure \ref{fig04}).
+Nebraska also has the smallest block sizes (measured via the proxy of average street segment length) while the largest concentrate in the Deep South, upper New England, and Utah (Figure \ref{fig:fig04}).
 
@@
 \centering
-\caption{Selected summary stats for all the neighborhood-scale street networks.}
-\label{table05}
-\begin{tabular}{llllll}
+\caption{Central tendency and statistical dispersion for selected measures of all the neighborhood-scale street networks: $\mu$ is the mean, $\sigma$ is the standard deviation, and $D$ is the dispersion index $\frac{\sigma ^ 2}{\mu}$.}
+\label{tab:measures_neighborhoods}
+\begin{tabular}{ l r r r r r r }
 	\toprule
-	& mean    & std dev       & min            & median         & max      \\
+	measure                                          & $\mu$   & $\sigma$& min            & median         & max      & $D$     \\
 	\midrule
-	Area (km\textsuperscript{2})                                  & 5.322   & 15.463  & 0.008          & 1.738          & 323.306  \\
-	Avg of the avg neighborhood degree          & 2.598   & 0.436   & \textless0.001 & 2.67           & 3.632    \\
-	Avg of the avg weighted neighborhood degree & 0.031   & 0.041   & \textless0.001 & 0.029          & 2.991    \\
-	Avg circuity                                & 1.08    & 0.411   & 1              & 1.044          & 24.29    \\
-	Avg clustering coefficient                  & 0.044   & 0.055   & \textless0.001 & 0.034          & 1        \\
-	Avg weighted clustering coefficient         & 0.01    & 0.027   & \textless0.001 & 0.005          & 0.799    \\
-	Intersection count                          & 173     & 379     & 0              & 76             & 8371     \\
-	Avg degree centrality                       & 0.13    & 0.27    & 0.001          & 0.054          & 4        \\
-	Edge density (km/km\textsuperscript{2})                       & 17.569  & 7.095   & 0.025          & 18.152         & 59.939   \\
-	Avg edge length (m)                         & 142.279 & 59.182  & 8.447          & 133.848        & 2231.331 \\
-	Total edge length (km)                      & 71.369  & 166.566 & 0.017          & 29.88          & 3563.409 \\
-	Proportion of dead-ends                     & 0.17    & 0.131   & \textless0.001 & 0.145          & 1        \\
-	Proportion of 3-way intersections           & 0.559   & 0.146   & \textless0.001 & 0.574          & 1        \\
-	Proportion of 4-way intersections           & 0.275   & 0.176   & \textless0.001 & 0.234          & 1        \\
-	Intersection density (per km\textsuperscript{2})              & 49.497  & 28.33   & \textless0.001 & 46.43          & 444.355  \\
-	Average node degree                         & 4.675   & 0.836   & 0.545          & 4.736          & 7.283    \\
-	m                                           & 5201    & 1185    & 1              & 217            & 27289    \\
-	n                                           & 208     & 459     & 2              & 90             & 9327     \\
-	Node density (per km\textsuperscript{2})                      & 58.677  & 31.802  & 0.063          & 55.626         & 499.9    \\
-	Max PageRank value                          & 0.055   & 0.086   & \textless0.001 & 0.026          & 0.889    \\
-	Min PageRank value                          & 0.01    & 0.041   & \textless0.001 & 0.002          & 0.5      \\
-	Self-loop proportion                        & 0.007   & 0.034   & \textless0.001 & \textless0.001 & 1        \\
-	Street density (km/km\textsuperscript{2})                     & 9.744   & 4.085   & 0.013          & 9.882          & 33.737   \\
-	Average street segment length (m)           & 143.664 & 60.023  & 7.376          & 134.877        & 2231.331 \\
-	Total street length (km)                    & 40.049  & 93.987  & 0.009          & 16.248         & 1960.643 \\
-	Street segment count                        & 288     & 656     & 1              & 119            & 14754    \\
-	Average streets per node                    & 2.925   & 0.408   & 1              & 2.944          & 4.026    \\
+	Area (km\textsuperscript{2})                     & 5.322   & 15.463  & 0.008          & 1.738          & 323.306  & 44.928  \\
+	Avg of the avg neighborhood degree               & 2.598   & 0.436   & \textless0.001 & 2.670          & 3.632    & 0.073   \\
+	Avg of the avg weighted neighborhood degree      & 0.031   & 0.041   & \textless0.001 & 0.029          & 2.991    & 0.054   \\
+	Avg circuity                                     & 1.080   & 0.411   & 1.000          & 1.044          & 24.290   & 0.157   \\
+	Avg clustering coefficient                       & 0.044   & 0.055   & \textless0.001 & 0.034          & 1.000    & 0.069   \\
+	Avg weighted clustering coefficient              & 0.010   & 0.027   & \textless0.001 & 0.005          & 0.799    & 0.076   \\
+	Intersection count                               & 173     & 379     & 0              & 76             & 8371     & 829.528 \\
+	Avg degree centrality                            & 0.130   & 0.270   & 0.001          & 0.054          & 4.000    & 0.561   \\
+	Edge density (km/km\textsuperscript{2})          & 17.569  & 7.095   & 0.025          & 18.152         & 59.939   & 2.866   \\
+	Avg edge length (m)                              & 142.279 & 59.182  & 8.447          & 133.848        & 2231.331 & 24.617  \\
+	Total edge length (km)                           & 71.369  & 166.566 & 0.017          & 29.880         & 3563.409 & 388.743 \\
+	Proportion of dead-ends                          & 0.170   & 0.131   & \textless0.001 & 0.145          & 1.000    & 0.101   \\
+	Proportion of 3-way intersections                & 0.559   & 0.146   & \textless0.001 & 0.574          & 1.000    & 0.038   \\
+	Proportion of 4-way intersections                & 0.275   & 0.176   & \textless0.001 & 0.234          & 1.000    & 0.112   \\
+	Intersection density (per km\textsuperscript{2}) & 49.497  & 28.330  & \textless0.001 & 46.430         & 444.355  & 16.216  \\
+	Average node degree                              & 4.675   & 0.836   & 0.545          & 4.736          & 7.283    & 0.150   \\
+	$m$                                              & 5201    & 1185    & 1              & 217            & 27289    & 2694.171\\
+	$n$                                              & 208     & 459     & 2              & 90             & 9327     & 1014.643\\
+	Node density (per km\textsuperscript{2})         & 58.677  & 31.802  & 0.063          & 55.626         & 499.900  & 17.237  \\
+	Max PageRank value                               & 0.055   & 0.086   & \textless0.001 & 0.026          & 0.889    & 0.133   \\
+	Min PageRank value                               & 0.010   & 0.041   & \textless0.001 & 0.002          & 0.500    & 0.161   \\
+	Self-loop proportion                             & 0.007   & 0.034   & \textless0.001 & \textless0.001 & 1.000    & 0.177   \\
+	Street density (km/km\textsuperscript{2})        & 9.744   & 4.085   & 0.013          & 9.882          & 33.737   & 1.712   \\
+	Average street segment length (m)                & 143.664 & 60.023  & 7.376          & 134.877        & 2231.331 & 25.078  \\
+	Total street length (km)                         & 40.049  & 93.987  & 0.009          & 16.248         & 1960.643 & 220.569 \\
+	Street segment count                             & 288     & 656     & 1              & 119            & 14754    & 1491.595\\
+	Average streets per node                         & 2.925   & 0.408   & 1.000          & 2.944          & 4.026    & 0.057   \\
 	\bottomrule
@@
 
-Table \ref{table05} presents summary statistics for these 6,857 neighborhoods.
+Table \ref{tab:measures_neighborhoods} presents summary statistics for these 6,857 neighborhoods.
 Compared to the metropolitan and municipal scales, we see much greater variance here, as expected, given the smaller network sizes at the neighborhood scale.
@@
 Right: top, downtown San Francisco, CA; bottom, its suburb, Concord.}
-	\label{fig05}
+	\label{fig:fig05}
 \end{figure*}
 
-To illustrate these morphologies, Figure \ref{fig05} compares one square mile of the centers of Philadelphia, Portland, and San Francisco with one square mile of each of their suburbs.
+To illustrate these morphologies, Figure \ref{fig:fig05} compares one square mile of the centers of Philadelphia, Portland, and San Francisco with one square mile of each of their suburbs.
 The connectedness and fine-grain of the central cities are clear, as are the disconnectivity and coarse-grain of their suburbs.
 In fact, the suburbs have more in common with one another---despite being hundreds or thousands of miles apart---than they do with their central city neighbors, suggesting that land use and an era's prevailing design paradigm is paramount to geographical localism and regional context.
-The top row of Figure \ref{fig05} represents an era of planning and development that preceded the automobile, while the bottom row reflects the exclusionary zoning and mid--late twentieth century era of automobility in residential suburb design---namely the \enquote{loops and lollipops} and the \enquote{lollipops on a stick} design patterns \cite{southworth_streets_1997}.
+The top row of Figure \ref{fig:fig05} represents an era of planning and development that preceded the automobile, while the bottom row reflects the exclusionary zoning and mid--late twentieth century era of automobility in residential suburb design---namely the \enquote{loops and lollipops} and the \enquote{lollipops on a stick} design patterns \cite{southworth_streets_1997}.
 
@@
 This differs from previous findings in the literature that relied on smaller sample sizes and examined European instead of US cities.
-We also find that most networks typically follow a lognormal distribution of street segment lengths.
-As discussed, this finding seems to make sense theoretically and is supported by these large-sample data at multiple scales, but obvious exceptions to lognormal distribution lie in those networks that exhibit substantial uniformity.
+We also find that most networks typically follow lognormal-style right-skewed distributions (particularly the exponentiated Weibull distribution) of street segment lengths.
+As discussed, this finding seems to make sense theoretically and is supported by these large-sample data at multiple scales, but obvious exceptions exist in those networks that exhibit substantial uniformity.
 At the neighborhood scale, examples include downtowns with consistent orthogonal grids, such as that of Portland, Oregon.
@@
 
-Future research should incorporate temporal analyses that goes beyond the present study's cross-sectional data.
+Future research could incorporate temporal analyses that goes beyond the present study's cross-sectional data.
 This empirical analysis emphasized network structure, but further linking structural complexity to the temporal complexity of city dynamics and processes lies ahead as critical work.
-As OSMnx can automatically calculate several dozen street network measures, future work can use dimensionality reduction to identify significant baskets of factors and cluster places into types.
+As OSMnx can automatically calculate several dozen street network measures, future work can use dimensionality reduction to identify significant baskets of indicators and cluster places into morphological types.
 These variables can also be used as advanced urban form measures in hedonic regressions and accessibility studies.
+Finally, future research can further explore urban spatial geometries such as block shapes and configurations, the statistical distributions of various indicators, and the comparative character of worldwide cities: this analysis of US urbanism and its specific empirical findings do not necessarily apply universally to cities elsewhere in the world.
 
@@
 This paper had three primary purposes.
-First, it examined urban morphology through metric and topological measures of 27,009 US street networks---particularly focusing on density, connectedness, and resilience.
+First, it examined urban morphology through metric and topological analyses of 27,000 US street networks---particularly focusing on density, connectedness, and resilience.
 Second, these methods and findings demonstrate the use of OSMnx as a new street network research toolkit, suggesting to urban planners and scholars new methods for acquiring and analyzing data.
-Third, it investigated claims about the distribution of street segment lengths and the relationship between the total street length and the number of nodes in a network.
+Third, it investigated past findings about the distribution of street segment lengths and the relationship between the total street length and the number of nodes in a network.
 This study has made all of these network datasets---for 497 urbanized areas, 19,655 cities and towns, and 6,857 neighborhoods---along with all of their attribute data and morphological measures available in an online public repository for other researchers to study and repurpose at [redacted].

### 2017-12-16 draft/v2/latex/paper.tex
update street network measures descriptions

@@
 \cite{trudeau_introduction_1994, albert_statistical_2002, dorogovtsev_evolution_2002, brandes_network_2005, costa_characterization_2007, newman_structure_2003, newman_networks:_2010, barthelemy_spatial_2011, cranmer_navigating_2017}.
-\emph{Average street length}, the mean edge length in the undirected representation of the graph, serves as a linear proxy for block size and indicates how fine-grained or coarse-grained the network is.
-\emph{Node density} is the number of nodes divided by the area covered by the network.
-\emph{Intersection density} is the density of the set of nodes with more than one street emanating from them (thus excluding dead-ends).
-\emph{Edge density} is the sum of all edge lengths divided by area, and the physical \emph{street density} is the sum of all edges in the undirected representation of the graph divided by area.
-These density measures suggest how fine-grained the network is.
-Finally, \emph{average circuity} represents the average ratio between each edge length and the straight-line distance between the two nodes it links (cf.
+The \emph{average street segment length} is a linear proxy for block size and specifies the network's grain.
+\emph{Node density} divides the node count by the network's area, while \emph{intersection density} excludes dead-ends to represent the density of street junctions.
+\emph{Edge density} divides the total directed network length by area, while \emph{street density} does the same for an undirected representation of the graph (to not double-count bidirectional streets).
+\emph{Average circuity} measures the ratio of edge lengths to the great-circle distances between the nodes these edges connect, indicating the street pattern's curvilinearity (cf.
 [redacted]).
 
-The \emph{average node degree}, or mean number of edges incident to each node, quantifies how well the nodes are connected.
-Similarly, but more concretely, the \emph{average streets per node} measures the mean number of physical streets (i.e., edges in the undirected representation of the graph) that emanate from each intersection and dead-end.
-This adapts the average node degree for physical form rather than directed flow.
-The statistical and spatial distributions of number of streets per node characterize the type, prevalence, and dispersion of intersection connectedness in the network.
-\emph{Connectivity} measures the minimum number of nodes or edges that must be removed from a connected network to disconnect it.
-The \emph{average node connectivity} (ANC) of a network---the mean number of internally node-disjoint paths between each pair of nodes---represents the expected number of nodes that must be removed to disconnect a randomly selected pair of non-adjacent nodes \cite{beineke_average_2002,dankelmann_bounds_2003}.
-
-The \emph{clustering coefficient} of a node is the ratio of the number of edges between its neighbors to the maximum possible number of edges that could exist between these neighbors \cite{jiang_topological_2004, opsahl_clustering_2009}.
-The weighted clustering coefficient weights this ratio by edge length and the average clustering coefficient is the mean of the clustering coefficients of all the network's nodes.
-\emph{Betweenness centrality} indicates the importance of nodes in a network by evaluating the number of shortest paths that pass through each node \cite{barthelemy_betweenness_2004,huang_trajgraph:_2016, zhong_revealing_2017}.
-The maximum betweenness centrality (MBC) in a network specifies the proportion of shortest paths that pass through the most important node.
-This is an indicator of resilience: networks with a high MBC are more prone to failure or inefficiency should this single choke point fail.
+The network's \emph{average node degree} quantifies connectedness in terms of the average number of edges incident to its nodes.
+The \emph{average streets per node} adapts this for physical form rather than directed circulation.
+It measures the average number of physical streets that emanate from each node (i.e., intersection or dead-end).
+The distribution and proportion of streets per node characterize the type, pervasiveness, and spatial dispersal of network connectedness and dead-ends.
+Connectivity represents the fewest number of nodes or edges that will disconnect the network if they are removed and is thus an indicator of resilience.
+A network's \emph{average node connectivity} (ANC)---the mean number of internally node-disjoint paths between each pair of nodes---more usefully represents how many nodes must be removed on average to disconnect a randomly selected pair of nodes \cite{beineke_average_2002,dankelmann_bounds_2003}.
+Brittle points of vulnerability characterize networks with low average connectivity.
+
+A node's \emph{clustering coefficient} represents the ratio between its neighbors' links and the maximum number of links that could exist between them \cite{jiang_topological_2004, opsahl_clustering_2009}.
+The weighted clustering coefficient weights this by edge length and the average clustering coefficient is the mean of the clustering coefficients of all the nodes.
+\emph{Betweenness centrality} evaluates how many of the network's shortest paths pass through some node (or edge) to indicate its importance \cite{barthelemy_betweenness_2004,huang_trajgraph:_2016, zhong_revealing_2017}.
+A network's \emph{maximum betweenness centrality} (MBC) measures the share of shortest paths that pass through the network's most important node: higher maximum betweenness centralities suggest networks more prone to inefficiency if this important choke point should fail.
 Finally, \emph{PageRank} ranks nodes based on the structure of incoming links and the rank of the source node \cite{brin_anatomy_1998, jiang_predicting_2008, agryzkov_algorithm_2012, chin_geographically_2015, gleich_pagerank_2015}.
@@
 \hline
-New York             & 8937          & 1.06         & 0.04                       & 0.18           & 0.62        & 0.2         & 34.44                 & 8.84                  & 148                   & 2.86             \\
-Atlanta              & 6850          & 1.1          & 0.04                       & 0.32           & 0.58        & 0.09        & 18.39                 & 6.16                  & 186                   & 2.45             \\
+New York             & 8937          & 1.06         & 0.04                       & 0.18           & 0.62        & 0.20        & 34.44                 & 8.84                  & 148                   & 2.86             \\
+Atlanta              & 6850          & 1.10         & 0.04                       & 0.32           & 0.58        & 0.09        & 18.39                 & 6.16                  & 186                   & 2.45             \\
 Chicago              & 6325          & 1.07         & 0.04                       & 0.17           & 0.57        & 0.25        & 27.05                 & 7.77                  & 163                   & 2.92             \\
-Philadelphia         & 5132          & 1.08         & 0.05                       & 0.17           & 0.63        & 0.2         & 26.65                 & 7.3                   & 159                   & 2.87             \\
-Boston               & 4852          & 1.09         & 0.05                       & 0.2            & 0.68        & 0.11        & 24.23                 & 6.44                  & 154                   & 2.71             \\
+Philadelphia         & 5132          & 1.08         & 0.05                       & 0.17           & 0.63        & 0.20        & 26.65                 & 7.30                  & 159                   & 2.87             \\
+Boston               & 4852          & 1.09         & 0.05                       & 0.20           & 0.68        & 0.11        & 24.23                 & 6.44                  & 154                   & 2.71             \\
 Dallas               & 4612          & 1.07         & 0.05                       & 0.15           & 0.61        & 0.23        & 34.16                 & 9.16                  & 156                   & 2.95             \\
 Los Angeles          & 4497          & 1.06         & 0.03                       & 0.21           & 0.56        & 0.22        & 39.45                 & 10.59                 & 151                   & 2.82             \\
-Houston              & 4303          & 1.08         & 0.04                       & 0.2            & 0.57        & 0.22        & 33.49                 & 8.62                  & 145                   & 2.83             \\
-Detroit              & 3461          & 1.07         & 0.04                       & 0.15           & 0.63        & 0.22        & 31.1                  & 8.56                  & 159                   & 2.93             \\
+Houston              & 4303          & 1.08         & 0.04                       & 0.20           & 0.57        & 0.22        & 33.49                 & 8.62                  & 145                   & 2.83             \\
+Detroit              & 3461          & 1.07         & 0.04                       & 0.15           & 0.63        & 0.22        & 31.10                 & 8.56                  & 159                   & 2.93             \\
 Washington           & 3424          & 1.09         & 0.04                       & 0.26           & 0.56        & 0.17        & 31.22                 & 8.26                  & 146                   & 2.66             \\
-Miami                & 3204          & 1.1          & 0.05                       & 0.17           & 0.59        & 0.23        & 40.54                 & 10.61                 & 149                   & 2.89             \\
-Phoenix              & 2968          & 1.09         & 0.05                       & 0.2            & 0.62        & 0.17        & 35.31                 & 9.1                   & 150                   & 2.77             \\
+Miami                & 3204          & 1.10         & 0.05                       & 0.17           & 0.59        & 0.23        & 40.54                 & 10.61                 & 149                   & 2.89             \\
+Phoenix              & 2968          & 1.09         & 0.05                       & 0.20           & 0.62        & 0.17        & 35.31                 & 9.10                  & 150                   & 2.77             \\
 Minneapolis          & 2647          & 1.08         & 0.05                       & 0.19           & 0.57        & 0.23        & 29.54                 & 8.68                  & 167                   & 2.84             \\
-Seattle              & 2617          & 1.07         & 0.03                       & 0.3            & 0.54        & 0.16        & 31.57                 & 8.2                   & 143                   & 2.57             \\
-Tampa                & 2479          & 1.1          & 0.05                       & 0.2            & 0.58        & 0.21        & 31.35                 & 8.46                  & 153                   & 2.83             \\
+Seattle              & 2617          & 1.07         & 0.03                       & 0.30           & 0.54        & 0.16        & 31.57                 & 8.20                  & 143                   & 2.57             \\
+Tampa                & 2479          & 1.10         & 0.05                       & 0.20           & 0.58        & 0.21        & 31.35                 & 8.46                  & 153                   & 2.83             \\
 St.
-Louis            & 2392          & 1.1          & 0.04                       & 0.22           & 0.62        & 0.15        & 29.68                 & 8.16                  & 154                   & 2.73             \\
-Pittsburgh           & 2345          & 1.09         & 0.04                       & 0.23           & 0.6         & 0.16        & 23.57                 & 6.71                  & 165                   & 2.72             \\
+Louis            & 2392          & 1.10         & 0.04                       & 0.22           & 0.62        & 0.15        & 29.68                 & 8.16                  & 154                   & 2.73             \\
+Pittsburgh           & 2345          & 1.09         & 0.04                       & 0.23           & 0.60        & 0.16        & 23.57                 & 6.71                  & 165                   & 2.72             \\
 San Juan             & 2245          & 1.11         & 0.02                       & 0.36           & 0.56        & 0.08        & 26.57                 & 6.43                  & 131                   & 2.36             \\
-Cincinnati           & 2040          & 1.07         & 0.03                       & 0.31           & 0.54        & 0.14        & 17.96                 & 6.1                   & 186                   & 2.51             \\
+Cincinnati           & 2040          & 1.07         & 0.03                       & 0.31           & 0.54        & 0.14        & 17.96                 & 6.10                  & 186                   & 2.51             \\
 Cleveland            & 2004          & 1.07         & 0.04                       & 0.19           & 0.66        & 0.14        & 19.13                 & 6.51                  & 198                   & 2.76             \\
-Charlotte            & 1920          & 1.08         & 0.04                       & 0.3            & 0.57        & 0.11        & 21                    & 6.43                  & 170                   & 2.51             \\
+Charlotte            & 1920          & 1.08         & 0.04                       & 0.30           & 0.57        & 0.11        & 21.00                 & 6.43                  & 170                   & 2.51             \\
 San Diego            & 1897          & 1.08         & 0.03                       & 0.28           & 0.54        & 0.17        & 28.89                 & 8.32                  & 159                   & 2.62             \\
 Baltimore            & 1857          & 1.09         & 0.04                       & 0.23           & 0.59        & 0.17        & 27.72                 & 7.56                  & 152                   & 2.72             \\
-Indianapolis         & 1828          & 1.08         & 0.05                       & 0.23           & 0.59        & 0.17        & 27.62                 & 7.63                  & 157                   & 2.7              \\
-Kansas City          & 1756          & 1.06         & 0.04                       & 0.21           & 0.58        & 0.2         & 32.09                 & 8.57                  & 152                   & 2.79             \\
-Denver               & 1729          & 1.07         & 0.05                       & 0.2            & 0.57        & 0.22        & 40.6                  & 9.84                  & 138                   & 2.84             \\
-Orlando              & 1548          & 1.11         & 0.06                       & 0.2            & 0.61        & 0.18        & 26.3                  & 7.44                  & 163                   & 2.79             \\
-San Antonio          & 1547          & 1.07         & 0.05                       & 0.17           & 0.6         & 0.21        & 28.33                 & 7.91                  & 162                   & 2.87             \\
-Nashville            & 1460          & 1.08         & 0.03                       & 0.27           & 0.59        & 0.14        & 19.08                 & 6.1                   & 181                   & 2.6              \\
-Milwaukee            & 1413          & 1.06         & 0.06                       & 0.14           & 0.55        & 0.3         & 28.27                 & 7.81                  & 157                   & 3.03 \\
+Indianapolis         & 1828          & 1.08         & 0.05                       & 0.23           & 0.59        & 0.17        & 27.62                 & 7.63                  & 157                   & 2.70             \\
+Kansas City          & 1756          & 1.06         & 0.04                       & 0.21           & 0.58        & 0.20        & 32.09                 & 8.57                  & 152                   & 2.79             \\
+Denver               & 1729          & 1.07         & 0.05                       & 0.20           & 0.57        & 0.22        & 40.60                 & 9.84                  & 138                   & 2.84             \\
+Orlando              & 1548          & 1.11         & 0.06                       & 0.20           & 0.61        & 0.18        & 26.30                 & 7.44                  & 163                   & 2.79             \\
+San Antonio          & 1547          & 1.07         & 0.05                       & 0.17           & 0.60        & 0.21        & 28.33                 & 7.91                  & 162                   & 2.87             \\
+Nashville            & 1460          & 1.08         & 0.03                       & 0.27           & 0.59        & 0.14        & 19.08                 & 6.10                  & 181                   & 2.60             \\
+Milwaukee            & 1413          & 1.06         & 0.06                       & 0.14           & 0.55        & 0.30        & 28.27                 & 7.81                  & 157                   & 3.03 \\
 \bottomrule

### 2018-02-14 draft/v3/main.tex
update figures and layout

--- 2017-12-16 draft/v2/latex/paper.tex
@@
-\documentclass[Afour,sageh,times]{sagej}
+%-------------------------------------------------------------------------------------------
+% enter the author and paper information below:
+%-------------------------------------------------------------------------------------------
+\newcommand{\myname}{Geoff Boeing}
+\newcommand{\myemail}{gboeing@berkeley.edu}
+\newcommand{\myaffiliation}{Department of City and Regional Planning\\University of California, Berkeley}
+\newcommand{\paperdate}{February 2018}
+\newcommand{\papertitle}{A Multi-Scale Analysis of 27,000 Urban Street Networks}
+\newcommand{\papercitation}{Boeing, G. 2018. \papertitle. Under review at \emph{Journal Name}.}
+\newcommand{\paperkeywords}{Urban Planning, Transportation, Data Science, Street Networks, Python, OpenStreetMap, Graph Theory, History}
 
+%-------------------------------------------------------------------------------------------
+% begin preamble setup
+%-------------------------------------------------------------------------------------------
+\RequirePackage[l2tabu,orthodox]{nag}
+\documentclass[11pt]{article}
+
+% import these packages first, in order
+\usepackage[utf8]{inputenc}
+\usepackage[strict,autostyle]{csquotes}
+\usepackage[USenglish]{babel}
+
+% import everything else
+\usepackage{authblk}
+\usepackage{booktabs}
+\usepackage{crimson}
+\usepackage{datetime}
+\usepackage[final]{draftwatermark}
+\usepackage{endnotes}
+\usepackage{geometry}
 \usepackage{graphicx}
-\usepackage{geometry}
-\usepackage{booktabs}
-\usepackage{pdflscape}
+\usepackage{hyperref}
+\usepackage{microtype}
+\usepackage{multicol}
+\usepackage{natbib}
+\usepackage{rotating}
 \usepackage{setspace}
-\usepackage[USenglish]{babel}
-\usepackage{csquotes}
-\usepackage{natbib}
 \usepackage{url}
 
-\usepackage{moreverb}
-\usepackage[colorlinks,bookmarksopen,bookmarksnumbered,citecolor=red,urlcolor=red]{hyperref}
-\newcommand\BibTeX{{\rmfamily B\kern-.05em \textsc{i\kern-.025em b}\kern-.08em
-		T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
-\def\volumeyear{2017}
+% location of figure files
+\graphicspath{{./figures/}}
 
-\renewcommand{\cite}{\citep}
-\doublespacing
+% configure the page layout
+\geometry{
+	paper=letterpaper,         % paper size
+	top=3cm,                   % margin sizes
+	bottom=3cm,
+	left=2.5cm,
+	right=2.5cm}
+\setlength{\columnsep}{0.75cm} % space between text columns
+\setstretch{1.00}              % line spacing
+\clubpenalty=10000             % prevent orphans
+\widowpenalty=10000            % prevent widows
+\let\footnote=\endnote         % turn footnotes into endnotes
 
+% configure pdf metadata and link handling
+\hypersetup{
+	pdfauthor={\myname},
+	pdftitle={\papertitle},
+	pdfsubject={\papertitle},
+	pdfkeywords={\paperkeywords},
+	pdffitwindow=true,         % window fit to page when opened
+	breaklinks=true,           % break links that overflow horizontally
+	hidelinks=true             % remove link color and border
+}
+
+% configure watermark appearance
+\SetWatermarkText{DRAFT}
+\SetWatermarkScale{1.3}
+\SetWatermarkLightness{0.9}
+
+%-------------------------------------------------------------------------------------------
+% end preamble setup, begin document
+%-------------------------------------------------------------------------------------------
 \begin{document}
-\title{A Multi-Scale Analysis of 27,000 Urban Street Networks}
-\author{Author Redacted}
-\date{March 2017}
+
+	\title{\papertitle\footnote{Manuscript under review; see \url{http://geoffboeing.com/publications/multiscale-analysis-street-networks/}}}
+	\date{\paperdate}
+	\author[]{\myname \thanks{Email: \href{mailto:\myemail}{\myemail}}}
+	\affil[]{\myaffiliation}
+
 \maketitle
@@
 \begin{abstract}
-
 OpenStreetMap offers an underexplored source of worldwide geospatial data useful to urban researchers.
@@
 These street networks and measures data sets have been shared in a public repository for other researchers to use.
+\end{abstract}
 
-\end{abstract}
+\vspace{0.25cm}
+
+\begin{multicols}{2}
 
@@
 
-On May 20, 1862, Abraham Lincoln signed the Homestead Act into law, making land across the United States' Midwest and Great Plains available for free to applicants \cite{porterfield_homestead_2005}.
-Under its auspices over the next 70 years, the federal government distributed 10\% of the entire US landmass to private owners in the form of 1.6 million homesteads \cite{lee_kansas_1979, sherraden_inclusion_2005}.
-New towns with gridiron street networks sprang up rapidly across the Great Plains and Midwest, due to both the prevailing urban design paradigm of the day and the standardized rectilinear town plats used repeatedly to lay out instant new cities \cite{southworth_streets_1997}.
+On May 20, 1862, Abraham Lincoln signed the Homestead Act into law, making land across the United States' Midwest and Great Plains available for free to applicants \citep{porterfield_homestead_2005}.
+Under its auspices over the next 70 years, the federal government distributed 10\% of the entire US landmass to private owners in the form of 1.6 million homesteads \citep{lee_kansas_1979, sherraden_inclusion_2005}.
+New towns with gridiron street networks sprang up rapidly across the Great Plains and Midwest, due to both the prevailing urban design paradigm of the day and the standardized rectilinear town plats used repeatedly to lay out instant new cities \citep{southworth_streets_1997}.
 Through path dependence, the spatial signatures of these land use laws, design paradigms, and planning instruments can still be seen today in these cities' urban forms and street networks.
@@
 
-Network analysis is a natural approach to the study of cities as complex systems \cite{masucci_random_2009}.
-The empirical literature on street networks is growing ever richer, but suffers from some limitations---discussed in detail in \cite{redacted_osmnx:_2017} and summarized here.
-First, sample sizes tend to be fairly small due to data availability, gathering, and processing constraints: most studies in this literature that conduct topological or metric analyses tend to have sample sizes ranging around 10 to 50 networks \cite{buhl_topological_2006, cardillo_structural_2006, barthelemy_modeling_2008, strano_urban_2013}, which may limit the generalizability and interpretability of findings.
+Network analysis is a natural approach to the study of cities as complex systems \citep{masucci_random_2009}.
+The empirical literature on street networks is growing ever richer, but suffers from some limitations---discussed in detail in \citep{boeing_osmnx:_2017} and summarized here.
+First, sample sizes tend to be fairly small due to data availability, gathering, and processing constraints: most studies in this literature that conduct topological or metric analyses tend to have sample sizes ranging around 10 to 50 networks \citep{buhl_topological_2006, cardillo_structural_2006, barthelemy_modeling_2008, strano_urban_2013}, which may limit the generalizability and interpretability of findings.
 Second, reproducibility has been difficult when the dozens of decisions that go into analysis---such as spatial extents, topological simplification and correction, definitions of nodes and edges, etc.---are ad hoc or only partly reported \citep[e.g.,][]{porta_network_2006, strano_urban_2013}.
-Third, and related to the first two, studies frequently oversimplify to planar or undirected primal graphs for tractability \cite[e.g.,][]{buhl_topological_2006, cardillo_structural_2006, barthelemy_modeling_2008, masucci_random_2009}, or use dual graphs despite the loss of geographic, metric information \cite{batty_network_2005, jiang_integration_2002, ratti_space_2004, crucitti_centrality_2006, crucitti_centrality_2006-1}.
+Third, and related to the first two, studies frequently oversimplify to planar or undirected primal graphs for tractability \citep[e.g.,][]{buhl_topological_2006, cardillo_structural_2006, barthelemy_modeling_2008, masucci_random_2009}, or use dual graphs despite the loss of geographic, metric information \citep{batty_network_2005, jiang_integration_2002, ratti_space_2004, crucitti_centrality_2006, crucitti_centrality_2006-1}.
 
@@
 It examines the street networks of every incorporated city and town (population size $N=19655$), census urbanized area ($N=497$), and Zillow-defined neighborhood ($N=6857$) in the US.
-To do so, it uses OSMnx---a new street network research toolkit \cite{redacted_osmnx:_2017}---to download, construct, and analyze these street networks at metropolitan, municipal, and neighborhood scales.
-These street networks and measures data sets have been compiled and shared in a public repository for other researchers to repurpose at [redacted].
+To do so, it uses OSMnx\footnote{OSMnx is freely available online at \url{https://github.com/gboeing/osmnx}}---a new street network research toolkit \citep{boeing_osmnx:_2017}---to download, construct, and analyze these street networks at metropolitan, municipal, and neighborhood scales.
+These street networks and measures data sets have been compiled and shared in a public repository at the Harvard Dataverse\footnote{\label{note_data_repo}Data repository available online at \url{https://dataverse.harvard.edu/dataverse/osmnx-street-networks}} for other researchers to repurpose.
 
@@
 A \emph{directed} network has directed edges: that is, edge $uv$ points one-way from node $u$ to node $v$, but there need not exist a reciprocal edge $vu$.
-A \emph{planar} network can be represented in two dimensions with its edges intersecting only at nodes \cite{viana_simplicity_2013, fischer_spatial_2014}.
-Most street networks are nonplanar---due to grade-separated expressways, overpasses, bridges, tunnels, etc.---but most quantitative studies of urban street networks represent them as planar \cite[e.g.,][]{buhl_topological_2006, cardillo_structural_2006, barthelemy_modeling_2008, masucci_random_2009, strano_urban_2013} for tractability because bridges and tunnels are uncommon in some cities.
-Planar graphs may reasonably model the street networks of old European town centers, but poorly model the street networks of modern autocentric cities like Los Angeles or Tokyo with many grade-separated expressways, bridges, and underpasses.
+A \emph{planar} network can be represented in two dimensions with its edges intersecting only at nodes \citep{viana_simplicity_2013, fischer_spatial_2014}.
+Most street networks are nonplanar---due to grade-separated expressways, overpasses, bridges, tunnels, etc.---but most quantitative studies of urban street networks represent them as planar \citep[e.g.,][]{buhl_topological_2006, cardillo_structural_2006, barthelemy_modeling_2008, masucci_random_2009, strano_urban_2013} for tractability because bridges and tunnels are uncommon in some cities.
+Planar graphs may reasonably model the street networks of old European town centers, but poorly model the street networks of modern autocentric cities like Los Angeles or Shanghai with many grade-separated expressways, bridges, and underpasses \citep{boeing_planarity_2018}.
 
@@
 This study uses OSMnx to download, construct, correct, analyze, and visualize street networks at metropolitan, municipal, and neighborhood scales.
-OSMnx is a Python-based research tool that easily downloads OpenStreetMap data for any place name, address, or polygon in the world, then constructs it into a spatially-embedded graph-theoretic object for analysis and visualization \cite{redacted_osmnx:_2017}.
-OpenStreetMap is a collaborative worldwide mapping project that makes its spatial data available via various APIs \cite{corcoran_analysing_2013, jokar_arsanjani_openstreetmap_2015}.
-These data are of high quality and compare favorably to CIA World Factbook estimates and US Census TIGER/Line data \cite{haklay_how_2010, over_generating_2010, zielstra_comparative_2011, maron_how_2015, wu_improving_2005, frizzelle_importance_2009}.
-In 2007, OpenStreetMap imported the TIGER/Line roads (2005 vintage) and since then, many community-led corrections and improvements have been made \cite{willis_openstreetmap_2008}.
+OSMnx is a Python-based research tool that easily downloads OpenStreetMap data for any place name, address, or polygon in the world, then constructs it into a spatially-embedded graph-theoretic object for analysis and visualization \citep{boeing_osmnx:_2017,boeing_methods_2017}.
+OpenStreetMap is a collaborative worldwide mapping project that makes its spatial data available via various APIs \citep{corcoran_analysing_2013, jokar_arsanjani_openstreetmap_2015}.
+These data are of high quality and compare favorably to CIA World Factbook estimates and US Census TIGER/Line data \citep{haklay_how_2010, over_generating_2010, zielstra_comparative_2011, maron_how_2015, wu_improving_2005, frizzelle_importance_2009}.
+In 2007, OpenStreetMap imported the TIGER/Line roads (2005 vintage) and since then, many community-led corrections and improvements have been made \citep{willis_openstreetmap_2008}.
 Many of these additions go beyond TIGER/Line's scope, including passageways between buildings, footpaths through parks, bike routes, and detailed feature attributes such as finer-grained street classifiers, speed limits, etc.
@@
 The first defines the metropolitan-scale study sites using the 2016 TIGER/Line shapefile of US census bureau urban areas.
-Each census-defined urban area comprises a set of tracts that meet a minimum density threshold \cite{u.s._census_bureau_2010_2010}.
+Each census-defined urban area comprises a set of tracts that meet a minimum density threshold \citep{u.s._census_bureau_2010_2010}.
 We retain only the \emph{urbanized areas} subset of these data (i.e., areas with greater than 50,000 population), discarding the small \emph{urban clusters} subset.
@@
 
-For each of these geometries, we use OSMnx to download the (drivable, public) street network within it, a process described in detail in \cite{redacted_osmnx:_2017} and summarized here.
+For each of these geometries, we use OSMnx to download the (drivable, public) street network within it, a process described in detail in \citet{boeing_osmnx:_2017} and summarized here.
 First OSMnx buffers each geometry by 0.5 km, then downloads the OpenStreetMap \enquote{nodes} and \enquote{ways} within this buffer.
@@
 Finally, it calculates metric and topological measures for each network, summarized below.
-Such measures extend the toolkit commonly used in urban form studies \cite{talen_measuring_2003, ewing_travel_2010}.
+Such measures extend the toolkit commonly used in urban form studies \citep{talen_measuring_2003, ewing_travel_2010}.
 
@@
 Brief descriptions of these OSMnx-calculated measures are discussed here, but extended technical definitions and algorithms can be found in e.g.
-\cite{trudeau_introduction_1994, albert_statistical_2002, dorogovtsev_evolution_2002, brandes_network_2005, costa_characterization_2007, newman_structure_2003, newman_networks:_2010, barthelemy_spatial_2011, cranmer_navigating_2017}.
+\citep{trudeau_introduction_1994, albert_statistical_2002, dorogovtsev_evolution_2002, brandes_network_2005, costa_characterization_2007, newman_structure_2003, newman_networks:_2010, barthelemy_spatial_2011, cranmer_navigating_2017}.
 The \emph{average street segment length} is a linear proxy for block size and specifies the network's grain.
@@
 \emph{Average circuity} measures the ratio of edge lengths to the great-circle distances between the nodes these edges connect, indicating the street pattern's curvilinearity (cf.
-[redacted]).
+\citet{boeing_morphology_2018}).
 
@@
 Connectivity represents the fewest number of nodes or edges that will disconnect the network if they are removed and is thus an indicator of resilience.
-A network's \emph{average node connectivity} (ANC)---the mean number of internally node-disjoint paths between each pair of nodes---more usefully represents how many nodes must be removed on average to disconnect a randomly selected pair of nodes \cite{beineke_average_2002,dankelmann_bounds_2003}.
+A network's \emph{average node connectivity} (ANC)---the mean number of internally node-disjoint paths between each pair of nodes---more usefully represents how many nodes must be removed on average to disconnect a randomly selected pair of nodes \citep{beineke_average_2002,dankelmann_bounds_2003}.
 Brittle points of vulnerability characterize networks with low average connectivity.
 
-A node's \emph{clustering coefficient} represents the ratio between its neighbors' links and the maximum number of links that could exist between them \cite{jiang_topological_2004, opsahl_clustering_2009}.
+A node's \emph{clustering coefficient} represents the ratio between its neighbors' links and the maximum number of links that could exist between them \citep{jiang_topological_2004, opsahl_clustering_2009}.
 The weighted clustering coefficient weights this by edge length and the average clustering coefficient is the mean of the clustering coefficients of all the nodes.
-\emph{Betweenness centrality} evaluates how many of the network's shortest paths pass through some node (or edge) to indicate its importance \cite{barthelemy_betweenness_2004,huang_trajgraph:_2016, zhong_revealing_2017}.
+\emph{Betweenness centrality} evaluates how many of the network's shortest paths pass through some node (or edge) to indicate its importance \citep{barthelemy_betweenness_2004,huang_trajgraph:_2016, zhong_revealing_2017}.
 A network's \emph{maximum betweenness centrality} (MBC) measures the share of shortest paths that pass through the network's most important node: higher maximum betweenness centralities suggest networks more prone to inefficiency if this important choke point should fail.
-Finally, \emph{PageRank} ranks nodes based on the structure of incoming links and the rank of the source node \cite{brin_anatomy_1998, jiang_predicting_2008, agryzkov_algorithm_2012, chin_geographically_2015, gleich_pagerank_2015}.
+Finally, \emph{PageRank} ranks nodes based on the structure of incoming links and the rank of the source node \citep{brin_anatomy_1998, jiang_predicting_2008, agryzkov_algorithm_2012, chin_geographically_2015, gleich_pagerank_2015}.
 
@@
 
-\begin{table*}
+\begin{table*}[htbp]
 \centering
@@
 \label{tab:measures_urban_areas}
+\small
 \begin{tabular}{ l r r r r r r }
@@
 
+\begin{figure*}[htbp]
+	\centering
+	\includegraphics[width=1\textwidth]{fig01.png}
+	\caption{Intersection density and average streets per node per urbanized area in the contiguous US.}
+	\label{fig:fig01}
+\end{figure*}
+
+\begin{figure*}[htbp]
+	\centering
+	\includegraphics[width=0.7\textwidth]{fig02.png}
+	\caption{Distribution of node types in 9 urbanized areas, with number of streets emanating from the node on the $x$-axis and proportion of nodes of this type on the $y$-axis }
+	\label{fig:fig02}
+\end{figure*}
+
 Nevertheless, looking across the data set provides a sense of the breadth of American metropolitan street networks.
@@
 In general, only the largest cities on the east coast (e.g., Boston, New York, Philadelphia, Washington) and Florida escape this trend.
-
-\begin{figure*}
-	\includegraphics[width=1\textwidth]{fig01.png}
-	\caption{Intersection density per urbanized area (left) and average streets per node per urbanized area (right) from lowest (pale yellow) to highest (dark red), in the contiguous US}
-	\label{fig:fig01}
-\end{figure*}
-
-\begin{figure*}
-	\includegraphics[width=0.7\textwidth]{fig02.png}
-	\caption{Distribution of node types in 9 urbanized areas, with number of streets emanating from the node on the $x$-axis and proportion of nodes of this type on the $y$-axis }
-	\label{fig:fig02}
-\end{figure*}
 
@@
 
-\newgeometry{margin=1in}
-\begin{landscape}
-\begin{table*}
+The dispersion index $D$ in Table \ref{tab:measures_urban_areas} demonstrates the heterogeneity of each indicator across the data set.
+Several \enquote{families} of indicators can be discerned by their heterogeneity.
+For instance, counts and totals such as $m$, $n$, and the total street length are extremely heterogeneous.
+Densities and average distances such as intersection density and the average street segment length exhibit only moderate heterogeneity.
+Finally, several topological measures such as the average clustering coefficient and PageRank are extremely homogeneous.
+Due to the substantial variation in urbanized area size, from 25 to 9,000 km\textsuperscript{2}, the preceding analysis covers a wide swath of metropolitan types.
+To better compare apples-to-apples, Table \ref{tab:largest_urbanized_areas} focuses on the 30 largest urbanized areas cross-sectionally to examine their metric and topological measures.
+This provides more consistent spatial scales and extents, while offering a window into the similarities and differences in the built forms of America's largest agglomerations.
+
+Among these urbanized areas, Milwaukee has the least circuitous network (6\% more circuitous than straight-line edges would be), and Orlando has the most (12\%).
+San Juan and Atlanta have the fewest streets per node on average (2.36 and 2.45, respectively), while Milwaukee has the most (3.03).
+Cincinnati has both the lowest intersection density (18/km\textsuperscript{2}) and street density (6.1 km/km\textsuperscript{2}) while Denver has the highest intersection density (40.6/km\textsuperscript{2}) and Miami and Los Angeles have the highest street density (10.6 km/km\textsuperscript{2}, apiece).
+In other words, Cincinnati has a particularly coarse-grained network with few connections and paths.
+The average street segment length, a proxy for block size, also reflects this: Cincinnati has the second highest (186 m), bested only by Cleveland (198 m).
+In contrast, the two lowest are Denver's 138-meter average and San Juan's 131-meter average.
+
+These metropolitan analyses consider trends in the built form at the scale of broad self-organized human systems and urbanized regions.
+However, they aggregate multiple heterogeneous municipalities and neighborhoods---the scales of human life, urban design projects, and planning jurisdiction---into single units of analysis.
+To disaggregate and analyze finer characteristics, the following subsections examine municipal- and neighborhood-scale street networks.
+
+\end{multicols}
+
+\begin{sidewaystable*}[p]
 \centering
@@
 \label{tab:largest_urbanized_areas}
+\small
 \begin{tabular}{p{2.5cm} p{1.5cm} p{1.5cm} p{1.5cm} p{1.55cm} p{1.5cm} p{1.5cm} p{1.5cm} p{1.5cm} p{1.5cm} p{1.5cm}}
@@
 \end{tabular}
-\end{table*}
-\end{landscape}
-\restoregeometry
+\end{sidewaystable*}
 
-The dispersion index $D$ in Table \ref{tab:measures_urban_areas} demonstrates the heterogeneity of each indicator across the data set.
-Several \enquote{families} of indicators can be discerned by their heterogeneity.
-For instance, counts and totals such as $m$, $n$, and the total street length are extremely heterogeneous.
-Densities and average distances such as intersection density and the average street segment length exhibit only moderate heterogeneity.
-Finally, several topological measures such as the average clustering coefficient and PageRank are extremely homogeneous.
-Due to the substantial variation in urbanized area size, from 25 to 9,000 km\textsuperscript{2}, the preceding analysis covers a wide swath of metropolitan types.
-To better compare apples-to-apples, Table \ref{tab:largest_urbanized_areas} focuses on the 30 largest urbanized areas cross-sectionally to examine their metric and topological measures.
-This provides more consistent spatial scales and extents, while offering a window into the similarities and differences in the built forms of America's largest agglomerations.
-
-Among these urbanized areas, Milwaukee has the least circuitous network (6\% more circuitous than straight-line edges would be), and Orlando has the most (12\%).
-San Juan and Atlanta have the fewest streets per node on average (2.36 and 2.45, respectively), while Milwaukee has the most (3.03).
-Cincinnati has both the lowest intersection density (18/km\textsuperscript{2}) and street density (6.1 km/km\textsuperscript{2}) while Denver has the highest intersection density (40.6/km\textsuperscript{2}) and Miami and Los Angeles have the highest street density (10.6 km/km\textsuperscript{2}, apiece).
-In other words, Cincinnati has a particularly coarse-grained network with few connections and paths.
-The average street segment length, a proxy for block size, also reflects this: Cincinnati has the second highest (186 m), bested only by Cleveland (198 m).
-In contrast, the two lowest are Denver's 138-meter average and San Juan's 131-meter average.
-
-These metropolitan analyses consider trends in the built form at the scale of broad self-organized human systems and urbanized regions.
-However, they aggregate multiple heterogeneous municipalities and neighborhoods---the scales of human life, urban design projects, and planning jurisdiction---into single units of analysis.
-To disaggregate and analyze finer characteristics, the following subsections examine municipal- and neighborhood-scale street networks.
+\begin{multicols}{2}
 
@@
 However, the latter's small sample size may limit the generalizability of this finding.
-To investigate this empirically, we examine the relationship between $L$ and $n$ for 19,655 US cities and towns and instead find a strong \emph{linear} relationship ($r^{2}=0.98$), as depicted in Figure \ref{fig:fig03}.
-We also find a similar linear relationship at the metropolitan and neighborhood scales.
+We examine the relationship between $L$ and $n$ across every US city and town and instead find a strong linear relationship ($r^{2}=0.98$), as depicted in Figure \ref{fig:fig03}.
+We also find a similar linear relationship at the metropolitan ($r^{2}=0.99$) and neighborhood ($r^{2}=0.98$) scales.
 
@@
 \label{tab:measures_cities}
+\small
 \begin{tabular}{ l r r r r r r }
@@
 
-Previous findings \citep[e.g.,][]{masucci_random_2009} also suggest street segment lengths in an urban network follow a power-law distribution.
-However, we find that these networks instead generally follow lognormal-style right-skewed distributions.
+Previous findings \citep[e.g.,][]{masucci_random_2009} suggest street segment lengths in an urban network follow a power-law distribution.
+We find that these networks instead generally follow lognormal-style right-skewed distributions.
 This makes theoretical sense as most street networks are not truly scale-free: for example, a typical street network might comprise very few very long street segments (e.g., 1 km), more medium-length segments (e.g., 250 m), many short segments (e.g., 80 m), but very few very short segments (e.g., 10 m).
@@
 The \emph{latter} results from the fact that most towns across the Great Plains exhibit minimal suburban sprawl.
-Thus, municipal boundaries snugly embrace the gridlike street network, without extending to accommodate a vast peripheral belt of twentieth century sprawl, circuity, and \enquote{loops and lollipops} \cite{southworth_streets_1997} that characterizes cities in e.g.
+Thus, municipal boundaries snugly embrace the gridlike street network, without extending to accommodate a vast peripheral belt of twentieth century sprawl, circuity, and \enquote{loops and lollipops} \citep{southworth_streets_1997} that characterizes cities in e.g.
 California that were settled in the same era but later subjected to substantial suburbanization.
 
-\begin{figure*}
-	\includegraphics[width=\textwidth]{fig03.png}
-	\caption{The linear relationship between total street length and number of nodes in the street networks of 6,857 US neighborhoods (left) and 19,655 US cities and towns (right).}
+\begin{figure*}[htbp]
+	\centering
+	\includegraphics[width=1\textwidth]{fig03.png}
+	\caption{The linear relationship between total street length and number of nodes in the street networks of every US urbanized area, city/town, and Zillow neighborhood.}
 	\label{fig:fig03}
@@
 
-\begin{figure*}
-	\includegraphics[width=\textwidth]{fig04.png}
-	\caption{Left: contiguous US states by median of mean streets per node in municipal street networks, from lowest (pale yellow) to highest (dark red).
-Right: contiguous US states by median of mean street segment length in municipal street networks, from longest (pale yellow) to shortest (dark red).}
+\begin{figure*}[htbp]
+	\centering
+	\includegraphics[width=1\textwidth]{fig04.png}
+	\caption{Contiguous US states by median of mean streets per node and by median of mean street segment length in municipal street networks.}
 	\label{fig:fig04}
@@
 
-\newgeometry{margin=2cm}
-\begin{table*}
+\begin{table*}[htbp]
 \centering
@@
 \label{tab:medians_state_aggregate}
+\footnotesize
 \begin{tabular}{ l r r r r }
@@
 \end{table*}
-\restoregeometry
 
@@
 \label{tab:measures_neighborhoods}
+\small
 \begin{tabular}{ l r r r r r r }
@@
 
+\begin{figure*}[htbp]
+	\centering
+	\includegraphics[width=0.8\textwidth]{fig05.png}
+	\caption{Square-mile comparisons of central cities and their suburbs.
+Left: top, downtown Philadelphia, PA; bottom, its suburb, King of Prussia.
+Middle: top, downtown Portland, OR; bottom, its suburb, Beaverton.
+Right: top, downtown San Francisco, CA; bottom, its suburb, Concord.}
+	\label{fig:fig05}
+\end{figure*}
+
 Due to the extreme values seen---resulting from the large variance in neighborhood size---we can filter the data set to examine only large neighborhoods (i.e., with area greater than the median value across the data set).
@@
 
-\begin{figure*}
-	\includegraphics[width=1\textwidth]{fig05.png}
-	\caption{Square-mile comparisons of central cities and their suburbs.
-Left: top, downtown Philadelphia, PA; bottom, its suburb, King of Prussia.
-Middle: top, downtown Portland, OR; bottom, its suburb, Beaverton.
-Right: top, downtown San Francisco, CA; bottom, its suburb, Concord.}
-	\label{fig:fig05}
-\end{figure*}
-
 To illustrate these morphologies, Figure \ref{fig:fig05} compares one square mile of the centers of Philadelphia, Portland, and San Francisco with one square mile of each of their suburbs.
@@
 In fact, the suburbs have more in common with one another---despite being hundreds or thousands of miles apart---than they do with their central city neighbors, suggesting that land use and an era's prevailing design paradigm is paramount to geographical localism and regional context.
-The top row of Figure \ref{fig:fig05} represents an era of planning and development that preceded the automobile, while the bottom row reflects the exclusionary zoning and mid--late twentieth century era of automobility in residential suburb design---namely the \enquote{loops and lollipops} and the \enquote{lollipops on a stick} design patterns \cite{southworth_streets_1997}.
+The top row of Figure \ref{fig:fig05} represents an era of planning and development that preceded the automobile, while the bottom row reflects the exclusionary zoning and mid--late twentieth century era of automobility in residential suburb design---namely the \enquote{loops and lollipops} and the \enquote{lollipops on a stick} design patterns \citep{southworth_streets_1997}.
 
@@
 Third, it investigated past findings about the distribution of street segment lengths and the relationship between the total street length and the number of nodes in a network.
-This study has made all of these network datasets---for 497 urbanized areas, 19,655 cities and towns, and 6,857 neighborhoods---along with all of their attribute data and morphological measures available in an online public repository for other researchers to study and repurpose at [redacted].
+This study has made all of these network datasets---for 497 urbanized areas, 19,655 cities and towns, and 6,857 neighborhoods---along with all of their attribute data and morphological measures available in an online public repository for other researchers to study and repurpose.
 
-\begin{thebibliography}{54}
-	\providecommand{\natexlab}[1]{#1}
-	\providecommand{\url}[1]{\texttt{#1}}
-	\expandafter\ifx\csname urlstyle\endcsname\relax
-	\providecommand{\doi}[1]{doi: #1}\else
-	\providecommand{\doi}{doi: \begingroup \urlstyle{rm}\Url}\fi
+% print the footnotes as endnotes, if any exist
+\IfFileExists{\jobname.ent}{\theendnotes}{}
 
-	\bibitem[Agryzkov et~al.(2012)Agryzkov, Oliver, Tortosa, and
-	Vicent]{agryzkov_algorithm_2012}
-	T.~Agryzkov, J.~L. Oliver, L.~Tortosa, and J.~F. Vicent.
-	\newblock An algorithm for ranking the nodes of an urban network based on the
-	concept of {PageRank} vector.
-	\newblock \emph{Applied Mathematics and Computation}, 219\penalty0
-	(4):\penalty0 2186--2193, 2012.
-	\newblock ISSN 00963003.
-	\newblock \doi{10.1016/j.amc.2012.08.064}.
+% print the bibliography
+\setlength{\bibsep}{0.00cm plus 0.05cm}
+\bibliographystyle{apalike}
+\bibliography{references}
 
-	\bibitem[Albert and Barabási(2002)]{albert_statistical_2002}
-	R.~Albert and A.-L. Barabási.
-	\newblock Statistical mechanics of complex networks.
-	\newblock \emph{Reviews of Modern Physics}, 74\penalty0 (1):\penalty0 47, 2002.
-	\newblock ISSN 0034-6861.
-	\newblock \doi{10.1103/RevModPhys.74.47}.
-
-	\bibitem[Albrecht and Abramovitz(2014)]{albrecht_indicator_2014}
-	J.~Albrecht and M.~Abramovitz.
-	\newblock Indicator {Analysis} for {Unpacking} {Poverty} in {New} {York}
-	{City}.
-	\newblock Technical report, CUNY, New York, NY, 2014.
-
-	\bibitem[Barthelemy(2004)]{barthelemy_betweenness_2004}
-	M.~Barthelemy.
-	\newblock Betweenness centrality in large complex networks.
-	\newblock \emph{The European Physical Journal B: Condensed Matter and Complex
-		Systems}, 38\penalty0 (2):\penalty0 163--168, 2004.
-	\newblock ISSN 1434-6028.
-	\newblock \doi{10.1140/epjb/e2004-00111-4}.
-
-	\bibitem[Barthelemy(2011)]{barthelemy_spatial_2011}
-	M.~Barthelemy.
-	\newblock Spatial networks.
-	\newblock \emph{Physics Reports}, 499\penalty0 (1-3):\penalty0 1--101, 2011.
-	\newblock ISSN 03701573.
-	\newblock \doi{10.1016/j.physrep.2010.11.002}.
-
-	\bibitem[Barthelemy and Flammini(2008)]{barthelemy_modeling_2008}
-	M.~Barthelemy and A.~Flammini.
-	\newblock Modeling {Urban} {Street} {Patterns}.
-	\newblock \emph{Physical Review Letters}, 100\penalty0 (13), 2008.
-	\newblock ISSN 0031-9007.
-	\newblock \doi{10.1103/PhysRevLett.100.138702}.
-
-	\bibitem[Batty(2005)]{batty_network_2005}
-	M.~Batty.
-	\newblock Network geography: {Relations}, interactions, scaling and spatial
-	processes in {GIS}.
-	\newblock In D.~J. Unwin and P.~Fisher, editors, \emph{Re-presenting
-		{Geographical} {Information} {Systems}}, pages 149--170. John Wiley \& Sons,
-	Chichester, England, 2005.
-
-	\bibitem[Beineke et~al.(2002)Beineke, Oellermann, and
-	Pippert]{beineke_average_2002}
-	L.~W. Beineke, O.~R. Oellermann, and R.~E. Pippert.
-	\newblock The {Average} {Connectivity} of a {Graph}.
-	\newblock \emph{Discrete Mathematics}, 252\penalty0 (1):\penalty0 31--45, 2002.
-	\newblock ISSN 0012-365X.
-	\newblock \doi{10.1016/S0012-365X(01)00180-7}.
-
-	\bibitem[Besbris et~al.(2015)Besbris, Faber, Rich, and
-	Sharkey]{besbris_effect_2015}
-	M.~Besbris, J.~W. Faber, P.~Rich, and P.~Sharkey.
-	\newblock Effect of neighborhood stigma on economic transactions.
-	\newblock \emph{Proceedings of the National Academy of Sciences}, 112\penalty0
-	(16):\penalty0 4994--4998, 2015.
-	\newblock ISSN 0027-8424.
-	\newblock \doi{10.1073/pnas.1414139112}.
-
-	\bibitem[Brandes and Erlebach(2005)]{brandes_network_2005}
-	U.~Brandes and T.~Erlebach, editors.
-	\newblock \emph{Network analysis: methodological foundations}.
-	\newblock Number 3418 in Lecture {Notes} in {Computer} {Science}. Springer,
-	Berlin, Germany, 2005.
-	\newblock ISBN 978-3-540-24979-5.
-
-	\bibitem[Brin and Page(1998)]{brin_anatomy_1998}
-	S.~Brin and L.~Page.
-	\newblock The anatomy of a large-scale hypertextual web search engine.
-	\newblock \emph{Computer Networks and ISDN Systems: Proceedings of the Seventh
-		International World Wide Web Conference}, 30\penalty0 (1-7):\penalty0
-	107--117, 1998.
-	\newblock ISSN 1389-1286.
-	\newblock \doi{10.1016/S0169-7552(98)00110-X}.
-
-	\bibitem[Buhl et~al.(2006)Buhl, Gautrais, Reeves, Solé, Valverde, Kuntz, and
-	Theraulaz]{buhl_topological_2006}
-	J.~Buhl, J.~Gautrais, N.~Reeves, R.~V. Solé, S.~Valverde, P.~Kuntz, and
-	G.~Theraulaz.
-	\newblock Topological patterns in street networks of self-organized urban
-	settlements.
-	\newblock \emph{The European Physical Journal B: Condensed Matter and Complex
-		Systems}, 49\penalty0 (4):\penalty0 513--522, 2006.
-	\newblock ISSN 1434-6028.
-	\newblock \doi{10.1140/epjb/e2006-00085-1}.
-
-	\bibitem[Cardillo et~al.(2006)Cardillo, Scellato, Latora, and
-	Porta]{cardillo_structural_2006}
-	A.~Cardillo, S.~Scellato, V.~Latora, and S.~Porta.
-	\newblock Structural properties of planar graphs of urban street patterns.
-	\newblock \emph{Physical Review E}, 73\penalty0 (6), 2006.
-	\newblock ISSN 1539-3755.
-	\newblock \doi{10.1103/PhysRevE.73.066107}.
-
-	\bibitem[Chin and Wen(2015)]{chin_geographically_2015}
-	W.-C.-B. Chin and T.-H. Wen.
-	\newblock Geographically {Modified} {PageRank} {Algorithms}: {Identifying} the
-	{Spatial} {Concentration} of {Human} {Movement} in a {Geospatial} {Network}.
-	\newblock \emph{PLoS ONE}, 10\penalty0 (10):\penalty0 e0139509, 2015.
-	\newblock ISSN 1932-6203.
-	\newblock \doi{10.1371/journal.pone.0139509}.
-
-	\bibitem[Corcoran et~al.(2013)Corcoran, Mooney, and
-	Bertolotto]{corcoran_analysing_2013}
-	P.~Corcoran, P.~Mooney, and M.~Bertolotto.
-	\newblock Analysing the growth of {OpenStreetMap} networks.
-	\newblock \emph{Spatial Statistics}, 3:\penalty0 21--32, 2013.
-	\newblock ISSN 22116753.
-	\newblock \doi{10.1016/j.spasta.2013.01.002}.
-
-	\bibitem[Costa et~al.(2007)Costa, Rodrigues, Travieso, and
-	Villas~Boas]{costa_characterization_2007}
-	L.~d.~F. Costa, F.~A. Rodrigues, G.~Travieso, and P.~R. Villas~Boas.
-	\newblock Characterization of complex networks: {A} survey of measurements.
-	\newblock \emph{Advances in Physics}, 56\penalty0 (1):\penalty0 167--242, 2007.
-	\newblock ISSN 0001-8732, 1460-6976.
-	\newblock \doi{10.1080/00018730601170527}.
-
-	\bibitem[Cranmer et~al.(2017)Cranmer, Leifeld, McClurg, and
-	Rolfe]{cranmer_navigating_2017}
-	S.~J. Cranmer, P.~Leifeld, S.~D. McClurg, and M.~Rolfe.
-	\newblock Navigating the range of statistical tools for inferential network
-	analysis.
-	\newblock \emph{American Journal of Political Science}, 61\penalty0
-	(1):\penalty0 237--251, 2017.
-	\newblock \doi{10.1111/ajps.12263}.
-
-	\bibitem[Crucitti et~al.(2006{\natexlab{a}})Crucitti, Latora, and
-	Porta]{crucitti_centrality_2006}
-	P.~Crucitti, V.~Latora, and S.~Porta.
-	\newblock Centrality measures in spatial networks of urban streets.
-	\newblock \emph{Physical Review E}, 73\penalty0 (3):\penalty0 036125,
-	2006{\natexlab{a}}.
-	\newblock ISSN 1539-3755.
-	\newblock \doi{10.1103/PhysRevE.73.036125}.
-
-	\bibitem[Crucitti et~al.(2006{\natexlab{b}})Crucitti, Latora, and
-	Porta]{crucitti_centrality_2006-1}
-	P.~Crucitti, V.~Latora, and S.~Porta.
-	\newblock Centrality in networks of urban streets.
-	\newblock \emph{Chaos: An Interdisciplinary Journal of Nonlinear Science},
-	16\penalty0 (1):\penalty0 015113, 2006{\natexlab{b}}.
-	\newblock ISSN 10541500.
-	\newblock \doi{10.1063/1.2150162}.
-
-	\bibitem[Dankelmann and Oellermann(2003)]{dankelmann_bounds_2003}
-	P.~Dankelmann and O.~R. Oellermann.
-	\newblock Bounds on the average connectivity of a graph.
-	\newblock \emph{Discrete Applied Mathematics}, 129\penalty0 (2):\penalty0
-	305--318, 2003.
-	\newblock ISSN 0166-218X.
-	\newblock \doi{10.1016/S0166-218X(02)00572-3}.
-
-	\bibitem[Dorogovtsev and Mendes(2002)]{dorogovtsev_evolution_2002}
-	S.~Dorogovtsev and J.~Mendes.
-	\newblock Evolution of networks.
-	\newblock \emph{Advances in Physics}, 51\penalty0 (4):\penalty0 1079--1187,
-	2002.
-	\newblock \doi{10.1080 /0001873011011251}.
-
-	\bibitem[Ewing and Cervero(2010)]{ewing_travel_2010}
-	R.~Ewing and R.~Cervero.
-	\newblock Travel and the {Built} {Environment}: {A} {Meta}-{Analysis}.
-	\newblock \emph{Journal of the American Planning Association}, 76\penalty0
-	(3):\penalty0 265--294, 2010.
-	\newblock ISSN 0194-4363.
-	\newblock \doi{10.1080/01944361003766766}.
-
-	\bibitem[Frizzelle et~al.(2009)Frizzelle, Evenson, Rodriguez, and
-	Laraia]{frizzelle_importance_2009}
-	B.~Frizzelle, K.~Evenson, D.~Rodriguez, and B.~Laraia.
-	\newblock The importance of accurate road data for spatial applications in
-	public health: customizing a road network.
-	\newblock \emph{International Journal of Health Geographics}, 8\penalty0 (24),
-	2009.
-	\newblock ISSN 1476-072X.
-	\newblock \doi{10.1186/1476-072X-8-24}.
-
-	\bibitem[Gleich(2015)]{gleich_pagerank_2015}
-	D.~F. Gleich.
-	\newblock {PageRank} {Beyond} the {Web}.
-	\newblock \emph{SIAM Review}, 57\penalty0 (3):\penalty0 321--363, 2015.
-	\newblock ISSN 0036-1445.
-	\newblock \doi{10.1137/140976649}.
-
-	\bibitem[Haklay(2010)]{haklay_how_2010}
-	M.~Haklay.
-	\newblock How {Good} is {Volunteered} {Geographical} {Information}? {A}
-	{Comparative} {Study} of {OpenStreetMap} and {Ordnance} {Survey} {Datasets}.
-	\newblock \emph{Environment and Planning B: Planning and Design}, 37\penalty0
-	(4):\penalty0 682--703, 2010.
-	\newblock ISSN 1472-3417.
-	\newblock \doi{10.1068/b35097}.
-
-	\bibitem[Huang et~al.(2016)Huang, Zhao, Ma, Yang, Ye, and
-	Zhang]{huang_trajgraph:_2016}
-	X.~Huang, Y.~Zhao, C.~Ma, J.~Yang, X.~Ye, and C.~Zhang.
-	\newblock {TrajGraph}: {A} {Graph}-{Based} {Visual} {Analytics} {Approach} to
-	{Studying} {Urban} {Network} {Centralities} {Using} {Taxi} {Trajectory}
-	{Data}.
-	\newblock \emph{IEEE Transactions on Visualization and Computer Graphics},
-	22\penalty0 (1):\penalty0 160--169, 2016.
-	\newblock ISSN 1077-2626.
-	\newblock \doi{10.1109/TVCG.2015.2467771}.
-
-	\bibitem[Jiang(2008)]{jiang_predicting_2008}
-	B.~Jiang.
-	\newblock Predicting human movement using {PageRank} in an urban environment.
-	\newblock 2008.
-
-	\bibitem[Jiang and Claramunt(2002)]{jiang_integration_2002}
-	B.~Jiang and C.~Claramunt.
-	\newblock Integration of space syntax into {GIS}: new perspectives for urban
-	morphology.
-	\newblock \emph{Transactions in GIS}, 6\penalty0 (3):\penalty0 295--309, 2002.
-	\newblock ISSN 1467-9671.
-	\newblock \doi{10.1111/1467-9671.00112}.
-
-	\bibitem[Jiang and Claramunt(2004)]{jiang_topological_2004}
-	B.~Jiang and C.~Claramunt.
-	\newblock Topological {Analysis} of {Urban} {Street} {Networks}.
-	\newblock \emph{Environment and Planning B: Planning and Design}, 31\penalty0
-	(1):\penalty0 151--162, 2004.
-	\newblock ISSN 0265-8135.
-	\newblock \doi{10.1068/b306}.
-
-	\bibitem[Jokar~Arsanjani et~al.(2015)Jokar~Arsanjani, Zipf, Mooney, and
-	Helbich]{jokar_arsanjani_openstreetmap_2015}
-	J.~Jokar~Arsanjani, A.~Zipf, P.~Mooney, and M.~Helbich, editors.
-	\newblock \emph{{OpenStreetMap} in {GIScience}}.
-	\newblock Lecture {Notes} in {Geoinformation} and {Cartography}. Springer
-	International, Cham, Switzerland, 2015.
-	\newblock ISBN 978-3-319-14279-1.
-
-	\bibitem[Lee(1979)]{lee_kansas_1979}
-	L.~B. Lee.
-	\newblock \emph{Kansas and the {Homestead} act, 1862-1905}.
-	\newblock The management of public lands in the {United} {States}. Ayer Company
-	Publishers, North Stratford, NH, 1979.
-	\newblock ISBN 978-0-405-11341-3.
-
-	\bibitem[Maron(2015)]{maron_how_2015}
-	M.~Maron.
-	\newblock How complete is {OpenStreetMap}?
-	\newblock \emph{Mapbox}, Nov. 2015.
-	\newblock URL \url{https://www.mapbox.com/blog/how-complete-is-openstreetmap/}.
-
-	\bibitem[Masucci et~al.(2009)Masucci, Smith, Crooks, and
-	Batty]{masucci_random_2009}
-	A.~P. Masucci, D.~Smith, A.~Crooks, and M.~Batty.
-	\newblock Random planar graphs and the {London} street network.
-	\newblock \emph{The European Physical Journal B: Condensed Matter and Complex
-		Systems}, 71\penalty0 (2):\penalty0 259--271, 2009.
-	\newblock ISSN 1434-6028.
-	\newblock \doi{10.1140/epjb/e2009-00290-4}.
-
-	\bibitem[Newman(2003)]{newman_structure_2003}
-	M.~E.~J. Newman.
-	\newblock The {Structure} and {Function} of {Complex} {Networks}.
-	\newblock \emph{SIAM Review}, 45\penalty0 (2):\penalty0 167--256, 2003.
-	\newblock ISSN 0036-1445.
-	\newblock \doi{10.1137/S003614450342480}.
-
-	\bibitem[Newman(2010)]{newman_networks:_2010}
-	M.~E.~J. Newman.
-	\newblock \emph{Networks: {An} {Introduction}}.
-	\newblock Oxford University Press, Oxford, England, 2010.
-	\newblock ISBN 978-0-19-920665-0.
-
-	\bibitem[Opsahl and Panzarasa(2009)]{opsahl_clustering_2009}
-	T.~Opsahl and P.~Panzarasa.
-	\newblock Clustering in weighted networks.
-	\newblock \emph{Social Networks}, 31\penalty0 (2):\penalty0 155--163, 2009.
-	\newblock ISSN 03788733.
-	\newblock \doi{10.1016/j.socnet.2009.02.002}.
-
-	\bibitem[O'Sullivan(2014)]{fischer_spatial_2014}
-	D.~O'Sullivan.
-	\newblock Spatial {Network} {Analysis}.
-	\newblock In M.~M. Fischer and P.~Nijkamp, editors, \emph{Handbook of
-		{Regional} {Science}}, pages 1253--1273. Springer-Verlag, Berlin, Germany,
-	2014.
-	\newblock ISBN 978-3-642-23429-3.
-
-	\bibitem[Over et~al.(2010)Over, Schilling, Neubauer, and
-	Zipf]{over_generating_2010}
-	M.~Over, A.~Schilling, S.~Neubauer, and A.~Zipf.
-	\newblock Generating web-based 3d {City} {Models} from {OpenStreetMap}: {The}
-	current situation in {Germany}.
-	\newblock \emph{Computers, Environment and Urban Systems}, 34\penalty0
-	(6):\penalty0 496--507, 2010.
-	\newblock ISSN 01989715.
-	\newblock \doi{10.1016/j.compenvurbsys.2010.05.001}.
-
-	\bibitem[Porta et~al.(2006)Porta, Crucitti, and Latora]{porta_network_2006}
-	S.~Porta, P.~Crucitti, and V.~Latora.
-	\newblock The network analysis of urban streets: {A} dual approach.
-	\newblock \emph{Physica A: Statistical Mechanics and its Applications},
-	369\penalty0 (2):\penalty0 853--866, 2006.
-	\newblock ISSN 03784371.
-	\newblock \doi{10.1016/j.physa.2005.12.063}.
-
-	\bibitem[Porterfield(2005)]{porterfield_homestead_2005}
-	J.~Porterfield.
-	\newblock \emph{The {Homestead} {Act} of 1862: {A} {Primary} {Source} {History}
-		of the {Settlement} of the {American} {Heartland} in the {Late} 19th
-		{Century}}.
-	\newblock Rosen Publishing Group, New York, NY, 1st edition, 2005.
-	\newblock ISBN 978-1-4042-0178-1.
-
-	\bibitem[Ratti(2004)]{ratti_space_2004}
-	C.~Ratti.
-	\newblock Space syntax: some inconsistencies.
-	\newblock \emph{Environment and Planning B: Planning and Design}, 31\penalty0
-	(4):\penalty0 487--499, 2004.
-	\newblock ISSN 1472-3417.
-	\newblock \doi{10.1068/b3019}.
-
-	\bibitem[Redacted(2017)]{redacted_osmnx:_2017}
-	R.~Redacted.
-	\newblock Title redacted.
-	\newblock \penalty0 (Redacted), 2017.
-	\newblock \doi{Redacted}.
-
-	\bibitem[Schernthanner et~al.(2016)Schernthanner, Asche, Gonschorek, and
-	Scheele]{schernthanner_spatial_2016}
-	H.~Schernthanner, H.~Asche, J.~Gonschorek, and L.~Scheele.
-	\newblock Spatial {Modeling} and {Geovisualization} of {Rental} {Prices} for
-	{Real} {Estate} {Portals}.
-	\newblock In O.~Gervasi, B.~Murgante, S.~Misra, A.~M. A.~C. Rocha, C.~M. Torre,
-	D.~Taniar, B.~O. Apduhan, E.~Stankova, and S.~Wang, editors,
-	\emph{Computational {Science} and {Its} {Applications} - {ICCSA} 2016},
-	number 9788 in Lecture {Notes} in {Computer} {Science}, pages 120--133.
-	Springer International, Cham, Switzerland, 2016.
-	\newblock ISBN 978-3-319-42110-0.
-	\newblock DOI: 10.1007/978-3-319-42111-7\_11.
-
-	\bibitem[Sherraden(2005)]{sherraden_inclusion_2005}
-	M.~Sherraden.
-	\newblock \emph{Inclusion in the {American} {Dream}: {Assets}, {Poverty}, and
-		{Public} {Policy}}.
-	\newblock Oxford University Press, Oxford, England, 2005.
-	\newblock ISBN 978-0-19-534709-8.
-
-	\bibitem[Southworth and Ben-Joseph(1997)]{southworth_streets_1997}
-	M.~Southworth and E.~Ben-Joseph.
-	\newblock \emph{Streets and the {Shaping} of {Towns} and {Cities}}.
-	\newblock McGraw-Hill, New York, NY, 1997.
-	\newblock ISBN 978-1-55963-916-3.
-
-	\bibitem[Strano et~al.(2013)Strano, Viana, da~Fontoura~Costa, Cardillo, Porta,
-	and Latora]{strano_urban_2013}
-	E.~Strano, M.~Viana, L.~da~Fontoura~Costa, A.~Cardillo, S.~Porta, and
-	V.~Latora.
-	\newblock Urban {Street} {Networks}, a {Comparative} {Analysis} of {Ten}
-	{European} {Cities}.
-	\newblock \emph{Environment and Planning B: Planning and Design}, 40\penalty0
-	(6):\penalty0 1071--1086, 2013.
-	\newblock ISSN 1472-3417.
-	\newblock \doi{10.1068/b38216}.
-
-	\bibitem[Talen(2003)]{talen_measuring_2003}
-	E.~Talen.
-	\newblock Measuring {Urbanism}: {Issues} in {Smart} {Growth} {Research}.
-	\newblock \emph{Journal of Urban Design}, 8\penalty0 (3):\penalty0 195--215,
-	2003.
-	\newblock ISSN 1357-4809.
-	\newblock \doi{10.1080/1357480032000155141}.
-
-	\bibitem[Trudeau(1994)]{trudeau_introduction_1994}
-	R.~J. Trudeau.
-	\newblock \emph{Introduction to {Graph} {Theory}}.
-	\newblock Dover Publications, New York, NY, 2nd edition, 1994.
-	\newblock ISBN 978-0-486-67870-2.
-
-	\bibitem[{U.S. Census Bureau}(2010)]{u.s._census_bureau_2010_2010}
-	{U.S. Census Bureau}.
-	\newblock 2010 {Census} {Urban} and {Rural} {Classification} and {Urban} {Area}
-	{Criteria}, 2010.
-	\newblock URL
-	\url{https://www.census.gov/geo/reference/ua/urban-rural-2010.html}.
-
-	\bibitem[Viana et~al.(2013)Viana, Strano, Bordin, and
-	Barthelemy]{viana_simplicity_2013}
-	M.~P. Viana, E.~Strano, P.~Bordin, and M.~Barthelemy.
-	\newblock The simplicity of planar networks.
-	\newblock \emph{Scientific Reports}, 3\penalty0 (3495):\penalty0 1--6, 2013.
-	\newblock ISSN 2045-2322.
-	\newblock \doi{10.1038/srep03495}.
-
-	\bibitem[Willis(2008)]{willis_openstreetmap_2008}
-	N.~Willis.
-	\newblock {OpenStreetMap} project completes import of {United} {States} {TIGER}
-	data.
-	\newblock \emph{Linux.com}, Jan. 2008.
-	\newblock URL
-	\url{https://www.linux.com/news/openstreetmap-project-completes-import-united-states-tiger-data}.
-
-	\bibitem[Wu et~al.(2005)Wu, Funk, Lurmann, and Winer]{wu_improving_2005}
-	J.~Wu, T.~H. Funk, F.~W. Lurmann, and A.~M. Winer.
-	\newblock Improving spatial accuracy of roadway networks and geocoded
-	addresses.
-	\newblock \emph{Transactions in GIS}, 9\penalty0 (4):\penalty0 585--601, 2005.
-	\newblock ISSN 1467-9671.
-	\newblock \doi{10.1111/j.1467-9671.2005.00236.x}.
-
-	\bibitem[Zhong et~al.(2017)Zhong, Schläpfer, Arisona, Batty, Ratti, and
-	Schmitt]{zhong_revealing_2017}
-	C.~Zhong, M.~Schläpfer, S.~M. Arisona, M.~Batty, C.~Ratti, and G.~Schmitt.
-	\newblock Revealing centrality in the spatial structure of cities from human
-	activity patterns.
-	\newblock \emph{Urban Studies}, 54\penalty0 (2):\penalty0 437--455, 2017.
-	\newblock ISSN 0042-0980.
-	\newblock \doi{10.1177/0042098015601599}.
-
-	\bibitem[Zielstra and Hochmair(2011)]{zielstra_comparative_2011}
-	D.~Zielstra and H.~Hochmair.
-	\newblock Comparative {Study} of {Pedestrian} {Accessibility} to {Transit}
-	{Stations} {Using} {Free} and {Proprietary} {Network} {Data}.
-	\newblock \emph{Transportation Research Record: Journal of the Transportation
-		Research Board}, 2217:\penalty0 145--152, 2011.
-	\newblock ISSN 0361-1981.
-	\newblock \doi{10.3141/2217-18}.
-
-\end{thebibliography}
-
+\end{multicols}
 \end{document}

### 2018-02-15 draft/v3/main.tex
update encodings, fonts, headings

@@
 %-------------------------------------------------------------------------------------------
-% enter the author and paper information below:
+% enter the author and paper information:
 %-------------------------------------------------------------------------------------------
@@
 
-% import these packages first, in order
+% import encoding and font packages for pdflatex, in order
+\usepackage[T1]{fontenc}
 \usepackage[utf8]{inputenc}
+\usepackage{crimson}
+\usepackage{helvet}
+
+% import language and regionalization packages, in order
 \usepackage[strict,autostyle]{csquotes}
 \usepackage[USenglish]{babel}
+\usepackage{microtype}
 
@@
 \usepackage{booktabs}
-\usepackage{crimson}
-\usepackage{datetime}
 \usepackage[final]{draftwatermark}
@@
 \usepackage{hyperref}
-\usepackage{microtype}
 \usepackage{multicol}
@@
 \usepackage{setspace}
+\usepackage{titlesec}
 \usepackage{url}
 
-% location of figure files
+% location of figure files, via graphicx package
 \graphicspath{{./figures/}}
 
-% configure the page layout
+% configure the page layout, via geometry package
 \geometry{
@@
     right=2.5cm}
-\setlength{\columnsep}{0.75cm} % space between text columns
+\setlength{\columnsep}{0.75cm} % space between columns
 \setstretch{1.00}              % line spacing
@@
 \widowpenalty=10000            % prevent widows
-\let\footnote=\endnote         % turn footnotes into endnotes
+
+% set section/subsection headings as the sans serif font
+\titleformat{\section}{\normalfont\sffamily\large\bfseries\color{black}}{\thesection.}{0.3em}{}
+\titleformat{\subsection}{\normalfont\sffamily\small\bfseries\color{black}}{\thesubsection.}{0.3em}{}
 
@@
 OpenStreetMap offers an underexplored source of worldwide geospatial data useful to urban researchers.
-This study uses the OSMnx software to automatically download and analyze 27,000 US street networks from OpenStreetMap at metropolitan, municipal, and neighborhood scales---namely, every US city and town ($N=19655$), census urbanized area ($N=497$), and Zillow-defined neighborhood ($N=6857$).
-It presents wide-ranging empirical findings on US urban form and street network characteristics, emphasizing measures relevant to graph theory, urban design, and morphology such as structure, connectedness, density, centrality, and resilience.
-In the past, street network data acquisition and processing has been challenging and ad hoc.
+This study uses the OSMnx software to automatically download and analyze 27,000 US street networks from OpenStreetMap at metropolitan, municipal, and neighborhood scales---namely, every US city and town, census urbanized area, and Zillow-defined neighborhood.
+It presents empirical findings on US urban form and street network characteristics, emphasizing measures relevant to graph theory, transportation, urban design, and morphology such as structure, connectedness, density, centrality, and resilience.
+In the past, street network data acquisition and processing have been challenging and ad hoc.
 This study illustrates the use of OSMnx and OpenStreetMap to consistently conduct street network analysis with extremely large sample sizes, with clearly defined network definitions and extents for reproducibility, and using nonplanar, directed graphs.
-These street networks and measures data sets have been shared in a public repository for other researchers to use.
+These street networks and measures data have been shared in a public repository for other researchers to use.
 \end{abstract}
@@
 
+% start two-column layout after the title and abstract
 \begin{multicols}{2}
@@
 It examines the street networks of every incorporated city and town (population size $N=19655$), census urbanized area ($N=497$), and Zillow-defined neighborhood ($N=6857$) in the US.
-To do so, it uses OSMnx\footnote{OSMnx is freely available online at \url{https://github.com/gboeing/osmnx}}---a new street network research toolkit \citep{boeing_osmnx:_2017}---to download, construct, and analyze these street networks at metropolitan, municipal, and neighborhood scales.
-These street networks and measures data sets have been compiled and shared in a public repository at the Harvard Dataverse\footnote{\label{note_data_repo}Data repository available online at \url{https://dataverse.harvard.edu/dataverse/osmnx-street-networks}} for other researchers to repurpose.
+To do so, it uses OSMnx\endnote{OSMnx is freely available online at \url{https://github.com/gboeing/osmnx}}---a new street network research toolkit \citep{boeing_osmnx:_2017}---to download, construct, and analyze these street networks at metropolitan, municipal, and neighborhood scales.
+These street networks and measures data sets have been compiled and shared in a public repository at the Harvard Dataverse\endnote{\label{note_data_repo}Data repository available online at \url{https://dataverse.harvard.edu/dataverse/osmnx-street-networks}} for other researchers to repurpose.
 
@@
 
+% switch back to one-column layout to rotate this wide table sideways
 \end{multicols}
-
 \begin{sidewaystable*}[p]
@@
 
+% done with wide rotated table, switch back to two columns
 \begin{multicols}{2}
